### 2021-04-27 manuscript.Rmd
initial commit

---
title             : "Sequential Hypothesis Testing: A Comparison of Group-Sequential Designs, the Sequential Bayes Factor, and the Independent Segments Procedure"
shorttitle        : "Sequential Hypothesis Testing"

author: 
  - name          : "Shilaan Alzahawi"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "655 Knight Way, Graduate School of Business, Stanford, CA 94305"
    email         : "shilaan@stanford.edu"

affiliation:
  - id            : "1"
    institution   : "Stanford University, Graduate School of Business"

note: "\\clearpage"

author_note:

abstract: >
  Sequential analysis allows researchers to perform well-powered experiments with smaller sample sizes. In this practical guide, I outline the commonalities and differences between three sequential hypothesis testing procedures: group-sequential designs, the sequential Bayes Factor, and the independent segments procedure. Using Monte-Carlo simulations, I compare the performance of these procedures in terms of their error control (i.e., false-negative and false-positive decision rates), efficiency (i.e., reduction in required sample size), and accuracy (i.e., bias and variance in effect size estimation). I conclude with practical recommendations for researchers who wish to utilize the benefits of sequential hypothesis testing.

keywords          : "sequential analysis, hypothesis testing, efficiency, effect size estimation, bias"
wordcount         : "-"

bibliography      : ["better.bib","r-refs.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
numbersections    : no
linenumbers       : no #add line numbers in the margins
mask              : no  #omit identifying info from the title page
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
knitr::opts_chunk$set()
library(papaja)
library(tidyverse)
library(ggplot2)
library(grid)
library(rpact)
library(pracma)
library(coda)
library(BayesFactor)
library(parallel)
library(metafor)
library(beepr)

r_refs(file = "r-refs.bib") #update later
default_plot_hook = knitr::knit_hooks$get('plot')

knitr::knit_hooks$set(plot = function(x, options) { #function to prevent pandoc issues when plotting
  default_md <- default_plot_hook(x, options)
  link_attr <- sprintf("{width=%sin height=%sin}", options$fig.width, options$fig.height)
  sub("(!\\[.*]\\(.*\\))", paste0("\\1", link_attr), default_md)
})
```

In 1929, Harold Dodge and Harry Romig introduced the idea of sequential analysis to the science of statistical quality control [@dodge1929]. Sequential analysis is a method of statistical inference that allows data to be analyzed several times during its collection, with the goal of optimizing the amount of resources spent in research [@beffarabret2018]. When Abraham Wald further developed this idea in the 1940s, his work was deemed so useful to the United States military that it was classified as Restricted under the Espionage Act [@wald1973].

Fast-forward 80 years, and we find ourselves in a decade of unparalleled methods reform in psychological science [@nelson2018]. At the receiving end of these proposed reforms we often find the statistical power of hypothesis tests [@bakker2012; @button2013]; that is, the long-run probability of observing a statistically significant result ($p < \alpha$) for a given number of subjects and a hypothesized effect size [@morey2016].

Several suggestions have been made to increase the statistical power of hypothesis tests conducted in psychological science. Not surprisingly, sequential analysis --- which allows researchers to perform well-powered experiments at lower cost --- has found its way into the limelight [e.g., @beffarabret2018; @lakens2014; @schonbrodt2017; @schnuerch2020; @miller2020].

While research efficiency is commendable [although not indisputably so; @tiokhin2020], it is not the only advantage conferred by sequential hypothesis testing. Under the Ethics Code of the American Psychological Association, psychological scientists have the ethical obligation to minimize research subjects' exposure to physical, emotional, or psychological harm [@apa2020]. Sequential analysis can contribute to this goal, by allowing researchers to stop an experiment when the interim data show clear evidence that effects are positive, negative, or largely absent [@proschan2006]. During clinical trials, stopping early can be a matter of life or death. Sequential analysis, it turns out, is just as useful today as it was during World War II.

## Sequential procedures

In the following section, I briefly outline the sequential procedures under discussion, highlighting several of their commonalities and differences. Next, I introduce the desirable properties of interest to this paper: error control, efficiency, and accuracy. By means of simulation, I will compare the sequential procedures to traditional, fixed-sample hypothesis testing, in terms of false-negative and false-positive error rates, sample size reduction, and the bias and variance involved in effect size estimation.

```{r, include = FALSE}
n_fixed = ceiling(power.t.test(delta = 0.5, sig.level = .05, power = 0.8, 
                                 type = "two.sample", alternative = "one.sided")$n)

gs = function(proc, return = "n") {
  bs = ifelse(proc == "asOF", "bsOF", "bsP")

  # Specify the design
  design = getDesignGroupSequential(sided = 1,
                                    alpha = 0.05, 
                                    beta = 0.2,
                                    kMax = 3, 
                                    typeOfDesign = proc,
                                    typeBetaSpending= bs)

  # Get parameters
  parameters = getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
  n_gs = ceiling(parameters$maxNumberOfSubjects/3/2) # n per look per group
  alpha = parameters$criticalValuesPValueScale
  futility = parameters$futilityBoundsPValueScale %>% round(2)

  if(return == "alpha") {return(alpha %>% signif(2) %>% as.character() %>% sub(".", "", .,))}
  if(return == "n") {return(n_gs)}
  if(return == "futility") {return(futility %>% as.character() %>% sub(".", "", .,))}
}

n_p = gs(proc = "asP")
alpha_p = gs(proc = "asP", return = "alpha")
futility_p = gs(proc = "asP", return = "futility")
n_of = gs(proc = "asOF")
alpha_of = gs(proc = "asOF", return = "alpha")
futility_of = gs(proc = "asOF", return = "futility")
```

## *Group-sequential procedures* 
The most common procedure for sequential hypothesis testing is a group-sequential (GS) design [@proschan2006; @schonbrodt2017]. The goal of GS procedures is to efficiently reject a null hypothesis in favor of an alternative hypothesis. In a prototypical group-sequential procedure [e.g., the Pocock correction or the O'Brien-Fleming correction, @pocock1977; @fleming1984], a researcher is allowed to look at the data several times during the course of a study.

It is well-known that multiple looks at data, conventionally, increase the type I error rate [@simmons2011]: i.e., the long-run rate at which a true null hypothesis is rejected. Sequential analysis procedures, however, allow a researcher to take multiple looks at data without inflating the type I error rate. Critical $\alpha$ levels are adjusted for each look so that the overall type I error rate is controlled at the desired level (for instance, at an $\alpha$ of 5%).

A general, sophisticated approach is provided by an "$\alpha$*-spending function*" [@lan1983; @wassmer2016]. This approach allows a researcher to have unequally spaced looks at their data (for instance, after the first 20% and the first 75% of the maximum sample size) and to adapt the number of looks as the experiment continues.

For example, a researcher who wishes to control their overall type I error rate at 5% and take $k_{max}$ = 3 looks (i.e., two interim looks and one final look) at the data can use the common Pocock-like $\alpha$-spending function to determine what the corrected $\alpha$ level should be for each look (referred to as $\alpha_{1,k}$). In this case, $\alpha_{1,1:2}$ = `r alpha_p[1]` for the first two looks and $\alpha_{1,3}$ = `r alpha_p[3]` for the last look (i.e., for a one-tailed test). If during any of the interim looks $p \le \alpha_1$, the researcher can reject the null hypothesis and end the experiment; this is known as "*stopping for efficacy*".

Because the use of group-sequential procedures allows for early stopping, it generally reduces the average sample size needed (with certain exceptions; e.g., when the true effect size is equal to or close to zero, the use of sequential procedures increases the average sample size needed). For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_+:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null-Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group. Using the Pocock-like GS design outlined above, the same study can be conducted with only `r n_p` participants per group per look.

In addition to stopping early for efficacy, a researcher can "*stop for futility*" if the observed effect size is too small to warrant further investigation. This can be done by manually specifying a lower bound (for example, a lower bound at *d* $\le$ 0 or, equivalently, at *p* $\ge$ .50) or by using a "$\beta$*-spending function*" to determine the critical lower bound for each look (referred to as $\alpha_{0,k}$). In the remainder of the text, I will use $\alpha_1$ to refer to the critical value that allows a researcher to stop for efficacy, and $\alpha_0$ to refer to the critical value that allows a researcher to stop for futility [@wassmer2016].

The logic of the $\beta$-spending function is similar to that of the $\alpha$-spending function: critical values that warrant stopping for futility are determined a priori while controlling the overall type II error rate at the desired level (for instance, at a $\beta$ of 20%). In other words, the bounds are determined such that the probability of a type II error (i.e., given the number of subjects, the long-run probability of failing to reject the null when the true population effect $\delta$ = 0.5) is $\beta$ %.

Figures 1A and 1B provide a visual overview of two possible group-sequential procedures (the Pocock-like and the O'Brien-Fleming-like designs) using $\alpha$- and $\beta$-spending functions, in the case of a one-tailed, two-sample *t*-test with $\alpha$ = 5%, $\beta$ = 20% for $\delta$ = 0.5, and $k_{max}$ = 3 looks. The main difference between the Pocock and the O'Brien-Fleming-like procedure is that the O'Brien-Fleming-like procedure sets a stricter $\alpha$ threshold in earlier looks, with the last look having a threshold closer to the overall $\alpha$. Consider the $\alpha$ levels for our example Pocock-like design described above: $\alpha_{1,1:2}$ = `r alpha_p[1]` for the first two looks and $\alpha_{1,3}$ = `r alpha_p[3]` for the last look. For the O'Brien-Fleming-like procedure, the critical levels are $\alpha_{1,1}$ = `r alpha_of[1]` for the first look, $\alpha_{1,2}$ = `r alpha_of[2]` for the second look, and $\alpha_{1,3}$ = `r alpha_of[3]` for the last look.

The futility bounds for our example Pocock-like design are $\alpha_{0,1}$ = `r futility_p[1]` for the first look and $\alpha_{0,2}$ = `r futility_p[2]` for the second look. For the O'Brien-Fleming-like design, the futility bounds are $\alpha_{0,1}$ = `r futility_of[1]` for the first look and $\alpha_{0,2}$ = `r futility_of[2]` for the second look. The researcher continues their study until the observed *p*-value crosses one of the prespecified $\alpha_1$ or $\alpha_0$ thresholds, or until the maximum *N* has been reached. Below, I briefly outline the decision-theoretic procedure of GS designs.

**Decision-theoretic procedure of GS designs.**

  *
1. Randomly sample $N_{gs}$ observations from the specified population(s) 
2. Run the test(s) of interest (e.g., a one-tailed, two-sample *t* test) on the obtained sample data
3. During the interim looks (1 to $k_{max} -1$),

  a. if $p \le \alpha_1$: decide to reject $H_0$ and end the experiment ("stop for efficacy")
  b. if $p > \ \alpha_0$: decide not to reject $H_0$ and end the experiment  ("stop for futility")
  c. if $\alpha_1 < p~\le~\alpha_0$: collect $N_{gs}$ additional observations and repeat step 2 on the cumulative data

4. During the final look $k_{max}$ (if reached),
  a. if $p \le \alpha_1$: decide to reject $H_0$ and end the experiment
  b. if $p > \ \alpha_1$: decide not to reject $H_0$ and end the experiment

## *Sequential Bayes Factor* 
@schonbrodt2017 propose an alternative approach to sequential hypothesis testing: the Sequential Bayes Factor (SBF). With the SBF procedure, decisions are made on the basis of interim Bayes Factors. The Bayes Factor is a measure of relative evidence for two competing hypotheses [e.g., the null and the alternative hypothesis; for an excellent introduction to Bayesian Hypothesis Testing, see @wagenmakers2018].

The SBF procedure has several properties of interest. First, it is symmetric: it allows researchers to examine whether the obtained data is better predicted by, for example, $H_+$ than $H_0$, or vice versa. Second, it is highly flexible: it allows researchers to take as many looks at incoming data as desired [even after every study participant; @schonbrodt2017]. Third, it does not allow type I and type II error rates to be defined a priori [although simulations can be run to choose an optimal design that controls expected error rates, @schonbrodt2017; @schonbrodt2018; @stefan2020]. Fourth, it allows (and requires) researchers to express their uncertainty about the expected effect size by means of a prior distribution.

Again, we consider the example of a researcher running a one-sided, two-sample *t* test between $H_0:~\delta$ = 0 and $H_+:~\delta > 0$. To follow the SBF procedure, the researcher needs to prespecify two things: (a) the prior distribution for the standardized effect size $\delta$ under $H_+$ and (b) the decision threshold _t_* for the Bayes Factor. Excellent guidance on these important decisions can be found in @schonbrodt2017. Ideally, a Bayes Factor Design Analysis or simulations can be run to choose an optimal design that controls expected error rates [@schonbrodt2018].

To reflect the expectation of a "medium" effect size, the researcher can, for example, choose the default prior distribution [i.e., a Cauchy distribution with scale parameter *r* = $\sqrt2/2$; @morey2011; @schnuerch2020; @schonbrodt2017]. In addition, the researcher could (for example, on the basis of simulations) choose $BF_{+0}$ = 3 as a decision threshold [commonly referred to as "*moderate evidence*"; @schonbrodt2017]. This decision threshold _t_* implies that the researcher is content to stop sampling once the data is 3 times more likely under $H_+$ than under $H_0$, or vice versa. Figure 2A (2B) provides a visual example of a situation in which the Bayes Factor indicates that the data is 3 times more likely under $H_+$ ($H_0$).

The Sequential Bayes Factor and the Group-Sequential Procedures address different questions, each interesting in their own right. Group-sequential procedures (as well as the traditional Null Hypothesis Significance Test and the Independent Segments Procedure described below) raise the question "*On the basis of a decision rule under which the long-run probability of rejecting a true null hypothesis will be no more than $\alpha$%, can the null hypothesis be rejected?*" The Sequential Bayes Factor procedure, on the other hand, considers "*On the basis of a prespecified threshold and desired error rates under the assumption of a specific $H_0$ and $H_+$, can we conclude that there is sufficient evidence for $H_0$ or $H_+$?*"

There are several possible ways to implement the SBF procedure [@schonbrodt2018]. An open-ended SBF design is one in which a researcher tests after each participant and stops data collection when sufficient evidence for either $H_1$ or $H_0$ has been gathered. Alternatively, a researcher can prespecify a maximum sample size and continue data collection until (a) sufficient evidence has been gathered or (b) the maximum sample size has been reached. In this paper, I focus on the latter procedure: a max-SBF design. Below, I briefly outline the decision-theoretic procedure of the max-SBF.

**Decision-theoretic procedure of the Sequential Bayes Factor with maximum N.**

  *
1. Randomly sample $N_{Bayes}$ observations from the specified population(s)
2. Run the test(s) of interest (e.g., a one-sided default Bayes Factor two-sample *t* test)
  a. if $BF_{+0} \ge t*$: decide to reject $H_0$ (support $H_+$) and end the experiment
  b. if $BF_{+0} \le 1/t*$: decide to reject $H_+$ (support $H_0$) and end the experiment
  c. if $1/t*~< BF_{+0} < t*$: collect additional observations and repeat step 2 on the cumulative data
3. Continue until the Bayes Factor crosses one of the thresholds, or until the maximum *N* has been reached

## *Independent segments procedure* 
Another sequential procedure recently proposed for use in psychological research is the Independent Segments Procedure [ISP, @miller2020]. A central property of the ISP is that data is collected in a series of independent experiments ('segments'). Using other sequential procedures (such as the GS design and the Sequential Bayes Factor), data collection and analysis is cumulative: when an experiment is continued (because the interim data did not cross the pre-specified boundaries for early stopping), a researcher collects more observations, adds these to the observations from previous looks, and calculates an interim test statistic on the basis of all the data collected so far. In contrast, when an independent segments experiment is continued, the researcher discards the previous batch of observations and calculates an interim test statistic on the basis of a new, completely independent, batch of observations. Unlike the SBF and GS-designs, the ISP ignores data from earlier looks. The ISP, by design, incorporates information loss.

In addition, the ISP does not have the same flexibility as the procedures described above, in two important regards. First, unlike the alternative research strategies discussed in this paper, the ISP requires that a researcher sets the exact number of participants and segments in advance, and all segments are assumed to be equally spaced (i.e., when $k_{max}$ = 2 segments, the first look at the data occurs when exactly 50% of the maximum sample size has been collected). Thus, the ISP will be less appropriate for researchers with practical constraints who, for example, cannot guarantee the exact same sample size in each respective segment. Again, consider our example of a one-tailed, two-sample *t* test with $k_{max}$ = 3 segments. Figures 3A and 3B provide a visual overview of the independent segments procedure, compared to the conventional fixed-sample Neyman-Pearson hypothesis test (Figure 3A shows the conceptual differences; Figure 3B shows our worked example). In order to achieve 80% power with an overall $\alpha$ of 5% and a hypothesized population effect size $\delta$ = 0.5, the required sample size $N_{ISP}$ = 25 per group per segment. If the researcher deviates in any way from having exactly 25 participants per group per segment, inferences on the basis of the ISP will be affected. Although this influence may generally be small for practical purposes, there are cases in which the effects of violating an assumption of equally sized stages are non-negligible [@wassmer2016].

Second, the ISP only applies to research contexts in which a single, key hypothesis is tested [@miller2020]. This, unfortunately, is unusually restrictive to the settings in which sequential hypothesis testing is common. For example, medical trials often have multiple "*endpoints*" that need to be monitored at the same time: a primary endpoint (e.g., treatment success or survival) and multiple secondary -- but exceedingly important -- efficacy or safety endpoints [@wassmer2016]. An enlightening example can be found in the FDA approval process for carvedilol (a drug developed for congestive heart failure), in which the endpoints were -- among other things -- mortality, distance that can be walked on a treadmill, exercise tolerance, cardiovascular hospitalizations, quality of life, and the patient and physician's global assessment [@fisher1999; @fisher1999a].

In sum, the ISP applies only to those settings in which a researcher is comfortable with discarding inconclusive data; can guarantee an exact number of participants, allocated equally across segments; and wishes to test a single, key hypothesis. Below, I briefly outline the decision-theoretic procedure of the ISP.

**Decision-theoretic procedure of the ISP.**

  *
1. Randomly sample $N_{ISP}$ observations from the specified population(s)
2. Run the test(s) of interest (e.g., a one-tailed, two-sample *t* test) on the obtained sample data
3. For segments 1:$k_{max} -1$,

  a. if $p \le \alpha_{strong}$: decide to reject $H_0$ and end the experiment 
  b. if $p > \alpha_{weak}$: decide not to reject $H_0$ and end the experiment
  c. if $\alpha_{strong} < p \le \alpha_{weak}$: discard the data, collect a new batch of $N_{ISP}$ observations, and repeat step 2 on this independent batch of data

4. For segment $k_{max}$ (if reached), repeat steps 1 and 2,
  a. if $p \le \alpha_{weak}$: decide to reject $H_0$ and end the experiment
  b. if $p > \alpha_{weak}$: decide not to reject $H_0$ and end the experiment

Table 1 summarizes a selection of commonalities and differences between the sequential hypothesis testing procedures under discussion.

## Common trade-offs in the scientific research process

The scientific research process often involves mutually incompatible goals [@mcgrath1981]. At times, scientific goals that are each worthy in their own right can conflict: optimizing for one goal tends to decrease our ability to achieve the other [e.g., @goodman2007]. In the following section, I introduce two such pairs of scientific goals. First, I elaborate on the trade-off between controlling errors in the long-run versus obtaining compelling evidence for the case at hand. Second, I describe the trade-off between efficiently testing hypotheses versus accurately estimating effect sizes. In addition, I explain why these trade-offs (long-run error control versus short-run evidence and efficiency versus accuracy, respectively) are pertinent to our discussion of sequential hypothesis testing.

## *Long-run error control versus short-run evidence*

"*Today, I speak to you of war. A war that has pitted statistician against statistician for nearly one hundred years. A mathematical conflict that has recently come to the attention of the normal people. And these normal people look on in fear, in horror, but mostly in confusion because they have no idea why we're fighting*" [@lawrencelivermorenationallaboratory2016]. Such was the introduction of statistician Kristin Lennox to a talk at the Lawrence Livermore National Laboratory. The war of which she speaks is the war between Bayesian and frequentist statistics.

In 1933, Jerzy Neyman and Egon Pearson introduced an ingenious device for testing competing hypotheses [@neyman1933]. The goal of Neyman-Pearson hypothesis testing is to decide between two competing hypotheses with explicit control over two types of error, introduced earlier in this paper: type I error (the long-run rate at which a true null hypothesis is rejected) and type II error (for a given nonzero population effect size, the long-run rate at which the null hypothesis is maintained).  By definition, this is a frequentist procedure: the goal is to devise a decision rule for choosing between competing hypotheses under which the decision maker, in the long run (i.e., when considering the frequency of all possible outcomes), will not be wrong too often [@neyman1933; @goodman1999].

This ingenious device, however, comes at a price that its users should be willing to pay. The outcome of the Neyman-Pearson hypothesis test is a behavior---a decision to reject or accept a hypothesis. This behavior has no short-run (i.e., experiment-specific) meaning: it is not an inference about the relative evidence for competing hypotheses provided by the experiment at hand. According to the Neyman-Pearson standpoint, it is impossible to provide any (relative) evidence of the truth or falsity of the hypotheses under study [@neyman1933]. All we can say is that a decision was made on the basis of a rule that, in the long-run, controls error probabilities [@dienes2011]. To enjoy the benefits of error control provided by Neyman-Pearson hypothesis testing, "[w]e must abandon our ability to measure evidence, or judge truth, in an individual experiment" [@goodman1999: p. 998].

It is at this junction --- between the ability to measure evidence in a single experiment and the ability to control the number of mistaken conclusions in the long-run --- that we meet the Bayesian statistician. According to the Bayesian, evidence is a property of the obtained data that makes us alter our belief about the hypotheses at hand. The Bayesian measure of evidence, previously encountered by the reader, is the Bayes Factor. There is, however, no free lunch in statistical inference [@rouder2016]. Because the Bayesian calculation considers no outcomes other than the one observed, robust error control can no longer be ensured [@dienes2011; @kruschke2018]. The consideration of outcomes that *could* be observed (i.e., the tail area of the sampling distribution) under a hypothesis of interest is essential to ensuring a test's long-run performance and associated error probabilities [@mayo2018].

To some, the use of tail areas in hypothesis testing is an admirable feature that ensures long-run error control; to others, it is a bug that defies our ability to articulate the evidence for competing hypotheses provided by the experiment at hand. Controlling (long-run) error rates and quantifying (short-run) evidential strength are both worthy scientific goals. However, you can't have one without implicitly abstaining from the other [unless you follow the Likelihood paradigm of statistical inference, which is beyond the scope of this paper; @blume2002; @goodman1988; @schnuerch2020]. Ultimately, the question is what is more important to the researcher [@lakens2020a]: controlling long-run error rates or quantifying short-run evidential strength [@dienes2011].

Why does this matter to our discussion of sequential analysis? It may seem that we have strayed far from the comparison of sequential hypothesis testing procedures. However, we are right at its core. As follows from the above discussion, the Sequential Bayes Factor -- unlike a GS design or the ISP -- is not guaranteed to control type I and type II error probabilities. Thus, it is interesting to evaluate the practical implications of the philosophical differences between frequentist and Bayesian hypothesis testing. I will return to this issue in the following section of the paper, in which I compare the error rates of the Sequential Bayes Factor to the fixed-sample Neyman-Pearson hypothesis test, the Pocock-like and O'Brien-Fleming like group-sequential procedures, and the independent segments procedure.

## *Efficiency versus accuracy*

Of special interest to our discussion of sequential analysis, a trade-off exists between efficiently testing hypotheses and accurately estimating an effect size [@schonbrodt2017]. Accurate parameter estimation often requires more observations than a (sequential) hypothesis tester is willing or able to collect. Efficiency and accuracy are each worthy scientific goals in their own right and, unfortunately, often conflict [@goodman2007].

The goal of sequential procedures is to efficiently test hypotheses; the goal is not to properly estimate effect sizes [@lakens2020; @schonbrodt2017]. It is well known that early stopping can seriously inflate effect sizes [@fan2004; @miller2020; @pocock1989]. However, when taking all studies into account (including those that did not terminate early), the bias inherent in both group-sequential designs [@goodman2007; @fan2004] and Sequential Bayes Factors [@schonbrodt2017] is small. Thus, group-sequential designs and Sequential Bayes Factors can increase research efficiency at only a slight cost to our ability to estimate the true effect size. No such investigation, however, has been conducted for the Independent Segments Procedure [@miller2020].

In the following section, I use Monte-Carlo simulations to compare group-sequential designs, Sequential Bayes Factors, and the Independent Segments Procedure to the fixed-sample Neyman-Pearson procedure, focusing on their relative error rates, efficiency, and accuracy. In a previous contribution, the Sequential Bayes Factor was compared to GS designs in terms of error rates, efficiency, and accuracy [@schonbrodt2017]. However, these comparisons were unbalanced in two regards: (a) the GS designs used did not allow for early stopping (i.e., they did not incorporate, for instance, a $\beta$-spending function to stop for futility) and (b) the number of looks was not equal across procedures (i.e., the Sequential Bayes Factor procedure was based on a near-unlimited number of looks until the decision threshold was met, while the chosen GS design incorporated four looks).

@schonbrodt2017 show that a highly flexible implementation of the Sequential Bayes Factor -- i.e., the SBF in all its glory, with as many looks as is desired -- is at least as efficient as a default group-sequential design (i.e., a group-sequential approach that does not incorporate binding futility bounds on the basis of a $\beta$-spending approach). In this paper, I implement the SBF with $k_{max}$ equally spaced looks, for two reasons: (a) to allow for maximum comparability between the sequential procedures under discussion, controlling for as many confounds as possible; and (b) to account for practical considerations: it may very well be impractical for researchers to implement the maximum-flexibility SBF approach and analyze data after every participant. Thus, I focus on a slightly different, pragmatic question: For a fixed number of equally spaced looks at data, how do these respective procedures perform in terms of their error rates, efficiency, and accuracy?

# Method

By means of simulation, I will demonstrate the properties of the procedures under discussion in the context of testing hypotheses about positive mean differences between two independent groups (i.e., a one-tailed, two-sample *t*-test). A one-tailed test was chosen to allow for comparability with the ISP, which has been developed for one-tailed hypothesis testing only. The procedures will be compared in terms of three main properties: (a) error rates, (b) efficiency, and (c) accuracy in effect size estimation. In the following section, I briefly outline how each property is evaluated.

## Properties of Interest: Error Rates, Efficiency, and Accuracy

Decision-theoretic error rates are evaluated by classifying test results as True Positive, True Negative, False Positive, False Negative, or Inconclusive. Below, I outline the cases in which the outcomes of a hypothesis test were labelled as True Positive, True Negative, False Positive, False Negative, or Inconclusive, respectively.

**True Positive. **

  *

1. For the fixed-sample hypothesis test, $\delta$ > 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha$)
2. For the Pocock and O'Brien-Fleming procedures, $\delta$ > 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha_{1,k}$)
3. For the SBF procedure, $\delta$ > 0 and the decision was made to reject the null hypothesis ($BF_{+0}~\ge$ _t_*)
4. For the Independent Segments Procedure, $\delta$ > 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha_{strong}$ in segments 1:$k_{max}-1$ or *p* $\le \alpha_{weak}$ in the final segment $k_{max}$)

**False Positive. **

  *

1. For the fixed-sample hypothesis test, $\delta$ = 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha$)
2. For the Pocock and O'Brien-Fleming procedures, $\delta$ = 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha_{1,k}$)
3. For the SBF procedure, $\delta$ = 0 and the decision was made to reject the null hypothesis ($BF_{+0} \ge$ _t_*)
4. For the Independent Segments Procedure, $\delta$ = 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha_{strong}$ in segments 1:$k_{max}-1$ or *p* $\le \alpha_{weak}$ in the final segment $k_{max}$)

**True Negative. **

  *

1. For the Pocock and O'Brien-Fleming procedures, $\delta$ = 0 and the decision was made to stop for futility (*p* $\ge \alpha_{0,k}$)
2. For the SBF procedure, $\delta$ = 0 and the decision was made to support the null hypothesis ($BF_{0+} \ge$ _t_*)
3. For the Independent Segments Procedure, $\delta$ = 0 and the decision was made to stop in an early segment (*p* $> \alpha_{weak}$)

**False Negative. **

  *

1. For the Pocock and O'Brien-Fleming procedures, $\delta$ > 0 and the decision was made to stop for futility (*p* $\ge \alpha_{0,k}$)
2. For the SBF procedure, $\delta$ > 0 and the decision was made to support the null hypothesis ($BF_{0+} \ge$ _t_*)
3. For the Independent Segments Procedure, $\delta$ > 0 and the decision was made to stop in an early segment (*p* $> \alpha_{weak}$)

**Inconclusive. **

  *

1. For the fixed-sample hypothesis test, the procedure ended in a failure to reject the null hypothesis 
2. For the Pocock and O'Brien-Fleming procedures, $k_{max}$ was reached and neither of the thresholds was met ($\alpha_1 < p~\le~\alpha_0$)
3. For the SBF procedure, $k_{max}$ was reached and neither of the thresholds was met ($1/t*~< BF_{+0} < t*$)
4. For the Independent Segments Procedure, $k_{max}$ was reached and ended in a failure to reject the null hypothesis (*p* > $\alpha_{weak}$)

Conventionally, the Null-Hypothesis Significance test is not able to distinguish between failures to reject the null hypothesis due to evidence of absence (a True Negative result) or absence of evidence (an Inconclusive result). Therefore, we labeled all failures to reject the null on the basis of a fixed-sample hypothesis test as Inconclusive. It will be noted, however, that it is possible to overcome this limitation of conventional practice by testing against range predictions rather than a nil-null hypothesis. Instead of testing the nil-null hypothesis that the effect is exactly zero, researchers can test whether effects more extreme than a range of predicted or practically important effect sizes can be rejected -- for example, by virtue of an equivalence test or a minimum-effect test [@lakens2018; @lakens2020a].

Procedural efficiency is evaluated by means of the average (expected) sample size across several true effect sizes for each procedure. Accuracy in effect size estimation is evaluated by means of the density of empirical effect size estimates, the bias in effect size estimates (i.e., the median estimated effect size compared to the true effect size), the mean squared error of effect size estimates (i.e., the sum of the squared bias and the variance of effect size estimates), and the meta-analytic effect size estimate derived from each procedure.

## Settings of the Simulation

I simulate populations with a specific standardized mean difference $\delta$ and examine the efficiency (i.e., average sample size required), error rates (i.e., rates of false-positive and false-negative evidence), and accuracy (i.e., of the simulation estimate of the population effect size) of five hypothesis testing procedures: the fixed-sample Neyman-Pearson hypothesis test, the independent segments procedure [@miller2020], a group-sequential design with Pocock-like $\alpha$- and $\beta$-spending functions [@pocock1977; @wassmer2016], a group-sequential design with O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions [@fleming1984; @wassmer2016], and the sequential Bayes Factor [@schonbrodt2017].

To facilitate discussion, I focus on one typical scenario in reporting the simulation results. Full results that vary additional parameters are reported in the online supplementary material, and reproducible analysis scripts are available at [link]. For the frequentist procedures (i.e., the fixed-sample N-P procedure, the ISP, and the Pocock-like and O'Brien-Fleming-like GS designs, respectively) under discussion, the chosen typical scenario corresponds to a one-tailed, two-sample hypothesis test powered to 80% ($\beta$ = 0.2) to detect a hypothesized population effect size of $\delta$ = 0.5 with $\alpha$ = .05. In the fixed-sample case, these specifications require a sample size of `r n_fixed` observations per group.

For maximum comparability, all sequential procedures (ISP, Pocock-like GS design, O'Brien-Fleming-like GS design, and the SBF) were set to a maximum number of 3 equally spaced looks. For the independent segments procedure with $\alpha_{total}$ = .05, $\alpha_{strong}$ = .025, $\alpha_{weak}$ = .28, $\beta_{total}$ = .2 and $k_{max}$ = 3 segments, the required sample size $N_{ISP}$ = 25 per group per segment.

For the GS designs with $\alpha_{total}$ = .05, $\beta_{total}$ = .20, $k_{max}$ = 3 looks, an $\alpha$-spending approach was used to determine critical $\alpha_{1,k}$ levels. For the example discussed here, the adjusted $\alpha_{1,k}$ levels are $\alpha_{1,1}$ = `r alpha_p[1]`, $\alpha_{1,2}$ = `r alpha_p[2]`, and $\alpha_{1,3}$ = `r alpha_p[3]` at the first, second, and third look for the Pocock-like design and $\alpha_{1,1}$ = `r alpha_of[1]`, $\alpha_{1,2}$ = `r alpha_of[2]`, and $\alpha_{1,3}$ = `r alpha_of[3]` at the first, second, and third look for the O'Brien-Fleming-like design.

In addition, a $\beta$-spending function was used to set critical $\alpha_{0,k}$ levels (i.e., binding futility bounds). For our worked example, the adjusted $\alpha_{0,k}$ levels correspond to $\alpha_{0,1}$ = `r futility_p[1]` and $\alpha_{0,2}$ = `r futility_p[2]` at the first and second look for the Pocock-like design and $\alpha_{0,1}$ = `r futility_of[1]` and $\alpha_{0,2}$ = `r futility_of[2]` at the first and second look for the O'Brien-Fleming-like design. For the Pocock-like design, the required sample size $N_{pocock}$ = `r n_p` per group per look; for the O'Brien-Fleming correction, the required sample size $N_{O'Brien-Fleming}$ = `r n_of` per group per look.

For the SBF, I ran simulations to choose an optimal sample size (with $k_{max}$ = 3 equally spaced looks), Bayes Factor threshold, and prior distribution, such that the overall error rate would approach that of the frequentist procedures (i.e., a false negative rate of 20% and a false positive rate of 5%). On the basis of these simulations, the minimum sample size was set to $N_{Bayes}$ = 22 per group per look with $k_{max}$ = 3 looks, and the Bayes Factor threshold was set to 3 for $H_+$ and 1/3 for $H_0$.

Of note, as discussed above, the Sequential Bayes Factor can be implemented much more flexibly (i.e., with as many looks as desired, without a maximum sample size). However, to make the sequential procedures as comparable as possible (and to account for practical considerations that researchers may have), each procedure is each implemented with $k_{max}$ = 3 equally spaced looks.

To calculate the Bayes Factor, I used the one-tailed default Bayes Factor for *t* tests [@morey2011] to compare $H_0:~\delta = 0$ to $H_+:~\delta > 0$. A default prior distribution was chosen for the standardized effect size $\delta$ under $H_+$ (i.e., a Cauchy distribution with scale parameter *r* = $\sqrt2/2$). This prior was chosen to reflect the expectation of a "medium" effect size [@morey2011; @schnuerch2020; @schonbrodt2017].

I drew random samples from two normal distributions with common variance $\sigma^2$ = 1 and means $\mu_1$ = 0 and $\mu_2 = \delta$ ($\delta$ = -0.2, 0, 0.2, 0.4, 0.5, 0.6, 0.8, 1). For each hypothesis testing procedure and true effect size $\delta$, 10,000 iterations were simulated (i.e., 5 procedures with 8 distinct true effect sizes and 10,000 replications each, leading to 400,000 observations in total). Below, I briefly outline each of the five decision-theoretic hypothesis testing procedures followed.

## *Fixed-sample hypothesis test* 
  *
1. Randomly sample $N_{fixed}$ observations from each of the two specified populations
2. Run a one-tailed, two-sample *t* test on the mean difference between the two samples
  a. If $p \le \alpha$: decide to reject $H_0$
  b. If $p > \alpha$: decide not to reject $H_0$
3. To estimate the standardized effect size, calculate Cohen's *d* (the difference in sample means divided by their pooled standard deviation)
4. Repeat 10,000 times
5. Compute the meta-analytic effect size estimate

## *Pocock-like (O'Brien-Fleming-like) Group-sequential procedures*
1. Randomly sample $N_{Pocock}$ ($N_{O'Brien-Fleming}$) observations from each of the two specified populations 
2. Run a one-tailed, two-sample *t* test on the mean difference between the two samples
3. For segments 1:$k_{max} -1$,

  a. if $p \le \alpha_{1}$: decide to reject $H_0$ and end the experiment ("stop for efficacy")
  b. if $p > \ \alpha_{0}$: decide not to reject $H_0$ and end the experiment  ("stop for futility")
  c. if $\alpha_{1} < p~\le~\alpha_{0}$: collect $N_{Pocock}$ ($N_{O'Brien-Fleming}$) additional observations and repeat step 2 on the cumulative data

4. For segment $k_{max}$ (if reached),
  a. if $p \le \alpha_{1}$: decide to reject $H_0$ and end the experiment
  b. if $p > \ \alpha_{1}$: decide not to reject $H_0$ and end the experiment
5. To estimate the standardized effect size, calculate Cohen's *d*
6. Repeat 10,000 times
7. Compute the meta-analytic effect size estimate

## *Sequential-Bayes Factor*
1. Randomly sample $N_{Bayes}$ observations from each of the two specified populations
2. Run a one-sided, two-tailed default Bayes Factor *t* test on the mean difference between the two samples
  a. if $BF_{+0} \ge 3$: decide to reject $H_0$ (support $H_+$) and end the experiment
  b. if $BF_{+0} \le 1/3$: decide to reject $H_+$ (support $H_0$) and end the experiment
  c. if $1/3 < BF_{10} < 3$: collect $N_{Bayes}$ additional observations and repeat step 2 on the cumulative data
3. Continue until the Bayes Factor crosses one of the thresholds, or until the maximum *N* ($N_{Bayes} * k_{max}$) has been reached
4. To estimate the standardized effect size, calculate the mean of the posterior distribution using a two-sided model [@vandoorn2020]
5. Repeat 10,000 times 
6. Compute the meta-analytic effect size estimate

## *Independent segments procedure* 
1. Randomly sample $N_{ISP}$ observations from each of the two specified populations
2. Run a one-tailed, two-sample *t* test on the mean difference between the two samples
3. For segments 1:$k_{max} -1$,

  a. if $p \le \alpha_{strong}$: decide to reject $H_0$ and end the experiment 
  b. if $p > \alpha_{weak}$: decide not to reject $H_0$ and end the experiment
  c. if $\alpha_{strong} < p \le \alpha_{weak}$: discard the data, collect $N_{ISP}$ new observations, and repeat step 2 on this independent batch of data

4. For segment $k_{max}$ (if reached), repeat steps 1 and 2,
  a. if $p \le \alpha_{weak}$: decide to reject $H_0$ and end the experiment
  b. if $p > \alpha_{weak}$: decide not to reject $H_0$ and end the experiment
5. To estimate the standardized effect size, calculate Cohen's *d*
6. Repeat 10,000 times
7. Compute the meta-analytic effect size estimate

# Results

## Error rates

The error rates involved in the procedures under discussion are shown in Table 2A, Table 2B, and Figure 4. Table 2A describes the rates of True Negative, Inconclusive, and False Positive results when the true population effect size $\delta$ = 0; Table 2B describes the rates of True Positive, Inconclusive, and False Negative results when the true population effect size $\delta$ = 0.5. As can be read from the tables, the frequentist procedures all perform as expected: 4-5% of simulations rejected the null hypothesis when $\delta$ = 0 (i.e., a type I error rate of 4-5%; see the False Positive column Table 2A) and 80-81% of simulations rejected the null hypothesis when $\delta$ = 0.5 (i.e., a type II error rate of 19-20%; see the True Positive column of Table 2B).

In addition, the results speak to the ability of using simulations to design an efficient Bayesian hypothesis testing procedure with attractive long-run properties. In this case, the False Positive rate of the SBF procedure is 4% and the True Positive rate of the SBF procedure is 78%. In addition, when a true population effect size exists, the SBF procedure consistently leads to a lower False Negative rate (and a higher rate of Inconclusive results) than its frequentist counterparts. This is in part due to the high(er) bar that the Bayes Factor requires to support the Null Hypothesis. In this case, the alternative hypothesis ($H_+:~\delta > 0$) has a within-model prior centered on zero (i.e., a Cauchy distribution with scale parameter *r* = $\sqrt2/2$). For this model to be dominated by the null hypothesis ($H_0:~\delta = 0$) when in fact a true population effect exists, it takes a large number of misleading samples in which the effect size is close to zero. Thus, a Bayesian will have a harder time concluding evidence in favor of the null (a False Negative) when a population effect exists.

In line with previous findings [@schonbrodt2017; @schonbrodt2018; @stefan2020], I find that simulation-based approaches allow researchers to choose an optimal Sequential Bayes Factor design that controls expected error rates. It has previously been shown that the maximum-flexibility SBF procedure -- with appropriate choices for the decision threshold and the prior distribution of effect sizes under the alternative hypothesis -- can perform (at least) as well as other (sequential) procedures in terms of false-positive and false-negative error rates [@schonbrodt2017; @stefan2020]. Here, I show that this observation holds when we use a restricted version of the SBF with $k_{max}$ = 3 equally spaced looks.

## Efficiency

Table 3 and Figure 5 show the average sample size required for the procedures under investigation, as a function of the true population effect size. As discussed, the fixed-sample procedure requires 51 participants per group. As can be read from Table 3 and Figure 5, the sequential procedures allow for a reduction in the average sample size required. Across the procedures, efficiency gains appear to be rather similar. When the true population effect size is relatively large, the O'Brien-Fleming approach is slightly less efficient than its counterparts. This can be explained by the relatively stringent $\alpha_1$ levels of the O'Brien-Fleming procedure in earlier segments (recall that, for the O'Brien-Fleming-like procedure, $\alpha_{1,1}$ = `r alpha_of[1]` and $\alpha_{1,2}$ = `r alpha_of[2]`).

Our sequential procedures are performing as expected, providing an appealing increase in the efficiency of hypothesis tests. Next, we consider whether --- and if so, to what extent --- these efficiency gains come at a cost to our ability to estimate the true effect size.

## Accuracy

Figure 6 shows the empirical distribution of effect size estimates stemming from the five hypothesis testing procedures when the true effect size $\delta$ = 0.5. Effect size estimates stemming from the fixed sample hypothesis test are, as expected, normally distributed around the true effect size ($\delta$ = 0.5). As we can see, the distribution of effect size estimates stemming from the sequential procedures have a somewhat odd shape. This is to be expected [see, for example, @goodman2007]. The distribution of observed effects is more spread out for the sequential procedures because the experiments that stopped early have smaller sample sizes and higher variability in effect size estimates. In addition, we observe left-tailed peaks in the distribution of observed effect sizes. This is a consequence of those experiments that stopped early with a decision to support the null (i.e., False Negative results).

For the SBF, the median effect size estimate (*d* = 0.51) is very close to the true value ($\delta$ = 0.5). The GS procedures each have a slight bias in the median effect size estimate: *d* = 0.54 for the O'Brien-Fleming-like procedure and *d* = 0.57 for the Pocock-like procedure. The bias in the median effect size estimate of the ISP, however, is severe (*d* = 0.66; an exaggeration ratio of 1.33). The distribution of effect size estimates stemming from the ISP is bi-modal.

To make sense of this result, we revisit Figure 3B. In the first and second segment of the ISP, the critical effect size (i.e., the effect size that needs to be observed to reject the null hypothesis) is *d* = 0.57. An effect $d \ge$ 0.57 leads to an early stop with a decision to reject the null hypothesis; an effect $d \le$ 0.16 leads to an early stop with a decision not to reject the null hypothesis. An observed effect size 0.16 $\le$ *d* < 0.57 is considered inconclusive, after which the data is discarded and a new batch of observations is collected. As a result, *effect sizes closest to the true population effect size are discarded*.

In the supplemental files, readers can find that this observation is not merely idiosyncratic to the test chosen (a two-sample *t* test with $\alpha$ = .05 and $\beta = .20$ for a hypothesized population effect size $\delta$ = 0.5). In contrast, across dozens of different test specifications, effect size estimates stemming from the ISP are consistently bi-modal and systematically exclude the true population effect size from being observed.

It is enlightening to consider that the Sequential Bayes Factor and Group-Sequential procedures are similarly biased in earlier segments. Early stopping acts as a filter that weeds out moderately sized effects: the only effect sizes that are maintained are the ones that are relatively small or in the wrong direction (e.g., when stopping for futility); or the ones that are really large (e.g., when stopping for efficacy). However, due to the cumulative nature of data collection and analysis, the bias inherent in early terminations tends to be compensated for by studies that terminate at later stages [@goodman2007; @fan2004; @schonbrodt2017]. As more observations are collected, the variability in effect size estimates decreases. Unfortunately, this is not the case for the Independent Segments Procedure.

The median effect size estimates across all possible values of $\delta$ are shown in Table 4. As expected, the median effect size estimates stemming from fixed-sample hypothesis tests are equal to the true population effect size. Median effect size estimates stemming from the Pocock-like design deviate only slightly from the true population effect size. In line with previous findings [@schonbrodt2017], median effect size estimates stemming from the Sequential Bayes Factor slightly underestimate the true population effect size for certain true effect sizes. This is because we use the mean of the posterior distribution as the effect size estimator, which shrinks the estimate in early terminations. Again, we see that the median effect size estimates stemming from the Independent Segments Procedure deviate most from the true population effect size.

Consider the columns of Table 4 that show the median effect size estimates for the respective procedures when the true population effect size $\delta$ = 0.2, 0.4, or 0.6. When the true population effect size $\delta$ = 0.2, the median effect size estimate on the basis of the ISP is *d* = 0.06 (compared to 0.13, 0.13, 0.16, and 0.20 for the SBF, Pocock-like, O'Brien-Fleming-like, and fixed sample procedures, respectively). When the true population effect size $\delta$ = 0.4, the median effect size estimate on the basis of the ISP is *d* = 0.57 (compared to 0.45, 0.42, 0.42, and 0.40 for the Pocock-like, O'Brien-Fleming-like procedure, SBF, and fixed sample procedures, respectively). When the true population effect size $\delta$ = 0.6, the median effect size estimate on the basis of the ISP is *d* = 0.74 (compared to 0.66, 0.62, 0.60, and 0.58 for the Pocock-like, O'Brien-Fleming-like procedure, fixed sample, and SBF procedures, respectively).

Figure 7 visualizes the information captured by Table 4, illustrating the differences between the median estimated effect sizes and true effect sizes. We can see that the Sequential Bayes Factor underestimates the true effect size for certain parameters and the Group-Sequential procedures slightly underestimate the true effect when it is smaller ($\delta \le$ 0.2) and slightly overestimate it when it is larger ($\delta \ge$ 0.4). As before, the ISP shows the greatest deviations from the true effect size.

Figure 8 shows the mean squared error of the four procedures. The mean squared error is a measure of the accuracy of an estimator, composed of two elements: variance (the spread of effect size estimates from one sample to the other) and squared bias (the difference between the mean estimated effect size and the true population effect size). We find that the Independent Segments Procedure has the greatest degree of mean squared error. In addition, we find that the error stems mostly from the higher variance of effect size estimates. As discussed before, the variance in the effect size estimates stemming from the ISP is caused by the process of conducting three independent mini-experiments, rather than a cumulative study.

To provide more meat to this argument, we separate the error of the procedures by segment in Figure 9. Here, we can clearly see that stopping after a first look at data causes error for all sequential procedures. For both the SBF and the Pocock-like procedure, however, error becomes near-negligible in the second and third segment. As a result, overall error (across all segments) is much lower. This is akin to the bias compensation process we have described above. For the Independent Segments Procedure, as we now know, error remains substantial in every segment, leading to non-negligible overall error.

Somewhat paradoxically, here we have an efficient hypothesis testing procedure that is inefficient in statistical parlance. An efficient estimator is one that has small variance [@fisher1922]. The ISP, which purports to increase efficiency, does just the opposite: the process of conducting independent mini-experiments and discarding a region of inconclusive data points, while maintaining only the most extreme data points, causes substantial variability in effect size estimates. More importantly, the region of inconclusive data points that the researcher is forced to discard are just those that are closest to the true effect size. As a result, it is near impossible to obtain a credible effect size estimate from the Independent Segments Procedure.

# Discussion

In this final section, I present a practical guide for researchers who wish to choose between the sequential procedures under discussion. Group-sequential designs, Sequential Bayes Factors, and the Independent Segments Procedure each have advantages and disadvantages. In deciding between sequential hypothesis testing procedures, researchers need to carefully weigh the advantages and disadvantages of the procedures of interest, with regard to the context in which they are conducting their research [@stefan2020]. As discussed earlier, the SBF allows researchers to specify their uncertainty regarding the effect size by means of a prior distribution. Thus, for researchers who wish to incorporate prior beliefs (e.g., because they work in a subject area with high uncertainty about expected effect sizes), the SBF presents an appealing research strategy. On the other hand, if a researcher wants robust (i.e., guaranteed) error control, GS designs present an appealing research strategy.

Although the ISP has some advantages, none of these advantages are unique: efficiency and explicit error control can be achieved through alternative methods, none of which pose the disadvantages unique to the ISP. The ISP applies only to settings in which a researcher can guarantee an exact number of participants -- allocated equally across segments -- and wishes to test a single hypothesis. More importantly, the researcher in question should be willing to discard valuable data and abstain from the ability to obtain a credible estimate of effect size.

@miller2020 present the independent nature of data collection and analysis built-in to the ISP as an attractive feature that creates simplicity and generality. Based on our simulations, however, we find that the information loss inherent to the procedure is a bug, not a feature, that leads to severe difficulty in effect size estimation. Given the existence of appealing alternatives that have the same advantages (of efficiently testing hypotheses with explicit error control) as the ISP, without its associated disadvantages (of severe difficulty in effect size estimation), the reader is advised to think twice before utilizing the Independent Segments Procedure.

Earlier, we introduced the reader to two trade-offs of potential importance to the sequential hypothesis tester: long-run error control vs. short-run evidence and efficiency in hypothesis testing vs. accuracy in parameter estimation, respectively. The short-run conception of evidence on which the (Sequential) Bayes Factor is based, has been accused of throwing out the error control baby with the bathwater [@mayo2016: p.2]. I show here -- in line with previous work -- that with appropriate design choices, the SBF procedure (implemented with a lower degree of flexibility than in previous investigations) can have comparable error rates to its frequentist alternatives. Moreover, despite the trade-off between accurate effect size estimation and efficient hypothesis testing [@goodman2007; @schonbrodt2017], I find that both the SBF procedure and the GS procedure can provide relatively accurate estimates of effect size. Thus, the Sequential Bayes Factor and Group-Sequential procedures can substantially increase research efficiency at only a slight cost to our ability to estimate an effect size.

# Conclusion

We have learned that scientific experiments can require substantial time, money, and effort. At times, the well-being of research subjects, human or non-human, is at stake [@stefan2020]. In these contexts, sequential hypothesis testing can provide substantial efficiency gains. Group-Sequential designs, the Sequential Probability Ratio Test, and the Sequential Bayes Factor procedure are valuable tools for psychological scientists to add to their statistical toolbox.

\newpage

## Acknowledgement
This manuscript was created using `r cite_r("r-refs.bib")`.

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>

\endgroup

\newpage

## Table 1
## *Commonalities and Differences between Three Sequential Hypothesis Testing Procedures*

|  | Group-Sequential Design | Independent Segments Procedure | Sequential Bayes Factor Design |
| ------------- | ------------- | ------------- | ------------- |
| **Aim** | Efficiently reject $H_0$ in favor of an alternative $H_1$ | Efficiently reject $H_0$ in favor of an alternative $H_1$ | Efficiently examine the evidence for $H_0$ relative to an alternative $H_1$, or vice versa |
| **Nature of data**  | Cumulative  | Independent  | Cumulative  | Cumulative |
| **Control of error rates** | Direct | Direct | Indirect (e.g., through simulation) |
| **Flexibility** | Moderately Flexible | Highly Restrictive | Highly Flexible

\newpage

## Table 2A
## *True Negative, False Positive, and Inconclusive Rate*

```{r include = F}
zz = gzfile("simulations/simulation001.csv.gz", 'rt')
df = read.csv(zz, header = T)

df = df %>% 
  mutate(name = case_when(proc == "Fixed" ~ "Fixed",
                           proc == "Bayes" ~ "Bayes",
                           proc == "ISP" ~ "ISP",
                           proc == "asP" ~ "Pocock",
                           proc == "asOF"~ "O'Brien-Fleming"),
         name = fct_relevel(name, "Fixed", "Pocock", "O'Brien-Fleming", "Bayes", "ISP"))

t1 = df %>% filter(d_forpower == d_actual) %>%
  group_by(name) %>% 
  summarize(true_positive = mean(decision == "Reject") %>% round(2),
            inconclusive = mean(decision == "FTR") %>% round(2),
            false_negative = mean(decision == "SupportNull") %>% round(2))

t2 = df %>% filter(d_actual == 0) %>% 
  group_by(name) %>% 
  summarize(true_negative = mean(decision == "SupportNull") %>% round(2),
            inconclusive = mean(decision == "FTR") %>% round(2),
            false_positive = mean(decision == "Reject") %>% round(2))
```

```{r, echo = FALSE}
knitr::kable(t2, 
             col.names = c("Procedure", "True Negative", "Inconclusive", "False Positive"), 
             align = "lccc")
```

*Note.* Error rates of the procedures when the true population effect size $\delta$ = 0.

\newpage

## Table 2B
## *True Positive, False Negative, and Inconclusive Rate*

```{r, echo = FALSE}
knitr::kable(t1, 
             col.names = c("Procedure", "True Positive", "Inconclusive", "False Negative"), 
             align = "lccc")
```

*Note.* Error rates of the procedures when the true population effect size $\delta$ = 0.5.

\newpage

## Table 3
## *Average Sample Size Required*

```{r include = F}
E_n = df %>% 
  group_by(name, d_actual) %>% 
  summarize(E_n = mean(n)) %>% 
  rename(Procedure = name)

t3 = E_n %>% mutate(E_n = E_n %>% round()) %>% 
  pivot_wider(names_from = d_actual, names_prefix = "*d* = ", values_from = E_n) 
```

```{r, echo = FALSE}
knitr::kable(t3)
```

\newpage

## Table 4
## *Median Effect Size Estimate*

```{r, include = FALSE}
med = df %>% 
  group_by(name, d_actual) %>% 
  summarize(medianES = median(ES)) %>% 
  rename(Procedure = name)

t4 = med %>% mutate(medianES = medianES %>% round(2)) %>%  
  pivot_wider(names_from = d_actual, names_prefix = "*d* = ", values_from = medianES)
```

```{r, echo = FALSE}
knitr::kable(t4) #perhaps add median fold inflation to the tables (i.e., exaggeration ratio)
```

*Note.* The table shows the median effect size estimate across simulations. For the Fixed, Pocock, O'Brien-Fleming, and ISP procedure, the effect size measure is Cohen's d. For the Sequential Bayes Factor, the effect size measure is the mean of the posterior distribution.

\newpage

## Figure 1A
## *Pocock-like Group-Sequential Design: A worked example* 
\noindent
```{r, fig.heigth = 1, fig.width = 12}
knitr::include_graphics("figures/figure1a.tiff", dpi = 1200)
```

*Note.* Pocock-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 looks. Critical values are determined on the basis of Pocock-like $\alpha$- and $\beta$-spending functions in *R* package *rpact* (Wassmer & Pahlke, 2020).

\newpage

## Figure 1B
## *O'Brien-Fleming-like Group-Sequential Design: A worked example* 
\noindent
```{r, fig.heigth = 1, fig.width = 12}
knitr::include_graphics("figures/figure1b.tiff", dpi = 1200)
```

*Note.* O'Brien-Fleming-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 looks. Critical values are determined on the basis of O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions in *R* package *rpact* (Wassmer & Pahlke, 2020).

\newpage

## Figure 2A
## *Bayes Favor in Favor of the Alternative Hypothesis* 
\noindent
```{r, fig.heigth = 10, fig.width = 4.5}
knitr::include_graphics("figures/figure2a.tiff", dpi = 1200)
```

*Note.* $BF_{+0}$ = 3 indicates that the alternative hypothesis ($H_+:~\delta > 0$) performs three times better at predicting the observed data than the null hypothesis ($H_0:~\delta = 0$). Figure based on Jeffreys' Amazing Statistics Program (JASP), adapted from Wagenmakers & Gronau [-@wagenmakers]

\newpage

## Figure 2B
## *Bayes Factor in Favor of the Null Hypothesis* 
\noindent
```{r, fig.heigth = 9, fig.width = 4.5}
knitr::include_graphics("figures/figure2b.tiff", dpi = 1200)
```

*Note.* $BF_{0+}$ = 3 indicates that the null hypothesis ($H_0:~\delta = 0$) performs three times better at predicting the observed data than the  alternative hypothesis ($H_+:~\delta > 0$). Figure based on JASP, adapted from Wagenmakers & Gronau [-@wagenmakers]

\newpage

## Figure 3A
## *Fixed Sample versus Independent Segments Hypothesis Testing: A conceptual overview* 
\noindent
```{r, fig.heigth = 1, fig.width = 12}
knitr::include_graphics("figures/figure3a.tiff", dpi = 1200)
```

*Note.* Comparing the general procedures.

\newpage

## Figure 3B
## *Fixed Sample versus Independent Segments Hypothesis Testing: A worked example* 
\noindent
```{r, fig.heigth = 3, fig.width = 12}
knitr::include_graphics("figures/figure3b.tiff", dpi = 1200)
```

*Note.* Specific example of fixed sample versus independent segments one-tailed, two-sample *t* tests using $\alpha$ = .05 and 1 - $\beta$ = 0.8 for $\delta$ = 0.5.

\newpage

## Figure 4
## *Error rates of the four procedures* 
\noindent
```{r, fig.heigth = 4.72, fig.width = 5}
knitr::include_graphics("figures/figure4.tiff", dpi = 1200)
```

*Note.* Red lines with crosses indicate the rates of false negative results; green lines with check marks indicate the rates of true positive results; grey, dashed lines with question marks indicate the rates of inconclusive results.

\newpage

## Figure 5
## *Efficiency of the four procedures* 
\noindent
```{r, fig.heigth = 4.72, fig.width = 5}
knitr::include_graphics("figures/figure5.tiff", dpi = 1200)
```

*Note.* Average sample size required across a range of true effect sizes. The solid, black line represents the fixed-sample Neyman-Pearson hypothesis test; the long-dashed, red line represents the Sequential Bayes Factor procedure; the two-dashed, green line represents the O-Brien-Fleming-like GS procedure; the solid, grey line represents the Pocock-like GS procedure; and the dotted, blue line represents the Independent Segments Procedure.

\newpage

## Figure 6
## *Distribution of Empirical Effect Size Estimates*   
\noindent
```{r, fig.heigth = 4.72, fig.width = 5}
knitr::include_graphics("figures/figure6.tiff", dpi = 300)
```

*Note.* One-tailed, two-sample *t* tests. Fixed Sample, Independent Segments, and Group-Sequential (Pocock- and O'Brien-Fleming-like) Hypothesis Tests are based on $\alpha$ = .05, 1 - $\beta$ = 0.8 for $\delta$ = 0.5. *N*s per group per look are 51, 25, 24, and 19, respectively, with a maximum number of 3 looks. For the Sequential Bayes Factor, data was collected in (at most three) batches of 22 subjects per group, and the procedure was terminated when the one-tailed Bayes Factor was greater than or equal to 3 in favor of the null or the alternative hypothesis. Figure includes the median effect size estimate. True effect size $\delta$ = 0.5

\newpage

## Figure 7

## *Bias in Empirical Effect Size Estimates*   
\noindent
```{r, fig.heigth = 4.72, fig.width = 5}
knitr::include_graphics("figures/figure7.tiff", dpi = 300)
```

*Note.* Figure displays the median estimated effect size minus the true effect size.

\newpage

## Figure 8
## *Bias-Variance in Empirical Effect Size Estimates*  
\noindent
```{r, fig.heigth = 4.72, fig.width = 5}
knitr::include_graphics("figures/figure8.tiff", dpi = 300)
```

*Note.*

\newpage

## Figure 9
## *Mean Squared Error in Empirical Effect Size Estimates by Segment*  
\noindent
```{r, fig.heigth = 4.72, fig.width = 8}
knitr::include_graphics("figures/figure9.tiff", dpi = 300)
```

*Note.*

### 2021-04-27 figures/figures.R

@@ -1,0 +1,991 @@
+rm(list = ls())
+#.rs.restartR() #restart when corrupt
+library(beepr)
+library(ggplot2)
+library(tidyverse)
+library(ggtext)
+library(grid)
+library(papaja)
+library(gridExtra)
+library(metafor) #For an overview and introduction to the package please type: help(metafor).
+library(rpact)
+library(BayesFactor)
+library(metafor)
+
+#Insert code section with Ctrl+Shift+R
+
+# TO DO -------------------------------------------------------------------
+# Increase number of simulations to 50,000 after Yen/Sherlock
+# Add simulation with SPRT - to this end, also change the long-run error control versus short-run evidence section of the paper. After SPRT data is collected, can have 6 panels for all figures. Also add Pramanik's 2020 modified SPRT.
+# Add simulation with equivalence test -- what should I set as the bounds? Notes meeting DL: "flip de toets. De type I fout van de equivalentie toets is gelijk aan de type II fout van de NHST. Sequential analyse jennison en turnboll 2000 group-sequential methods with applications to clinical trials. De simpelste combinatie van een equivalentie toets en de NHST"
+# Perhaps change settings for O'Brien-Fleming: look after 50%, 75%, 100%.
+# Implement bias corrections for GS designs - adjusted ES from rpact
+# For the Bayes procedure, also add the normal effect size estimate (Cohen's d) - this would even speed up your code. 
+# Include meta-analytic effect size 
+# Add a practical considerations sections under the discussion
+  ## A great advantage to group-sequential designs (compared to the SPRT or the SBF) is that, unlike the SPRT and
+  ## the SBF, it is not open-ended: you know exactly how many participants you at most will need to sample.
+  ## How many interim looks can researchers actually take? Look at how SBF/SPRT are applied in practice. In SBF
+  ## designs, how often is data analyzed in practice? Look at papers that use it. 
+  ## "Whereas continuous monitoring is desirable, it is often impractical" (Jennison and Turnbull)
[…]

### 2021-04-27 simulations/simulations.R

@@ -1,0 +1,433 @@
+########################## SET UP ########################## 
+rm(list = ls())
+library(beepr)
+library(rpact)
+library(pracma)
+library(BayesFactor)
+library(coda)
+library(tidyverse)
+library(parallel)
+
+########################## READ ISP FILES ##########################
+source("simulations/engine.R") # Obtained from Ulrich & Miller 2020
+
+########################## FIND ALPHA-WEAK FUNCTION ##########################
+
+find_alpha_weak <- function(alpha_total, max_n_segments, alpha_strong) {
+  # Obtained from Ulrich & Miller 2020
+  # Use numerical search to find the appropriate alpha_weak value
+  # that will produce the desired overall_alpha level for the indicated
+  # values of n_segments and alpha_strong
+  compute_error <- function(try_aw) {
+    diff <- try_aw - alpha_strong
+    diff_accumulated <- diff^(max_n_segments - 1)
+    alpha_strong * (1-diff_accumulated)/(1-diff)+
+      try_aw * diff_accumulated-alpha_total 
+  }
+  aw_range <- c(alpha_strong, alpha_total^(1/max_n_segments))
+  final_aw <- fzero(compute_error, aw_range)
+  final_aw$x
+}
+
+#find_alpha_weak(0.05, 3, .025) #returns .28
+
+########################## SIMULATION FUNCTIONS ##########################
+
+generate_data = function(nsims, n, d) { 
+  # Generate raw data
+  raw = lapply(1:nsims, function(i) {
+    data.frame(nsim = i, n = n, d = d,
+               m1 = rnorm(n = n, mean = 0),
+               m2 = rnorm(n = n, mean = 0) + d)})
+  bind_rows(raw)
+}
+
+run_tests = function(df, n, proc, segment = 1, alpha) {
+  # Run t-tests on raw data
+  p = lapply(1:max(df$nsim), function(i) {
+    df %>% filter(nsim == i) %>% #run 1 analysis per simulation
+      summarize(p = t.test(m2, m1, alternative = "greater", var.equal = T)$p.value)
+    })
+  
+  # Create final data-frame
+  df %>% group_by(nsim) %>% 
+    summarize(proc = proc, segment = segment, 
+              alpha = alpha, n = mean(n),
+              sd1 = sd(m1), sd2 = sd(m2),
+              m1 = mean(m1), m2 = mean(m2)) %>% 
[…]

### 2021-04-27 manuscript.Rmd
adjusted information rates for O'Brien-Fleming

@@
 
-bibliography      : ["better.bib","r-refs.bib"]
+bibliography      : ["references.bib","r-refs.bib"]
 
@@
 library(beepr)
+library(citr)
 
@@
 
+# Practical considerations section
+
+"The SPRT has the theoretical optimal property that, among all tests with error probabilities not exceeding alpha and beta, it attains the smallest possible expected sample size ...
+when either H0 or H1 is true.
+However, the SPRT is an "open" procedure, i.e., the sample size is not bounded, and a consequence of this is that the distribution of sample size can be quite skewed with a large variance." And the ASN can be large when theta is not equal to theta0 or theta1 [@jennison2000: p.23]
+
+A great advantage to group-sequential designs (compared to the SPRT or the SBF) is that, unlike the SPRT and the SBF, it is not open-ended: you know exactly how many participants you at most will need to sample.
+How many interim looks can researchers actually take?
+Look at how SBF/SPRT are applied in practice.
+In SBF designs, how often is data analyzed in practice?
+Look at papers that use it.
+GS-designs are closed - there is a guaranteed upper limit on sample size.
+"Whereas continuous monitoring is desirable, it is often impractical" [@jennison2000: p.21] Ethical considerations: when an experiment is highly costly or may be unsafe, it is an ethical imperative to terminate a trial as soon as possible - in this case, "ethical considerations prescribe that accumulating data be evaluated frequently" [@jennison2000: p.19]
+
+"Armitage (1954, 1958b, 1975) and Bross (1952, 1958) pioneered the use of sequential methods in the medical field, particularly for comparative clinical trials.
+Initially these plans were fully sequential and did not receive widespread acceptance, perhaps because continuous assessment of study results was often impractical."[@jennison2000: p.24]
+
+Re.
+O'Brien-Fleming: discuss the very conservative stopping boundary at early analysis, which also the decision rule to be similar to the fixed sample test if the last stage is reached -- "features which have turned out to be very appealing to practitioners"[@jennison2000: p.25]
+
+Methods were "extended to the often more realistic setting in which group sizes, or more generally increments in information, are unequal and unpredictable" [@jennison2000: p.26] (ISP can not be applied in realistic settings.)
+
 \newpage

### 2021-04-27 figures/figures.R

@@ -15,20 +15,13 @@ library(metafor)
 #Insert code section with Ctrl+Shift+R
 
 # TO DO -------------------------------------------------------------------
-# Increase number of simulations to 50,000 after Yen/Sherlock
-# Add simulation with SPRT - to this end, also change the long-run error control versus short-run evidence section of the paper. After SPRT data is collected, can have 6 panels for all figures. Also add Pramanik's 2020 modified SPRT.
+# Add simulation with SPRT - to this end, also change the long-run error control versus short-run evidence section of the paper. After SPRT data is collected, can have 6 panels for all figures. Also add Pramanik's 2020 modified SPRT. 
 # Add simulation with equivalence test -- what should I set as the bounds? Notes meeting DL: "flip de toets. De type I fout van de equivalentie toets is gelijk aan de type II fout van de NHST. Sequential analyse jennison en turnboll 2000 group-sequential methods with applications to clinical trials. De simpelste combinatie van een equivalentie toets en de NHST"
-# Perhaps change settings for O'Brien-Fleming: look after 50%, 75%, 100%.
+# Change settings for O'Brien-Fleming: look after 50%, 75%, 100%.
 # Implement bias corrections for GS designs - adjusted ES from rpact
 # For the Bayes procedure, also add the normal effect size estimate (Cohen's d) - this would even speed up your code. 
 # Include meta-analytic effect size 
-# Add a practical considerations sections under the discussion
-  ## A great advantage to group-sequential designs (compared to the SPRT or the SBF) is that, unlike the SPRT and
-  ## the SBF, it is not open-ended: you know exactly how many participants you at most will need to sample.
-  ## How many interim looks can researchers actually take? Look at how SBF/SPRT are applied in practice. In SBF
-  ## designs, how often is data analyzed in practice? Look at papers that use it. 
-  ## "Whereas continuous monitoring is desirable, it is often impractical" (Jennison and Turnbull)
-  ## Ethical considerations: when an experiment is highly costly or may be unsafe, it is an ethical imperative to terminate a trial as soon as possible - in this case, "ethical considerations prescribe that accumulating data be evaluated frequently" (Jennison and Turnbull)
+
 # Add in discussion/limitation section: We do not discuss adaptive designs, which extend GS designs and allow for even more flexibility while controlling error rates. Or: Implement adaptive designs???
 # Perhaps change prior for the SBF: r scale value of 0.5 - problem is power is too low. 
[…]

### 2021-04-27 simulations/simulations.R

@@ -63,8 +63,77 @@ run_tests = function(df, n, proc, segment = 1, alpha) {
 
 ########################## SEQUENTIAL PROCEDURE FUNCTION ########################## 
 
-group_sequential = function(proc, target_power, d_forpower, d_actual) {
+#########  FUNCTION WITH ADJUSTED INFORMATION RATES FOR O'BRIEN-FLEMING #########
+group_sequential = function(proc, target_power, d_forpower, d_actual) { #adjusted information rates for O'Brien-Fleming
   bs = ifelse(proc == "asOF", "bsOF", "bsP")
+  if(proc == "asOF") {rates = c(0.50, 0.75, 1)} else{rates = c(1/3, 2/3, 1)}
+  
+  # Specify the design
+  design = getDesignGroupSequential(sided = 1,
+                                    alpha = alpha_total, 
+                                    beta = 1 - target_power,
+                                    kMax = max_n_segments, 
+                                    typeOfDesign = proc,
+                                    typeBetaSpending= bs,
+                                    informationRates = rates) 
+  
+  # Get parameters
+  parameters = getSampleSizeMeans(design = design, 
+                                  groups = 2, 
+                                  alternative = d_forpower)
+  n_gs = ceiling(c(parameters$numberOfSubjects[1], diff(parameters$numberOfSubjects))/2) # n per look per group
+  alpha = parameters$criticalValuesPValueScale
+  futility = parameters$futilityBoundsPValueScale
+  
+  # RUN LOOK 1
+  set.seed(1234*(d_forpower*target_power))
+  raw1 = generate_data(nsims = nsims, n = n_gs[1], d = d_actual)
+  gs1 = run_tests(df = raw1, n = n_gs[1], proc = proc, alpha = alpha[1]) %>% 
+    mutate(decision = ifelse(p <= alpha, "Reject", 
+                             ifelse(p > futility[1], "SupportNull", "Continue")))
+  #Support null not to be taken literally; stopped for futility based on beta-spending
+  
+  keep1 = gs1 %>% filter(decision == "Continue") %>% select(nsim) %>% pull()
+  
+  # RUN LOOK 2
+  gs1keep = raw1 %>% 
+    filter(nsim %in% keep1) %>% 
[…]

### 2021-04-28 figures/figures.R
added bias correction for GS-designs

@@ -17,7 +17,8 @@ library(metafor)
 # TO DO -------------------------------------------------------------------
 # Add simulation with SPRT - to this end, also change the long-run error control versus short-run evidence section of the paper. After SPRT data is collected, can have 6 panels for all figures. Also add Pramanik's 2020 modified SPRT. 
 # Add simulation with equivalence test -- what should I set as the bounds? Notes meeting DL: "flip de toets. De type I fout van de equivalentie toets is gelijk aan de type II fout van de NHST. Sequential analyse jennison en turnboll 2000 group-sequential methods with applications to clinical trials. De simpelste combinatie van een equivalentie toets en de NHST"
-# Change settings for O'Brien-Fleming: look after 50%, 75%, 100%.
+# Rewrite in manuscript: Settings for O'Brien-Fleming: look after 50%, 75%, 100%.
+# Rerun simulations on Sherlock.
 # Implement bias corrections for GS designs - adjusted ES from rpact
 # For the Bayes procedure, also add the normal effect size estimate (Cohen's d) - this would even speed up your code. 
 # Include meta-analytic effect size 

### 2021-04-28 simulations/simulations.R

@@ -37,8 +37,8 @@ generate_data = function(nsims, n, d) {
   # Generate raw data
   raw = lapply(1:nsims, function(i) {
     data.frame(nsim = i, n = n, d = d,
-               m1 = rnorm(n = n, mean = 0),
-               m2 = rnorm(n = n, mean = 0) + d)})
+               m2 = rnorm(n = n, mean = 0),
+               m1 = rnorm(n = n, mean = 0) + d)})
   bind_rows(raw)
 }
 
@@ -46,22 +46,149 @@ run_tests = function(df, n, proc, segment = 1, alpha) {
   # Run t-tests on raw data
   p = lapply(1:max(df$nsim), function(i) {
     df %>% filter(nsim == i) %>% #run 1 analysis per simulation
-      summarize(p = t.test(m2, m1, alternative = "greater", var.equal = T)$p.value)
+      summarize(p = t.test(m1, m2, alternative = "greater", var.equal = T)$p.value)
     })
   
   # Create final data-frame
   df %>% group_by(nsim) %>% 
     summarize(proc = proc, segment = segment, 
-              alpha = alpha, n = mean(n),
+              alpha = alpha, n = mean(n), 
               sd1 = sd(m1), sd2 = sd(m2),
               m1 = mean(m1), m2 = mean(m2)) %>% 
     mutate(sdpooled = sqrt((sd1^2 +sd2^2)/2)) %>% 
-    mutate(ES = (m2 - m1) / sdpooled) %>% #ES estimate = Cohen's d
+    mutate(ES = (m1 - m2) / sdpooled) %>% #ES estimate = Cohen's d
     cbind(tibble(p) %>% 
             unnest(cols = c(p))) #add p-value
 } 
 
 ########################## SEQUENTIAL PROCEDURE FUNCTION ########################## 
+
+#########  FUNCTION WITH BIAS ADJUSTMENT AND ADJUSTED INFORMATION RATES FOR O'BRIEN-FLEMING #########  
+
+group_sequential = function(proc, target_power, d_forpower, d_actual) { 
+
+bs = ifelse(proc == "asOF", "bsOF", "bsP") #Specify beta-spending function
+if(proc == "asOF") {rates = c(0.50, 0.75, 1)} else{rates = c(1/3, 2/3, 1)} #specify information rates for Pocock vs OBF
+
+# Specify the design
+design = getDesignGroupSequential(sided = 1,
+                                  alpha = alpha_total, 
[…]

### 2021-04-28 manuscript.Rmd
+ Bayes naive effect size estimate for consistency - + n per look and cumulative n to all procedures

@@
 library(metafor)
-library(beepr)
 library(citr)

### 2021-04-28 figures/figures.R

@@ -19,7 +19,7 @@ library(metafor)
 # Add simulation with equivalence test -- what should I set as the bounds? Notes meeting DL: "flip de toets. De type I fout van de equivalentie toets is gelijk aan de type II fout van de NHST. Sequential analyse jennison en turnboll 2000 group-sequential methods with applications to clinical trials. De simpelste combinatie van een equivalentie toets en de NHST"
 # Rewrite in manuscript: Settings for O'Brien-Fleming: look after 50%, 75%, 100%.
 # Rerun simulations on Sherlock.
-# Implement bias corrections for GS designs - adjusted ES from rpact
+# Rewrite in manuscript: Implemented bias corrections for GS designs - median unbiased ES from rpact
 # For the Bayes procedure, also add the normal effect size estimate (Cohen's d) - this would even speed up your code. 
 # Include meta-analytic effect size 
 

### 2021-04-28 simulations/simulations.R

@@ -1,6 +1,5 @@
 ########################## SET UP ########################## 
 rm(list = ls())
-library(beepr)
 library(rpact)
 library(pracma)
 library(BayesFactor)
@@ -113,7 +112,7 @@ gs1keep = raw1 %>% #keep raw data of unfinished trials after look 1
 set.seed(1512*(d_forpower*target_power))
 raw2 = generate_data(nsims = sum(gs1$decision=="Continue"), n = n_gs[2], d = d_actual) %>% 
   mutate(id = rep(keep1, each = n_gs[2]), 
-         n_cumulative = n_gs[1] + n_gs[2],  #Do I need to adjust n here?
+         n_cumulative = n_gs[1] + n_gs[2],  
          look = 2) 
 
 gs2 = bind_rows(gs1keep, raw2) %>% 
@@ -134,7 +133,7 @@ gs2keep = raw2 %>% filter(nsim %in% keep2) %>%
 
 set.seed(0304*(d_forpower*target_power))
 raw3 = generate_data(nsims = sum(gs2$decision=="Continue"), n = n_gs[3], d = d_actual) %>% 
-  mutate(id = rep(keep1[keep2], each = n_gs[3]), #mistake here 
+  mutate(id = rep(keep1[keep2], each = n_gs[3]), 
          n_cumulative = sum(n_gs),
          look = 3)
 
@@ -145,8 +144,9 @@ gs3 = bind_rows(gs1keep, gs2keep, raw3) %>%
          n_cumulative = sum(n_gs), 
          decision = ifelse(p <= alpha, "Reject", "FTR"))
 
-# Bias-correction with rpact
+##############################
 
+# Bias-correction with rpact
 raw_merged = bind_rows(raw1, raw2, raw3) %>% 
   group_by(id, look) %>% 
   summarize(n = mean(n),
@@ -158,7 +158,6 @@ raw_merged = bind_rows(raw1, raw2, raw3) %>%
          ES = (m1 - m2)/ sdpooled) #ES = uncorrected Cohen's d
 
 # Get median unbiased estimate
-
 unbiased = raw_merged %>% 
   split(.$id) %>% 
   map(~ getDataset(
@@ -174,21 +173,211 @@ unbiased = raw_merged %>%
     dataInput = .x
   )) %>%
   map("medianUnbiased") %>% 
-  data_frame(ES_corrected = .) 
[…]

### 2021-04-28 simulations/simulations.R
reran simulations with bias corrections

@@ -347,11 +347,13 @@ alpha_weak= find_alpha_weak(alpha_total, max_n_segments, alpha_strong) #calculat
 
 system.time(dt <- mclapply(1:length(d_actual), 
                            function(i) procedure(d_forpower[i], d_actual[i]),
-                           mc.cores = 2)) #takes ~56mins to run when nsim = 10,000 & ncores = 2
-                                          #takes ~7.9 hours to run when nsim = 50,000 & ncores = 2
-beep()
+                           mc.cores = 2)) #takes ~13 hours to run with nsim = 50,000 and ncores = 2
+system("say Your code has finished running!")
 df = dt %>% bind_rows()
 write.csv(df, file=gzfile("simulations/simulation001.csv.gz"), row.names = F) #write in zip to compress
+
+########################## READ IN DATA ##########################
+
 zz = gzfile("simulations/simulation001.csv.gz", 'rt')
 df = read.csv(zz, header = T) #read in data
 

### 2021-04-28 manuscript.Rmd
added modified SPRT

@@
 library(citr)
+library(MSPRT)
 
@@
 
+"From a practical perspective, there are many hypothesis testing contexts where it is not feasible to implement a SPRT.
+For instance, a SPRT cannot be applied when data are not collected sequentially.
+Similarly, it cannot be applied when it is not possible to perform the evaluation as soon as participants are treated...
+To address these challenges, group sequential designs have been developed to allow for the evaluation of patient outcomes only after groups of patients have been observed or at scheduled interim analysis times" [@pramanik2021: p.
+4-5]
+
+# To add
+
+In this paper, the example we use looks at the performance of these procedures under ideal conditions (an oracle prior setup), i.e.
+the performance of these procedures when the true effect size matches the model expectations as closely as possible: the population effect size is either equal to zero or it matches the effect size parameter (SPRT, Fixed, GS-designs) or the mode of the prior distribution on effect size (SBF) in the alternative model [@stefan2020: p.
+15].
+
 \newpage

### 2021-04-28 figures/figures.R

@@ -20,11 +20,12 @@ library(metafor)
 # Rewrite in manuscript: Settings for O'Brien-Fleming: look after 50%, 75%, 100%.
 # Rerun simulations on Sherlock.
 # Rewrite in manuscript: Implemented bias corrections for GS designs - median unbiased ES from rpact
-# For the Bayes procedure, also add the normal effect size estimate (Cohen's d) - this would even speed up your code. 
+# For the Bayes procedure, also add the normal effect size estimate (Cohen's d). 
 # Include meta-analytic effect size 
+# Perhaps change prior for the SBF: r scale value of 0.5 - problem is power is too low. Perhaps use prior from Stefan et al. 2020 - informed normal distribution centered on the population effect size. 
+# Optimize Bayes thresholds using the R Package NMOF (differential evolution algorithm - see Stefan)
 
 # Add in discussion/limitation section: We do not discuss adaptive designs, which extend GS designs and allow for even more flexibility while controlling error rates. Or: Implement adaptive designs???
-# Perhaps change prior for the SBF: r scale value of 0.5 - problem is power is too low. 
 # Emphasize that there is no bias correction available for the ISP. Also emphasize that the ISP is the only procedure that is *not* median unbiased.
 # Change tables: show results for all effect sizes
 # Rephrase wording: instead of fail to reject, simply refer to Rejecting the Null and Rejecting the Alternative.

### 2021-04-28 simulations/simulations.R

@@ -6,6 +6,7 @@ library(BayesFactor)
 library(coda)
 library(tidyverse)
 library(parallel)
+library(MSPRT)
 
 ########################## READ ISP FILES ##########################
 source("simulations/engine.R") # Obtained from Ulrich & Miller 2020
@@ -46,7 +47,7 @@ run_tests = function(df, n, proc, segment = 1, alpha) {
   p = lapply(1:max(df$nsim), function(i) {
     df %>% filter(nsim == i) %>% #run 1 analysis per simulation
       summarize(p = t.test(m1, m2, alternative = "greater", var.equal = T)$p.value)
-    })
+  }) %>% map("p") %>% unlist()
   
   # Create final data-frame
   df %>% group_by(nsim) %>% 
@@ -55,9 +56,8 @@ run_tests = function(df, n, proc, segment = 1, alpha) {
               sd1 = sd(m1), sd2 = sd(m2),
               m1 = mean(m1), m2 = mean(m2)) %>% 
     mutate(sdpooled = sqrt((sd1^2 +sd2^2)/2)) %>% 
-    mutate(ES = (m1 - m2) / sdpooled) %>% #ES estimate = Cohen's d
-    cbind(tibble(p) %>% 
-            unnest(cols = c(p))) #add p-value
+    mutate(ES = (m1 - m2) / sdpooled, #ES estimate = Cohen's d #Change to Hedges g?
+           p = p) #add p-value
 } 
 
 ########################## SEQUENTIAL PROCEDURE FUNCTION ########################## 
@@ -158,7 +158,7 @@ raw_merged = bind_rows(raw1, raw2, raw3) %>%
          ES = (m1 - m2)/ sdpooled) #ES = uncorrected Cohen's d
 
 # Get median unbiased estimate
-unbiased = raw_merged %>% 
+ES_corrected = raw_merged %>% 
   split(.$id) %>% 
   map(~ getDataset(
     n1 = .$n,
@@ -168,12 +168,12 @@ unbiased = raw_merged %>%
     stDevs1 = .$sd1,
     stDevs2 = .$sd2
   )) %>% 
-    map(~ getFinalConfidenceInterval(
+  map(~ getFinalConfidenceInterval(
     design = design,
     dataInput = .x
   )) %>%
   map("medianUnbiased") %>% 
-  tibble() %>% unnest(cols = c(.)) %>% rename(ES_corrected = ".")
+  unlist()
 
[…]

### 2021-04-29 manuscript.Rmd
added equivalence test to fixed sample NHST

@@
 library(MSPRT)
+library(TOSTER)
 
@@
 
+"Continuous data monitoring can be a serious practical burden, and the introduction of group sequential tests has led to much wider use of sequential methods.
+(...) Not only are group sequential tests convenient to conduct, they also provide ample opportunity for early stopping and can achieve most of the benefits of fully sequential tests in terms of lower expected sample sizes and shorter average study lengths." [@jennison2000: p.
+38]
+
 # To add
 
-In this paper, the example we use looks at the performance of these procedures under ideal conditions (an oracle prior setup), i.e.
+In this paper, the example we use throughout looks at the performance of these procedures under ideal conditions (an oracle prior setup), i.e.
 the performance of these procedures when the true effect size matches the model expectations as closely as possible: the population effect size is either equal to zero or it matches the effect size parameter (SPRT, Fixed, GS-designs) or the mode of the prior distribution on effect size (SBF) in the alternative model [@stefan2020: p.
 15].
+
+Add writing on the modified SPRT.
 

### 2021-04-29 figures/figures.R

@@ -15,7 +15,7 @@ library(metafor)
 #Insert code section with Ctrl+Shift+R
 
 # TO DO -------------------------------------------------------------------
-# Add simulation with SPRT - to this end, also change the long-run error control versus short-run evidence section of the paper. After SPRT data is collected, can have 6 panels for all figures. Also add Pramanik's 2020 modified SPRT. 
+# Make decision wording consistent across procedures. 
 # Add simulation with equivalence test -- what should I set as the bounds? Notes meeting DL: "flip de toets. De type I fout van de equivalentie toets is gelijk aan de type II fout van de NHST. Sequential analyse jennison en turnboll 2000 group-sequential methods with applications to clinical trials. De simpelste combinatie van een equivalentie toets en de NHST"
 # Rewrite in manuscript: Settings for O'Brien-Fleming: look after 50%, 75%, 100%.
 # Rerun simulations on Sherlock.

### 2021-04-29 simulations/simulations.R

@@ -7,6 +7,7 @@ library(coda)
 library(tidyverse)
 library(parallel)
 library(MSPRT)
+library(TOSTER)
 
 ########################## READ ISP FILES ##########################
 source("simulations/engine.R") # Obtained from Ulrich & Miller 2020
@@ -97,9 +98,8 @@ raw1 = generate_data(nsims = nsims, n = n_gs[1], d = d_actual) %>%
 gs1 = run_tests(df = raw1, n = n_gs[1], proc = proc, alpha = alpha[1]) %>% 
   mutate(id = nsim, 
          n_cumulative = n_gs[1],
-         decision = ifelse(p <= alpha, "Reject",
-                           ifelse(p > futility[1], "SupportNull", "Continue")))
-#Support null not to be taken literally; stopped for futility based on beta-spending
+         decision = ifelse(p <= alpha, "reject.null",
+                           ifelse(p > futility[1], "reject.alt", "Continue"))) 
 
 keep1 = gs1 %>% filter(decision == "Continue") %>% select(nsim) %>% pull()
 
@@ -119,9 +119,9 @@ gs2 = bind_rows(gs1keep, raw2) %>%
   run_tests(n = n_gs[2], proc = proc, alpha = alpha[2], segment = 2) %>% 
   mutate(id = keep1,
          n = n_gs[2],
-         n_cumulative = n_gs[1] + n_gs[2],
-         decision = ifelse(p <= alpha, "Reject", ifelse(p > futility[2], "SupportNull", "Continue")))
-#Support null not to be taken literally; stopped for futility based on beta-spending
+         n_cumulative = n_gs[1] + n_gs[2], 
+         decision = ifelse(p <= alpha, "reject.null", 
+                           ifelse(p > futility[2], "reject.alt", "Continue")))
 
 keep2 = gs2 %>% filter(decision == "Continue") %>% select(nsim) %>% pull()
 
@@ -142,7 +142,7 @@ gs3 = bind_rows(gs1keep, gs2keep, raw3) %>%
   mutate(id = keep1[keep2],
          n = n_gs[3],
          n_cumulative = sum(n_gs), 
-         decision = ifelse(p <= alpha, "Reject", "FTR"))
+         decision = ifelse(p <= alpha, "reject.null", "inconclusive")) 
 
 ##############################
 
@@ -203,14 +203,44 @@ procedure = function(d_forpower, d_actual) {
[…]

### 2021-04-29 simulations/simulations.R
reran simulations with modified sprt and equivalence test

@@ -441,10 +441,28 @@ alpha_weak= find_alpha_weak(alpha_total, max_n_segments, alpha_strong) #calculat
 
 system.time(dt <- mclapply(1:length(d_actual), 
                            function(i) procedure(d_forpower[i], d_actual[i]),
-                           mc.cores = 2)) #takes ~13 hours to run with nsim = 50,000 and ncores = 2
+                           mc.cores = 2)) #takes ~14 hours to run with nsim = 50,000 and ncores = 2
 system("say Your code has finished running!")
 df = dt %>% bind_rows()
 write.csv(df, file=gzfile("simulations/simulation001.csv.gz"), row.names = F) #write in zip to compress
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+########################## READ IN DATA ##########################
+
+zz = gzfile("simulations/simulation001.csv.gz", 'rt')
+df = read.csv(zz, header = T) #read in data
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+
+############################################################################
+############################################################################
+############################################################################
+
+############################################################################
+############################################################################
+############################################################################
+# ARCHIVED FUNCTIONS ------------------------------------------------------
 
 # MSPRT --------------------------------------------------------------------
 
@@ -454,7 +472,7 @@ MSPRT = function(d_forpower, d_actual) {
   maxN = 23*3
   
   cumulativeNs = seq(minN, maxN, minN)
-
+  
   design_SPRT = design.MSPRT(test.type = "twoT",
                              theta1 = d_forpower,
                              Type1.target = alpha_total,
@@ -537,23 +555,6 @@ df = bind_rows(df, sprt)
 # Write data
 write.csv(df, file=gzfile("simulations/simulation001.csv.gz"), row.names = F) #write in zip to compress
 
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-########################## READ IN DATA ##########################
-
-zz = gzfile("simulations/simulation001.csv.gz", 'rt')
-df = read.csv(zz, header = T) #read in data
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-
-############################################################################
-############################################################################
-############################################################################
-
-############################################################################
-############################################################################
-############################################################################
[…]

### 2021-05-03 manuscript.Rmd
implemented data.table, BFDA, modified Bayes + SPRT with sample sizes matched to OBF

@@
 library(TOSTER)
+library(BFDA)
 
@@
 Add writing on the modified SPRT. 
+Add writing on the combination of fixed sample NHST and equivalence test.
+Rewrite settings for O'Brien-Fleming: look after 50%, 75%, 100%.
+Add writing on the implemented bias corrections for GS designs - median unbiased ES from rpact
+Add writing on naive effect size estimate Bayes
+Rephrase decision wording: refer to rejecting the null, rejecting the alternative, and inconclusive.
+Change writing modified SBF prior from Stefan et al. 2020 - informed normal distribution centered on the population effect size (mean 0.5, variance 0.3).
+
+Here's some examples on how SBFs are used in practice https://github.com/nicebread/BFDA/blob/master/BFDA_examples.md
 
@@
 ```{r include = F}
-zz = gzfile("simulations/simulation001.csv.gz", 'rt')
-df = read.csv(zz, header = T)
+df = read.csv("simulations/simulation001.csv")
+#zipped.df = gzfile("simulations/simulation001.csv.gz", 'rt')
+#df = read.csv(zipped.df, header = T)
 
@@
   group_by(name) %>% 
-  summarize(true_positive = mean(decision == "Reject") %>% round(2),
-            inconclusive = mean(decision == "FTR") %>% round(2),
-            false_negative = mean(decision == "SupportNull") %>% round(2))
+  summarize(true_positive = mean(decision == "reject.null") %>% round(2),
+            inconclusive = mean(decision == "inconclusive") %>% round(2),
+            false_negative = mean(decision == "reject.alt") %>% round(2))
 
@@
   group_by(name) %>% 
-  summarize(true_negative = mean(decision == "SupportNull") %>% round(2),
-            inconclusive = mean(decision == "FTR") %>% round(2),
-            false_positive = mean(decision == "Reject") %>% round(2))
+  summarize(true_negative = mean(decision == "reject.alt") %>% round(2),
+            inconclusive = mean(decision == "inconclusive") %>% round(2),
+            false_positive = mean(decision == "reject.null") %>% round(2))
 ```

### 2021-05-03 figures/figures.R

@@ -15,27 +15,21 @@ library(metafor)
 #Insert code section with Ctrl+Shift+R
 
 # TO DO -------------------------------------------------------------------
-# Make decision wording consistent across procedures. 
-# Add simulation with equivalence test -- what should I set as the bounds? Notes meeting DL: "flip de toets. De type I fout van de equivalentie toets is gelijk aan de type II fout van de NHST. Sequential analyse jennison en turnboll 2000 group-sequential methods with applications to clinical trials. De simpelste combinatie van een equivalentie toets en de NHST"
-# Rewrite in manuscript: Settings for O'Brien-Fleming: look after 50%, 75%, 100%.
+# Bayes: add effect size correction
+# Run meta-analysis and include meta-analytic effect size figure 
+# Take figure 9 out of paper
+# Maybe change figures (higher, less wide, labels with \n)
 # Rerun simulations on Sherlock.
-# Rewrite in manuscript: Implemented bias corrections for GS designs - median unbiased ES from rpact
-# For the Bayes procedure, also add the normal effect size estimate (Cohen's d). 
-# Include meta-analytic effect size 
-# Perhaps change prior for the SBF: r scale value of 0.5 - problem is power is too low. Perhaps use prior from Stefan et al. 2020 - informed normal distribution centered on the population effect size. 
-# Optimize Bayes thresholds using the R Package NMOF (differential evolution algorithm - see Stefan)
-
+# Run with 90% power (ideally, both?)
 # Add in discussion/limitation section: We do not discuss adaptive designs, which extend GS designs and allow for even more flexibility while controlling error rates. Or: Implement adaptive designs???
 # Emphasize that there is no bias correction available for the ISP. Also emphasize that the ISP is the only procedure that is *not* median unbiased.
 # Change tables: show results for all effect sizes
-# Rephrase wording: instead of fail to reject, simply refer to Rejecting the Null and Rejecting the Alternative.
-# Make example in the paper 90% power?
 # Go through comments Jake - Chunchen - Charles
-# Speed up code: Datatable? 
+# Speed up cod w Datatable 
 # Rewrite to call MSE reducible error 
 # Make density by segment plot
[…]

### 2021-05-03 simulations/engine.R

@@ -1,2531 +1,0 @@
-####ISP Engine provided by Ulrich & Miller at https://github.com/milleratotago/Independent_Segments_R
-
-#setwd("/Users/shilaan/Library/Mobile Documents/com~apple~CloudDocs/PhD/ISP")
-
-# Class SegmentedHypTestEngine
-
-# Runs (computes the proababilistic results of) complete segmented hypothesis testing scenarios
-
-#########################################################################################################
-# Creation methods - following Wickham
-#########################################################################################################
-#========================================================================================================
-# new_SegmentedHypTestEngine
-#========================================================================================================
-new_SegmentedHypTestEngine <- function()
-{
-  
-  # No fields needed at creation
-  new_SegmentedHypTestEngine <- list()
-  
-  
-  attr(new_SegmentedHypTestEngine, "class") <- "SegmentedHypTestEngine"
-  
-  # Return instance
-  return(new_SegmentedHypTestEngine)
-  
-} # new_SegmentedHypTestEngine
-#========================================================================================================
-
-#========================================================================================================
-# validate_SegmentedHypTestEngine
-#========================================================================================================
-# Checks additional properties of elements, for example, that values are in allowed ranges -- business rules
-validate_SegmentedHypTestEngine <- function(segmented_hyp_test_engine)
-{
-  # More elaborate checking in here if needed. Stop if things go wrong.
-  
-} # end validate_SegmentedHypTestEngine
-#========================================================================================================
-
-#========================================================================================================
-# SegmentedHypTestEngine - Public ctor
-#========================================================================================================
-#-----------------
-# Roxygen comments
-
-#' SegmentedHypTestEngine
-#'
-#' \code{SegmentedHypTestEngine}  is the public-facing SegmentedHypTestEngine
-#' constructor. It accepts no parameters. This object exposes the primary methods
-#' for numerically determing the expected outcomes of a segmented hypothesis testing
-#' study. See run_scenario.SegmentedHypTestEngine for an example.
-#' @return Validated SegmentedHypTestEngine instance
-#'
-#' @export
-SegmentedHypTestEngine <- function()
-{
-  # Create instance
-  instance = new_SegmentedHypTestEngine()
-  
-  # Validate instance. Execution halts here if things go wrong
-  validate_SegmentedHypTestEngine(instance)
-  
[…]

### 2021-05-03 simulations/isp-engine.R

@@ -1,0 +1,2531 @@
+####ISP Engine provided by Ulrich & Miller at https://github.com/milleratotago/Independent_Segments_R
+
+#setwd("/Users/shilaan/Library/Mobile Documents/com~apple~CloudDocs/PhD/ISP")
+
+# Class SegmentedHypTestEngine
+
+# Runs (computes the proababilistic results of) complete segmented hypothesis testing scenarios
+
+#########################################################################################################
+# Creation methods - following Wickham
+#########################################################################################################
+#========================================================================================================
+# new_SegmentedHypTestEngine
+#========================================================================================================
+new_SegmentedHypTestEngine <- function()
+{
+  
+  # No fields needed at creation
+  new_SegmentedHypTestEngine <- list()
+  
+  
+  attr(new_SegmentedHypTestEngine, "class") <- "SegmentedHypTestEngine"
+  
+  # Return instance
+  return(new_SegmentedHypTestEngine)
+  
+} # new_SegmentedHypTestEngine
+#========================================================================================================
+
+#========================================================================================================
+# validate_SegmentedHypTestEngine
+#========================================================================================================
+# Checks additional properties of elements, for example, that values are in allowed ranges -- business rules
+validate_SegmentedHypTestEngine <- function(segmented_hyp_test_engine)
+{
+  # More elaborate checking in here if needed. Stop if things go wrong.
+  
+} # end validate_SegmentedHypTestEngine
+#========================================================================================================
+
+#========================================================================================================
+# SegmentedHypTestEngine - Public ctor
+#========================================================================================================
+#-----------------
+# Roxygen comments
+
+#' SegmentedHypTestEngine
+#'
+#' \code{SegmentedHypTestEngine}  is the public-facing SegmentedHypTestEngine
+#' constructor. It accepts no parameters. This object exposes the primary methods
+#' for numerically determing the expected outcomes of a segmented hypothesis testing
+#' study. See run_scenario.SegmentedHypTestEngine for an example.
+#' @return Validated SegmentedHypTestEngine instance
+#'
+#' @export
+SegmentedHypTestEngine <- function()
+{
+  # Create instance
+  instance = new_SegmentedHypTestEngine()
+  
+  # Validate instance. Execution halts here if things go wrong
+  validate_SegmentedHypTestEngine(instance)
+  
[…]

### 2021-05-03 simulations/simulations.R

@@ -1,5 +1,6 @@
 ########################## SET UP ########################## 
 rm(list = ls())
+library(data.table)
 library(rpact)
 library(pracma)
 library(BayesFactor)
@@ -8,338 +9,536 @@ library(tidyverse)
 library(parallel)
 library(MSPRT)
 library(TOSTER)
+library(BFDA)
+library(microbenchmark)
+library(NCmisc)
+library(here)
 
-########################## READ ISP FILES ##########################
-source("simulations/engine.R") # Obtained from Ulrich & Miller 2020
-
-########################## FIND ALPHA-WEAK FUNCTION ##########################
-
-find_alpha_weak <- function(alpha_total, max_n_segments, alpha_strong) {
-  # Obtained from Ulrich & Miller 2020
-  # Use numerical search to find the appropriate alpha_weak value
-  # that will produce the desired overall_alpha level for the indicated
-  # values of n_segments and alpha_strong
-  compute_error <- function(try_aw) {
-    diff <- try_aw - alpha_strong
-    diff_accumulated <- diff^(max_n_segments - 1)
-    alpha_strong * (1-diff_accumulated)/(1-diff)+
-      try_aw * diff_accumulated-alpha_total 
-  }
-  aw_range <- c(alpha_strong, alpha_total^(1/max_n_segments))
-  final_aw <- fzero(compute_error, aw_range)
-  final_aw$x
-}
-
-#find_alpha_weak(0.05, 3, .025) #returns .28
-
-########################## SIMULATION FUNCTIONS ##########################
+########################## GLOBAL HELPER FUNCTIONS ##########################
 
 generate_data = function(nsims, n, d) { 
   # Generate raw data
   raw = lapply(1:nsims, function(i) {
-    data.frame(nsim = i, n = n, d = d,
+    data.table(nsim = i, n = n, d = d,
                m2 = rnorm(n = n, mean = 0),
                m1 = rnorm(n = n, mean = 0) + d)})
-  bind_rows(raw)
+  data.table::rbindlist(raw)
 }
 
 run_tests = function(df, n, proc, segment = 1, alpha) {
-  # Run t-tests on raw data
-  p = lapply(1:max(df$nsim), function(i) {
-    df %>% filter(nsim == i) %>% #run 1 analysis per simulation
-      summarize(p = t.test(m1, m2, alternative = "greater", var.equal = T)$p.value)
-  }) %>% map("p") %>% unlist()
[…]

### 2021-05-06 manuscript.Rmd
Added bias-corrected ES estimate for SBF

@@
 ---
-title             : "Sequential Hypothesis Testing: A Comparison of Group-Sequential Designs, the Sequential Bayes Factor, and the Independent Segments Procedure"
+title             : "Group-Wise Sequential Hypothesis Testing"
 shorttitle        : "Sequential Hypothesis Testing"
@@
   Sequential analysis allows researchers to perform well-powered experiments with smaller sample sizes.
-In this practical guide, I outline the commonalities and differences between three sequential hypothesis testing procedures: group-sequential designs, the sequential Bayes Factor, and the independent segments procedure.
+In this practical guide, I outline the commonalities and differences between four sequential hypothesis testing procedures: group-sequential designs, the sequential Bayes Factor, the independent segments procedure, and the modified sequential probability ratio test.
+I illustrate how these procedures can be implemented in a group-wise manner.
 Using Monte-Carlo simulations, I compare the performance of these procedures in terms of their error control (i.e., false-negative and false-positive decision rates), efficiency (i.e., reduction in required sample size), and accuracy (i.e., bias and variance in effect size estimation).

### 2021-05-06 figures/figures.R

@@ -15,7 +15,6 @@ library(metafor)
 #Insert code section with Ctrl+Shift+R
 
 # TO DO -------------------------------------------------------------------
-# Bayes: add effect size correction
 # Run meta-analysis and include meta-analytic effect size figure 
 # Take figure 9 out of paper
 # Maybe change figures (higher, less wide, labels with \n)

### 2021-05-06 simulations/bayes-helpers.R

@@ -1,0 +1,101 @@
+# ==============================================================================
+# Acknowledgement: 
+# All functions obtained from the BFDA package (https://github.com/nicebread/BFDA)
+# With the exception of the EV function (to calculate the mean of the posterior distribution),
+# Which was generously given to me by Angelika Stefan. Thank you, Angelika!
+
+# ==============================================================================
+# Functions obtained from the BFDA package
+# ==============================================================================
+# These are the functions for t-tests with informed priors
+# ==============================================================================
+# see also https://arxiv.org/abs/1704.02479 for the formulae
+
+#' @import hypergeo
+
+# helper functions for the computation of the Bayes factor with informed priors
+
+A <- function(t, n, nu, mu.delta, g) {
+  
+  Re(hypergeo::genhypergeo(U = (nu + 1)/2, L = 1/2,
+                           z = mu.delta^2*t^2/
+                             (2*(1/n + g)*((1 + n*g)*nu + t^2))))
+  
+}
+
+B <- function(t, n, nu, mu.delta, g) {
+  
+  out <- mu.delta*t/sqrt(1/2*(1/n + g)*((1 + n*g)*nu + t^2)) *
+    exp(lgamma((nu + 2)/2) - lgamma((nu + 1)/2)) *
+    Re(hypergeo::genhypergeo(U = (nu + 2)/2, L = 3/2,
+                             z = mu.delta^2*t^2/
+                               (2*(1/n + g)*((1 + n*g)*nu + t^2))))
+  
+  return(out)
+  
+}
+
+
+C <- function(delta, t, n, nu) {
+  
+  Re(hypergeo::genhypergeo(U = (nu + 1)/2, L = 1/2,
+                           z = n*t^2*delta^2/(2*(nu + t^2))))
+  
+}
+
+D <- function(delta, t, n, nu) {
+  
+  out <- t*delta*sqrt(2*n/(nu + t^2))*
+    exp(lgamma((nu + 2)/2) - lgamma((nu + 1)/2))*
+    Re(hypergeo::genhypergeo(U = (nu + 2)/2, L = 3/2,
+                             z = n*t^2*delta^2/(2*(nu + t^2))))
+  
+  return(out)
+  
+}
+
+term_normalprior <- function(t, n, nu, mu.delta, g) {
+  
+  (1 + n*g)^(-1/2) * exp(-mu.delta^2/(2*(1/n + g))) *
[…]

### 2021-05-06 simulations/simulations.R

@@ -15,7 +15,6 @@ library(NCmisc)
 library(here)
 
 ########################## GLOBAL HELPER FUNCTIONS ##########################
-
 generate_data = function(nsims, n, d) { 
   # Generate raw data
   raw = lapply(1:nsims, function(i) {
@@ -43,34 +42,11 @@ run_tests = function(df, n, proc, segment = 1, alpha) {
   
 } 
 
-# To do:------
-# Reconcile ES estimate from BFDA and other functions 
-# Note that BFDA calculates Cohen's d from the t-statistic, which leads to a slight overestimation of ES 
-
-#run_tests = function(df, n, proc, segment = 1, alpha) {
-#  
-#  # Run t-tests and create data-frame
-#  df %>% group_by(nsim) %>% 
-#    summarize(p = t.test(m1, m2, alternative = "greater", var.equal = T)$p.value,
-#              t = t.test(m1, m2, alternative = "greater", var.equal = T)$statistic,
-#              proc = proc, 
-#              segment = segment, 
-#              alpha = alpha, 
-#              n = mean(n), 
-#              sd1 = sd(m1), 
-#              sd2 = sd(m2),
-#              m1 = mean(m1), 
-#              m2 = mean(m2)) %>% 
-#    mutate(sdpooled = sqrt((sd1^2 +sd2^2)/2)) %>% 
-#    mutate(ES = (m1 - m2) / sdpooled,
-#           emp.ES = 2*t / sqrt(2*n-2)) #from BFDA
-#} 
 ########################## GLOBAL PARAMETERS ##########################
 max_n_segments = 3
 alpha_total = 0.05
 target_power = 0.8
 d_forpower = 0.5
-
 
 ########################## FULL PROCEDURE FUNCTION ##########################
 
@@ -322,7 +298,7 @@ run_all_procedures = function(nsims){
   ########################## INDEPENDENT SEGMENTS PROCEDURE FUNCTION ########################## 
 
   ########################## READ ISP FUNCTIONS ##########################
-  source("simulations/isp-engine.R") # Obtained from Ulrich & Miller 2020
+  source("simulations/isp-helpers.R") # Obtained from Ulrich & Miller 2020
 
   # Find alpha weak
   find_alpha_weak <- function(alpha_total, max_n_segments, alpha_strong) {
@@ -453,8 +429,15 @@ run_all_procedures = function(nsims){
              ),
              segment = n_cumulative/n,
              d_forpower = d_forpower,
-             power = target_power,
[…]

### 2021-11-03 manuscript.Rmd
picked up this project after 6 centuries: made to do list, took out MSPRT, reran with 10k sims, reformatted some code

@@
 library(papaja)
+library(beepr)
 library(tidyverse)
@@
 library(citr)
-library(MSPRT)
 library(TOSTER)
@@
 
-While research efficiency is commendable [although not indisputably so; @tiokhin2020], it is not the only advantage conferred by sequential hypothesis testing.
+While research efficiency is commendable, it is not the only advantage conferred by sequential hypothesis testing.
 Under the Ethics Code of the American Psychological Association, psychological scientists have the ethical obligation to minimize research subjects' exposure to physical, emotional, or psychological harm [@apa2020].
@@
 
+  # Add binding futility bounds and the different rates
+
   # Get parameters
@@
 The goal of GS procedures is to efficiently reject a null hypothesis in favor of an alternative hypothesis.
-In a prototypical group-sequential procedure [e.g., the Pocock correction or the O'Brien-Fleming correction, @pocock1977; @fleming1984], a researcher is allowed to look at the data several times during the course of a study.
+In a prototypical group-sequential procedure [e.g., the Pocock correction or the O'Brien-Fleming correction\; @fleming1984; @pocock1977], a researcher is allowed to look at the data several times during the course of a study.
 
@@
 Second, it is highly flexible: it allows researchers to take as many looks at incoming data as desired [even after every study participant; @schonbrodt2017].
-Third, it does not allow type I and type II error rates to be defined a priori [although simulations can be run to choose an optimal design that controls expected error rates, @schonbrodt2017; @schonbrodt2018; @stefan2020].
+Third, it does not allow type I and type II error rates to be defined a priori [although simulations can be run to choose an optimal design that controls expected error rates, @schonbrodt2018; @schonbrodt2017; @stefan2020].
 Fourth, it allows (and requires) researchers to express their uncertainty about the expected effect size by means of a prior distribution.
@@
 In a previous contribution, the Sequential Bayes Factor was compared to GS designs in terms of error rates, efficiency, and accuracy [@schonbrodt2017].
-However, these comparisons were unbalanced in two regards: (a) the GS designs used did not allow for early stopping (i.e., they did not incorporate, for instance, a $\beta$-spending function to stop for futility) and (b) the number of looks was not equal across procedures (i.e., the Sequential Bayes Factor procedure was based on a near-unlimited number of looks until the decision threshold was met, while the chosen GS design incorporated four looks).
+However, these comparisons were unbalanced in two regards: (a) the GS designs used did not allow for early stopping (i.e., they did not incorporate, for instance, a $\beta$-spending function to stop for futility) and (b) the number of looks was not equal across procedures (i.e., the Sequential Bayes Factor procedure was based on a near-unlimited number of looks until the decision threshold was met, while the chosen GS design incorporated only four looks).
 
@@
 In this paper, I implement the SBF with $k_{max}$ equally spaced looks, for two reasons: (a) to allow for maximum comparability between the sequential procedures under discussion, controlling for as many confounds as possible; and (b) to account for practical considerations: it may very well be impractical for researchers to implement the maximum-flexibility SBF approach and analyze data after every participant.
-Thus, I focus on a slightly different, pragmatic question: For a fixed number of equally spaced looks at data, how do these respective procedures perform in terms of their error rates, efficiency, and accuracy?
+Thus, I focus on a slightly different, pragmatic question: For a fixed number of equally spaced looks at data (i.e., what I term a "group-wise" implementation), how do these respective procedures perform in terms of their error rates, efficiency, and accuracy?
 
@@
 1. For the fixed-sample hypothesis test, the procedure ended in a failure to reject the null hypothesis 
-2. For the Pocock and O'Brien-Fleming procedures, $k_{max}$ was reached and neither of the thresholds was met ($\alpha_1 < p~\le~\alpha_0$)
+2. For the Pocock and O'Brien-Fleming procedures, $k_{max}$ was reached and ended in a failure to reject the null hypothesis ($p > \ \alpha_1$)
 3. For the SBF procedure, $k_{max}$ was reached and neither of the thresholds was met ($1/t*~< BF_{+0} < t*$)
@@
 The error rates involved in the procedures under discussion are shown in Table 2A, Table 2B, and Figure 4.
-Table 2A describes the rates of True Negative, Inconclusive, and False Positive results when the true population effect size $\delta$ = 0; Table 2B describes the rates of True Positive, Inconclusive, and False Negative results when the true population effect size $\delta$ = 0.5.
+Table 2A describes the rates of True Negative, Inconclusive, and False Positive results when the true population effect size $\delta$ = 0; Table 2B describes the rates of True Positive, Inconclusive, and False Negative results when the true population effect size $\delta$ = 0.50.
 As can be read from the tables, the frequentist procedures all perform as expected: 4-5% of simulations rejected the null hypothesis when $\delta$ = 0 (i.e., a type I error rate of 4-5%; see the False Positive column Table 2A) and 80-81% of simulations rejected the null hypothesis when $\delta$ = 0.5 (i.e., a type II error rate of 19-20%; see the True Positive column of Table 2B).
@@
 To provide more meat to this argument, we separate the error of the procedures by segment in Figure 9.
-Here, we can clearly see that stopping after a first look at data causes error for all sequential procedures.
+Here, we can clearly see that stopping after a first look causes error for all sequential procedures.
 For both the SBF and the Pocock-like procedure, however, error becomes near-negligible in the second and third segment.
@@
 In these contexts, sequential hypothesis testing can provide substantial efficiency gains.
-Group-Sequential designs, the Sequential Probability Ratio Test, and the Sequential Bayes Factor procedure are valuable tools for psychological scientists to add to their statistical toolbox.
+Group-Sequential designs and the Sequential Bayes Factor procedure are valuable tools for psychological scientists to add to their statistical toolbox.
 
-# Practical considerations section
+<!-- # Practical considerations section  -->
 
-"The SPRT has the theoretical optimal property that, among all tests with error probabilities not exceeding alpha and beta, it attains the smallest possible expected sample size ...
-when either H0 or H1 is true.
-However, the SPRT is an "open" procedure, i.e., the sample size is not bounded, and a consequence of this is that the distribution of sample size can be quite skewed with a large variance." And the ASN can be large when theta is not equal to theta0 or theta1 [@jennison2000: p.23]
-
-A great advantage to group-sequential designs (compared to the SPRT or the SBF) is that, unlike the SPRT and the SBF, it is not open-ended: you know exactly how many participants you at most will need to sample.
+<!-- A great advantage to group-sequential designs (compared to the SPRT or the SBF) is that, unlike the SPRT and the SBF, it is not open-ended: you know exactly how many participants you at most will need to sample.
 How many interim looks can researchers actually take?
@@
 GS-designs are closed - there is a guaranteed upper limit on sample size.
-"Whereas continuous monitoring is desirable, it is often impractical" [@jennison2000: p.21] Ethical considerations: when an experiment is highly costly or may be unsafe, it is an ethical imperative to terminate a trial as soon as possible - in this case, "ethical considerations prescribe that accumulating data be evaluated frequently" [@jennison2000: p.19]
+"Whereas continuous monitoring is desirable, it is often impractical" [@jennison2000: p.21] Ethical considerations: when an experiment is highly costly or may be unsafe, it is an ethical imperative to terminate a trial as soon as possible - in this case, "ethical considerations prescribe that accumulating data be evaluated frequently" [@jennison2000: p.19] -->
 
-"Armitage (1954, 1958b, 1975) and Bross (1952, 1958) pioneered the use of sequential methods in the medical field, particularly for comparative clinical trials.
-Initially these plans were fully sequential and did not receive widespread acceptance, perhaps because continuous assessment of study results was often impractical."[@jennison2000: p.24]
+<!-- "Armitage (1954, 1958b, 1975) and Bross (1952, 1958) pioneered the use of sequential methods in the medical field, particularly for comparative clinical trials.
+Initially these plans were fully sequential and did not receive widespread acceptance, perhaps because continuous assessment of study results was often impractical."[@jennison2000: p.24] -->
 
-Re.
-O'Brien-Fleming: discuss the very conservative stopping boundary at early analysis, which also the decision rule to be similar to the fixed sample test if the last stage is reached -- "features which have turned out to be very appealing to practitioners"[@jennison2000: p.25]
+<!-- Re.
+O'Brien-Fleming: discuss the very conservative stopping boundary at early analysis, which also makes the decision rule similar to the fixed sample test if the last stage is reached -- "features which have turned out to be very appealing to practitioners"[@jennison2000: p.25] -->
 
-Methods were "extended to the often more realistic setting in which group sizes, or more generally increments in information, are unequal and unpredictable" [@jennison2000: p.26] (ISP can not be applied in realistic settings.)
+<!-- Methods were "extended to the often more realistic setting in which group sizes, or more generally increments in information, are unequal and unpredictable" [@jennison2000: p.26] (ISP can not be applied in realistic settings.) -->
 
-"From a practical perspective, there are many hypothesis testing contexts where it is not feasible to implement a SPRT.
+<!-- "From a practical perspective, there are many hypothesis testing contexts where it is not feasible to implement a SPRT.
 For instance, a SPRT cannot be applied when data are not collected sequentially.
@@
 To address these challenges, group sequential designs have been developed to allow for the evaluation of patient outcomes only after groups of patients have been observed or at scheduled interim analysis times" [@pramanik2021: p.
-4-5]
+4-5] -->
 
-"Continuous data monitoring can be a serious practical burden, and the introduction of group sequential tests has led to much wider use of sequential methods.
+<!-- "Continuous data monitoring can be a serious practical burden, and the introduction of group sequential tests has led to much wider use of sequential methods.
 (...) Not only are group sequential tests convenient to conduct, they also provide ample opportunity for early stopping and can achieve most of the benefits of fully sequential tests in terms of lower expected sample sizes and shorter average study lengths." [@jennison2000: p.
-38]
+38] -->
 
-# To add
+<!-- # To add -->
 
-In this paper, the example we use throughout looks at the performance of these procedures under ideal conditions (an oracle prior setup), i.e.
+<!-- Add in discussion/limitation section: We do not discuss adaptive designs, which extend GS designs and allow for even more flexibility while controlling error rates.  -->
+<!-- Address supplemental materials / OSF links, etc.  -->
+<!-- Fix Table 3  -->
+<!-- Change writing on Bayes True Negative. Because I used an oracle prior setup, true negative rates actually look great -->
+<!-- Also refer to MSE as reducible error  -->
+<!-- Maybe add the figure of bimodality for all ISP parameters? Look through previous commits for this. -->
+<!-- Emphasize that there is no bias correction available for the ISP. Also emphasize that the ISP is the only procedure that is *not* median unbiased. Check if Lakens' comment makes this point? -->
+<!-- Incorporate points from three new Psych Methods papers (Lakens, Sneurch/Erdfelder, Ulrich/Miller) -->
+<!-- Add writing on the combination of fixed sample NHST and equivalence test. -->
+<!-- Rephrase decision wording for the Fixed Sample Test combined with equivalence test: refer to rejecting the null, rejecting the alternative, and inconclusive. -->
+<!-- Rewrite settings for O'Brien-Fleming: look after 50%, 75%, 100%. -->
+<!-- Add writing on the implemented bias corrections for GS designs - median unbiased ES from rpact -->
+<!-- Add writing on naive effect size estimate Bayes -->
+<!-- Change writing modified SBF prior from Stefan et al. 2020 - informed normal distribution centered on the population effect size (mean 0.5, variance 0.3).  -->
+<!-- Explain oracle prior setup: the population effect size is equal to the mode of the prior distribution on effect size in the alternative model [@stefan2020:p.15]  -->
+<!-- In this paper, the example we use throughout looks at the performance of these procedures under ideal conditions (an oracle prior setup), i.e.
 the performance of these procedures when the true effect size matches the model expectations as closely as possible: the population effect size is either equal to zero or it matches the effect size parameter (SPRT, Fixed, GS-designs) or the mode of the prior distribution on effect size (SBF) in the alternative model [@stefan2020: p.
 15].
+-->
 
-Add writing on the modified SPRT. 
-Add writing on the combination of fixed sample NHST and equivalence test.
-Rewrite settings for O'Brien-Fleming: look after 50%, 75%, 100%.
-Add writing on the implemented bias corrections for GS designs - median unbiased ES from rpact
-Add writing on naive effect size estimate Bayes
-Rephrase decision wording: refer to rejecting the null, rejecting the alternative, and inconclusive.
-Change writing modified SBF prior from Stefan et al. 2020 - informed normal distribution centered on the population effect size (mean 0.5, variance 0.3).
-
-Here's some examples on how SBFs are used in practice https://github.com/nicebread/BFDA/blob/master/BFDA_examples.md
+<!-- Here's some examples on how SBFs are used in practice https://github.com/nicebread/BFDA/blob/master/BFDA_examples.md  -->
 
@@
   group_by(name, d_actual) %>% 
-  summarize(E_n = mean(n)) %>% 
+  summarize(E_n = mean(n_cumulative)) %>% 
   rename(Procedure = name)
@@
 ```{r, fig.heigth = 4.72, fig.width = 5}
-knitr::include_graphics("figures/figure7.tiff", dpi = 300)
+knitr::include_graphics("figures/figure7a.tiff", dpi = 300)
 ```
 
-*Note.* Figure displays the median estimated effect size minus the true effect size.
+*Note.* Figure displays the mean estimated effect size minus the true effect size.
 
@@
 \noindent
-```{r, fig.heigth = 4.72, fig.width = 8}
+```{r, fig.heigth = 4, fig.width = 8}
 knitr::include_graphics("figures/figure9.tiff", dpi = 300)

### 2021-11-03 figures/figures.R

@@ -1,4 +1,3 @@
-rm(list = ls())
 #.rs.restartR() #restart when corrupt
 library(beepr)
 library(ggplot2)
@@ -15,20 +14,24 @@ library(metafor)
 #Insert code section with Ctrl+Shift+R
 
 # TO DO -------------------------------------------------------------------
+# Figure 6: Add legend for corrected ES estimate 
+# Add line for correct ES estimate BFDA?
+# Check why citations don't render correctly (perhaps add re-add those papers to Zotero?)
+# Change figures after taking out MSPRT (height/width/labels with \n)
+    # Set Fixed Sample Figure to left, and the Sequential Figures to right in a 2x2
+    # OR: Add extra panel for Fixed Sample Test only?
+# Change group-sequential designs figures (binding futility bounds + different rates for OBF changed the N). # Change this^ in manuscript too
+# Change table 2a and 2b code in manuscript
+
+# MAYBE -------------------------------------------------------------------
 # Run meta-analysis and include meta-analytic effect size figure 
 # Take figure 9 out of paper
-# Maybe change figures (higher, less wide, labels with \n)
-# Rerun simulations on Sherlock.
+# Rerun simulations on Sherlock?
 # Run with 90% power (ideally, both?)
-# Add in discussion/limitation section: We do not discuss adaptive designs, which extend GS designs and allow for even more flexibility while controlling error rates. Or: Implement adaptive designs???
-# Emphasize that there is no bias correction available for the ISP. Also emphasize that the ISP is the only procedure that is *not* median unbiased.
-# Change tables: show results for all effect sizes
-# Go through comments Jake - Chunchen - Charles
-# Speed up cod w Datatable 
-# Rewrite to call MSE reducible error 
 # Make density by segment plot
-# Also vary number of looks (and power)
-# Change group-sequential designs figures (binding futility bounds changed the N)
+# Also vary number of looks 
+# Add a table that show results for all effect sizes (maybe for supplemental?)
[…]

### 2021-11-03 simulations/simulations.R

@@ -1,5 +1,4 @@
 ########################## SET UP ########################## 
-rm(list = ls())
 library(data.table)
 library(rpact)
 library(pracma)
@@ -7,7 +6,7 @@ library(BayesFactor)
 library(coda)
 library(tidyverse)
 library(parallel)
-library(MSPRT)
+#library(MSPRT)
 library(TOSTER)
 library(BFDA)
 library(microbenchmark)
@@ -38,14 +37,14 @@ run_tests = function(df, n, proc, segment = 1, alpha) {
               m1 = mean(m1), 
               m2 = mean(m2)) %>% 
     mutate(sdpooled = sqrt((sd1^2 +sd2^2)/2)) %>% 
-    mutate(ES = (m1 - m2) / sdpooled) #ES estimate = Cohen's d #Change to Hedges g?
+    mutate(ES = (m1 - m2) / sdpooled) #ES estimate = Cohen's d
   
 } 
 
 ########################## GLOBAL PARAMETERS ##########################
 max_n_segments = 3
 alpha_total = 0.05
-target_power = 0.8
+target_power = 0.8 #might also add 0.9
 d_forpower = 0.5
 
 ########################## FULL PROCEDURE FUNCTION ##########################
@@ -59,8 +58,13 @@ run_all_procedures = function(nsims){
     ########################## FIXED SAMPLE ##########################
     
     # Calculate fixed n per group for the one-tailed two-sample t-test
-    n_fixed = ceiling(power.t.test(delta = d_forpower, sig.level = alpha_total, power = target_power, 
-                                   type = "two.sample", alternative = "one.sided")$n)
+    n_fixed = ceiling(
+      power.t.test(
+        delta = d_forpower, 
+        sig.level = alpha_total, 
+        power = target_power, 
+        type = "two.sample", 
+        alternative = "one.sided")$n)
     
     # RUN FIXED SAMPLE NHST
     set.seed(1234*(d_forpower*target_power))
@@ -105,9 +109,10 @@ run_all_procedures = function(nsims){
     # Create final data-frame
     fixed = fixed %>% 
       mutate(equivalence = eq_tests$outcome) %>% 
-      mutate(decision = ifelse(p <= alpha, "reject.null", 
-                               ifelse(equivalence == TRUE, "reject.alt",
-                                      "inconclusive"))) 
+      mutate(decision = ifelse(
+        p <= alpha, "reject.null", 
+        ifelse(equivalence == TRUE, "reject.alt",
+               "inconclusive"))) 
     
[…]

### 2021-11-04 manuscript.Rmd
fixed coding error in group-sequential designs; first draft supplemental materials

@@
   Sequential analysis allows researchers to perform well-powered experiments with smaller sample sizes.
-In this practical guide, I outline the commonalities and differences between four sequential hypothesis testing procedures: group-sequential designs, the sequential Bayes Factor, the independent segments procedure, and the modified sequential probability ratio test.
+In this practical guide, I outline the commonalities and differences between three sequential hypothesis testing procedures: group-sequential designs, the sequential Bayes Factor, and the Independent Segments Procedure.
 I illustrate how these procedures can be implemented in a group-wise manner.
@@
 
-keywords          : "sequential analysis, hypothesis testing, efficiency, effect size estimation, bias"
+keywords          : "sequential analysis, hypothesis testing, efficiency, estimation, bias"
 wordcount         : "-"
@@
 By means of simulation, I will demonstrate the properties of the procedures under discussion in the context of testing hypotheses about positive mean differences between two independent groups (i.e., a one-tailed, two-sample *t*-test).
-A one-tailed test was chosen to allow for comparability with the ISP, which has been developed for one-tailed hypothesis testing only.
 The procedures will be compared in terms of three main properties: (a) error rates, (b) efficiency, and (c) accuracy in effect size estimation.
@@
 <!-- Add in discussion/limitation section: We do not discuss adaptive designs, which extend GS designs and allow for even more flexibility while controlling error rates.  -->
-<!-- Address supplemental materials / OSF links, etc.  -->
-<!-- Fix Table 3  -->
-<!-- Change writing on Bayes True Negative. Because I used an oracle prior setup, true negative rates actually look great -->
+<!-- Discuss the power of the individual segments.
+Practical advice (can lead to guidance on choosing alpha strong; make sure that the alpha strong you set does not exclude the effect size you are interested in.
+Write reveal.effectsize function): You have to make sure that the individual segments are each powered to at least 50 percent to detect the hypothesized effect size: otherwise, the critical effect size will be higher than the hypothesized effect size, and you will thus systematically discard the effect sizes that you intended to detect!
+Show that with the example used, the first two independent segments are powered to ~30%.
+In all but the last segment, the critical effect size (the effect size you need to observe to claim significance) is a factor larger than the actual population effect that you powered your test for.
+It is thus impossible to get an unbiased estimate out of anything but the last segment.
+When you obtain an effect size close to the true effect, the ISP procedure forces you to continue to the last segment, and *throw out that data*.
+(You can even graph this exaggeration ratio: create a table that shows the exact critical effect size for all possible combinations, and then you can just read off that plot that you're setting yourself up for overestimation).
+-->
+<!-- Discuss conditional bias (bias in effect size estimates conditional on stopping early) and unconditional bias (bias in effect size estimates across all studies).
+All other procedures have low unconditional bias, when you take the late termination studies into account, because they do cumulative analyses — the late termination studies are able to compensate for the bias in the early stage terminations.
+However, because the independent segments procedure is independent, each individual segment is biased and this is never accounted or compensated for.
+Thus, the ISP is unique among sequential procedures in that it not only has strong conditional bias (like most sequential procedures): it also has strong unconditional bias.
+[In this paper, I focused on two properties: conditional bias (bias in effect size estimates conditional on stopping early) and unconditional bias (bias in effect size estimates across all studies).
+I show that the independent segments procedure is unique among sequential procedures in that it not only has strong conditional bias (like all sequential procedures); it also has strong unconditional bias.] Emphasize that this is also what we see in Figure 3B: we only reject when d > 0.57.
+(This is not the case for, e.g., Pocock)   -->
+<!-- In the language of decision theory, the ISP is inadmissible: its total squared error risk exceeds that of other hypothesis testing procedures no matter what the true effect size may be.  -->
+<!-- Change writing on Bayes True Negative. Because I switched to an oracle prior setup, true negative rates actually look great -->
 <!-- Also refer to MSE as reducible error  -->
 <!-- Maybe add the figure of bimodality for all ISP parameters? Look through previous commits for this. -->
-<!-- Emphasize that there is no bias correction available for the ISP. Also emphasize that the ISP is the only procedure that is *not* median unbiased. Check if Lakens' comment makes this point? -->
+<!-- Emphasize that there is no bias correction available for the ISP, while there is for the SBF and GS.
+Also emphasize that the ISP is the only procedure that is *not* median unbiased.
+Check if Lakens' comment makes this point?
+-->
 <!-- Incorporate points from three new Psych Methods papers (Lakens, Sneurch/Erdfelder, Ulrich/Miller) -->
@@
 
-<!-- Here's some examples on how SBFs are used in practice https://github.com/nicebread/BFDA/blob/master/BFDA_examples.md  -->
+<!-- (To make the point that the setup of the simulations is more practical and realistic than previous ones), Show some data on what scientists actually do, e.g., how often they look at their data when they use the Sequential Bayes Factor.
+Here's some examples on how SBFs are used in practice https://github.com/nicebread/BFDA/blob/master/BFDA_examples.md  -->
 
@@
 | **Control of error rates** | Direct | Direct | Indirect (e.g., through simulation) |
-| **Flexibility** | Moderately Flexible | Highly Restrictive | Highly Flexible
+| **Bias correction procedure**| Available | Does not exist | Available |
+| **Flexibility** | Moderately flexible | Highly restrictive | Highly flexible |
+| **Acceptance of the null** | Possible | Impossible | Possible |
 
@@
 ```{r include = F}
-df = read.csv("simulations/simulation001.csv")
-#zipped.df = gzfile("simulations/simulation001.csv.gz", 'rt')
-#df = read.csv(zipped.df, header = T)
+df = read.csv("simulations/data.csv")
 
 df = df %>% 
-  mutate(name = case_when(proc == "Fixed" ~ "Fixed",
+  mutate(
+    name = case_when(
+      proc == "Fixed" ~ "Fixed",
       proc == "Bayes" ~ "Bayes",
@@
       proc == "asP" ~ "Pocock",
-                           proc == "asOF"~ "O'Brien-Fleming"),
-         name = fct_relevel(name, "Fixed", "Pocock", "O'Brien-Fleming", "Bayes", "ISP"))
+      proc == "asOF"~ "O'Brien-Fleming"
+      ),
+    name = fct_relevel(name, "Fixed", "Pocock", "O'Brien-Fleming", "Bayes", "ISP")
+    )
 
@@
 \noindent
-```{r, fig.heigth = 4, fig.width = 8}
+```{r, fig.heigth = 4, fig.width = 6}
 knitr::include_graphics("figures/figure9.tiff", dpi = 300)

### 2021-11-04 figures/figures.R

@@ -47,7 +47,7 @@ alpha_gs = function(proc, return = "n") {
     beta = 0.2,
     kMax = 3, 
     typeOfDesign = proc,
-    typeBetaSpending= bs) 
+    typeBetaSpending= bs) #also specify information rates
   
   # Get parameters
   parameters = getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
@@ -723,10 +723,7 @@ dev.off()
 ########################## FIGURE 4: ERROR RATES ########################## 
 #=========================================================================#
 # WITHOUT THE MSPRT
-rm(list = ls())
-df = read.csv("simulations/simulation001.csv")
-#zipped.df = gzfile("simulations/simulation001.csv.gz", 'rt')
-#df = read.csv(zipped.df, header = T)
+df = read.csv("simulations/data.csv")
 
 df = df %>% 
   mutate(
@@ -790,7 +787,7 @@ ggplot(data = ER_summary %>% filter(d_actual > 0), mapping = aes(x = d_actual, y
        y = "True Positives, False Negatives, Inconclusive Evidence")
 dev.off()
 
-# Error Rates Table 2a [see code in manuscript.Rmd] #change code in manuscript
+# Error Rates Table 2a [see code in manuscript.Rmd] 
 t1 = df %>% filter(d_forpower == d_actual) %>%
   group_by(proc) %>% 
   summarize(true_positive = mean(decision == "reject.null"),
@@ -809,10 +806,6 @@ t2 = df %>% filter(d_actual == 0) %>%
 ########################## FIGURE 4: ERROR RATES ########################## 
 #=========================================================================#
 # WITH MSPRT
-# rm(list = ls())
-# df = read.csv("simulations/simulation001.csv")
-# #zipped.df = gzfile("simulations/simulation001.csv.gz", 'rt')
-# #df = read.csv(zipped.df, header = T)
 # 
 # df = df %>% 
 #   mutate(
@@ -900,8 +893,8 @@ ggplot(data = E_n, mapping = aes(x = d_actual, y = E_n, group = proc, linetype =
   annotate(geom = "text", x = 0.3, y = 47.7, label = "Bayes", color = "red") +
   #annotate(geom = "segment", x = 0.54, y = 48, xend = 0.66, yend = 48, linetype = "longdash", color = "red", size = 1) +
[…]

### 2021-11-04 simulations/simulations.R

@@ -6,7 +6,6 @@ library(BayesFactor)
 library(coda)
 library(tidyverse)
 library(parallel)
-#library(MSPRT)
 library(TOSTER)
 library(BFDA)
 library(microbenchmark)
@@ -26,16 +25,19 @@ generate_data = function(nsims, n, d) {
 run_tests = function(df, n, proc, segment = 1, alpha) {
   
   # Run t-tests and create data-frame
-  df %>% group_by(nsim) %>% 
-    summarize(p = t.test(m1, m2, alternative = "greater", var.equal = T)$p.value,
-              proc = proc, 
-              segment = segment, 
-              alpha = alpha, 
-              n = mean(n), 
-              sd1 = sd(m1), 
-              sd2 = sd(m2),
-              m1 = mean(m1), 
-              m2 = mean(m2)) %>% 
+  df %>% 
+    group_by(nsim) %>% 
+    summarize(
+      p = t.test(m1, m2, alternative = "greater", var.equal = T)$p.value,
+      proc = proc, 
+      segment = segment, 
+      alpha = alpha, 
+      n = mean(n), 
+      sd1 = sd(m1), 
+      sd2 = sd(m2),
+      m1 = mean(m1), 
+      m2 = mean(m2)
+      ) %>% 
     mutate(sdpooled = sqrt((sd1^2 +sd2^2)/2)) %>% 
     mutate(ES = (m1 - m2) / sdpooled) #ES estimate = Cohen's d
   
@@ -68,18 +70,24 @@ run_all_procedures = function(nsims){
     
     # RUN FIXED SAMPLE NHST
     set.seed(1234*(d_forpower*target_power))
-    fixed = generate_data(nsims = nsims,
-                          n = n_fixed, 
-                          d = d_actual) %>% 
-      run_tests(n = n_fixed, 
-                proc = "Fixed", 
-                alpha = alpha_total) %>% 
-      mutate(id = nsim, 
-             n_cumulative = n, 
-             ES_corrected = ES,
-             d_forpower = d_forpower, 
-             d_actual = d_actual, 
-             power = target_power)
+    fixed = generate_data(
+      nsims = nsims,
+      n = n_fixed, 
+      d = d_actual
+      ) %>% 
+      run_tests(
+        n = n_fixed, 
+        proc = "Fixed",
+        alpha = alpha_total
+        ) %>% 
+      mutate(
+        id = nsim,
+        n_cumulative = n, 
[…]

### 2021-11-04 supplemental/figures/figures_supplemental.R

@@ -1,0 +1,79 @@
+# Supplemental Materials
+# Add figures comparing ISP to Fixed (code for these sims in 'FixedvsISPsims.Rmd')
+# Add Google Drive > Independent Segments Procedure > Simulations > Figure 1 (Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing). Make ISP solid, fixed dash. Code in 'Figures.R'
+# Add Google Drive > Independent Segments Procedure > Simulations > Figure 4 (Bias in Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing). Code in 'Figures.Rmd'
+# Add Google Drive > Independent Segments Procedure > Simulations > Figure 7 (Bias-Variance in Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing).  Code in 'Figures.Rmd'
+# Add Google Drive > Independent Segments Procedure > Simulations > Figure 6 (Estimated Cohen's d using Fixed versus Independent Segments Hypothesis Testing).  Code in 'Figures Appendix.Rmd'
+
+
+# Set up ------------------------------------------------------------------
+
+library(ggplot2)
+library(tidyverse)
+library(ggtext)
+library(grid)
+
+df = read.csv("supplemental/simulations/data_supplemental.csv")
+
+# Figure S1 ----------------------------------------------------------------
+# Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing
+
+plt = function(f){
+  D = mean(f$d) #get separate plot for each combo of power and d
+  target_power = mean(f$power)
+  t = sprintf("\u03b4 = %0.1f, 1 - *\u03b2* = %0.1f", D, target_power)
+  
+  ggplot(data = f, aes(x = ES, group = proc, linetype = proc)) +
+    geom_density() +
+    theme_classic() +
+    theme(
+      legend.position = "none", 
+      plot.title = ggtext::element_markdown(size = 11, hjust = 0.5, face = "bold"),
+      axis.title = element_blank(), 
+      text = element_text(size = 11), 
+      axis.text.y = element_blank(),
+      axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1), 
+      axis.ticks.y = element_blank()
+      ) +
+    scale_x_continuous(breaks = c(0, 0.5*D, D, 1.5*D),
+                       limits = c(-D, 2.5*D)) +
+    scale_linetype_manual(values=c("dashed", "solid"))+
+    labs(title = t)
+}
+
[…]

### 2021-11-04 supplemental/simulations/simulations_supplemental.R

@@ -1,0 +1,289 @@
+########################## SET UP ########################## 
+library(data.table)
+library(rpact)
+library(pracma)
+#library(coda)
+library(tidyverse)
+library(parallel)
+#library(TOSTER)
+#library(NCmisc)
+library(here)
+
+########################## GLOBAL HELPER FUNCTIONS ##########################
+generate_data = function(nsims, n, d) { 
+  # Generate raw data
+  raw = lapply(1:nsims, function(i) {
+    data.table(nsim = i, n = n, d = d,
+               m2 = rnorm(n = n, mean = 0),
+               m1 = rnorm(n = n, mean = 0) + d)})
+  data.table::rbindlist(raw)
+}
+
+run_tests = function(df, n, proc, segment = 1, alpha) {
+  
+  # Run t-tests and create data-frame
+  df %>% 
+    group_by(nsim) %>% 
+    summarize(
+      p = t.test(m1, m2, alternative = "greater", var.equal = T)$p.value,
+      proc = proc, 
+      segment = segment, 
+      alpha = alpha, 
+      n = mean(n), 
+      sd1 = sd(m1), 
+      sd2 = sd(m2),
+      m1 = mean(m1), 
+      m2 = mean(m2)
+    ) %>% 
+    mutate(sdpooled = sqrt((sd1^2 +sd2^2)/2)) %>% 
+    mutate(ES = (m1 - m2) / sdpooled) #ES estimate = Cohen's d
+  
+} 
+
+########################## GLOBAL PARAMETERS ##########################
+max_n_segments = 3
+alpha_total = 0.05
+d_forpower = 0.5
+
+########################## FULL PROCEDURE FUNCTION ##########################
+
+run_all_procedures = function(nsims){
+  
+  ########################## FIXED SAMPLING PROCEDURE FUNCTION ########################## 
+  
+  run_fixed = function(d_actual, target_power) {
+    
+    ########################## FIXED SAMPLE ##########################
+    
+    # Calculate fixed n per group for the one-tailed two-sample t-test
+    n_fixed = ceiling(
+      power.t.test(
+        delta = d_forpower, 
+        sig.level = alpha_total, 
+        power = target_power, 
+        type = "two.sample", 
+        alternative = "one.sided")$n)
+    
+    # RUN FIXED SAMPLE NHST
+    set.seed(1234*(d_forpower*target_power))
+    fixed = generate_data(
+      nsims = nsims,
[…]

### 2021-11-04 simulations/simulations.R
added supplemental figures 1 and 2

@@ -219,28 +219,12 @@ run_all_procedures = function(nsims){
             ) 
         )
     }) %>% 
-      data.table::rbindlist() 
-    
-    conclusive <- tests %>% 
-      filter(decision!= "inconclusive") %>% 
-      group_by(nsim) %>% 
-      slice_head() #keep all data up to the first segment in which decision was made
-    
-    inconclusive <- tests %>% 
-      filter(decision=="inconclusive" & !nsim %in% conclusive$nsim) %>% 
-      group_by(nsim) %>% 
-      slice_tail() 
-    
-    tests <- bind_rows(conclusive, inconclusive) %>% 
-      arrange(nsim)
-    
-    # maybe just change to this:
-    # data.table::rbindlist() %>% 
-    #   arrange(nsim) %>% 
-    #   group_by(nsim) %>% 
-    #   mutate(final_segment = ifelse(decision != "inconclusive" | segment == 3, 1, 0)) %>% 
-    #   filter(final_segment == 1) %>% 
-    #   slice_head() %>% 
+      data.table::rbindlist() %>% 
+      arrange(nsim) %>%
+      group_by(nsim) %>%
+      mutate(final_segment = ifelse(decision != "inconclusive" | segment == 3, 1, 0)) %>%
+      filter(final_segment == 1) %>%
+      slice_head()
     
     #Get bias corrected ES estimate - first split data into individual segments
     raw_data = lapply(1:length(ns), function(i){

### 2021-11-04 supplemental/figures/figures_supplemental.R

@@ -1,25 +1,28 @@
 # Supplemental Materials
 # Add figures comparing ISP to Fixed (code for these sims in 'FixedvsISPsims.Rmd')
-# Add Google Drive > Independent Segments Procedure > Simulations > Figure 1 (Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing). Make ISP solid, fixed dash. Code in 'Figures.R'
-# Add Google Drive > Independent Segments Procedure > Simulations > Figure 4 (Bias in Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing). Code in 'Figures.Rmd'
 # Add Google Drive > Independent Segments Procedure > Simulations > Figure 7 (Bias-Variance in Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing).  Code in 'Figures.Rmd'
 # Add Google Drive > Independent Segments Procedure > Simulations > Figure 6 (Estimated Cohen's d using Fixed versus Independent Segments Hypothesis Testing).  Code in 'Figures Appendix.Rmd'
 
-
 # Set up ------------------------------------------------------------------
-
 library(ggplot2)
 library(tidyverse)
 library(ggtext)
 library(grid)
+library(gridExtra)
 
-df = read.csv("supplemental/simulations/data_supplemental.csv")
+zipped_df = gzfile("supplemental/simulations/data_supplemental.csv.gz", "rt")
+df = read.csv(zipped_df, header = T) 
 
 # Figure S1 ----------------------------------------------------------------
 # Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing
+# Expected effect sizes only (i.e., d powered for == actual d)
+# Google Drive > Independent Segments Procedure > Simulations > Figure 1 (Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing). Code in 'Figures.R'
+
+df_expected <- df %>% 
+  filter(d_forpower == d_actual)
 
 plt = function(f){
-  D = mean(f$d) #get separate plot for each combo of power and d
+  D = mean(f$d_actual) #get separate plot for each combo of power and d
   target_power = mean(f$power)
   t = sprintf("\u03b4 = %0.1f, 1 - *\u03b2* = %0.1f", D, target_power)
   
@@ -41,30 +44,113 @@ plt = function(f){
     labs(title = t)
 }
 
[…]

### 2021-11-04 supplemental/simulations/simulations_supplemental.R

@@ -2,11 +2,9 @@
 library(data.table)
 library(rpact)
 library(pracma)
-#library(coda)
 library(tidyverse)
 library(parallel)
-#library(TOSTER)
-#library(NCmisc)
+library(TOSTER)
 library(here)
 
 ########################## GLOBAL HELPER FUNCTIONS ##########################
@@ -43,7 +41,8 @@ run_tests = function(df, n, proc, segment = 1, alpha) {
 ########################## GLOBAL PARAMETERS ##########################
 max_n_segments = 3
 alpha_total = 0.05
-d_forpower = 0.5
+
+# Simulations for supplemental data ---------------------------------------
 
 ########################## FULL PROCEDURE FUNCTION ##########################
 
@@ -51,7 +50,7 @@ run_all_procedures = function(nsims){
   
   ########################## FIXED SAMPLING PROCEDURE FUNCTION ########################## 
   
-  run_fixed = function(d_actual, target_power) {
+  run_fixed = function(d_actual, d_forpower, target_power) {
     
     ########################## FIXED SAMPLE ##########################
     
@@ -59,13 +58,14 @@ run_all_procedures = function(nsims){
     n_fixed = ceiling(
       power.t.test(
         delta = d_forpower, 
+        #delta = d_actual, 
         sig.level = alpha_total, 
         power = target_power, 
         type = "two.sample", 
         alternative = "one.sided")$n)
     
     # RUN FIXED SAMPLE NHST
-    set.seed(1234*(d_forpower*target_power))
+    set.seed(1234*(d_actual*target_power))
     fixed = generate_data(
       nsims = nsims,
       n = n_fixed, 
@@ -81,6 +81,7 @@ run_all_procedures = function(nsims){
         n_cumulative = n, 
         ES_corrected = ES,
         d_forpower = d_forpower,
+        #d_forpower = d_actual,
         d_actual = d_actual, 
         power = target_power
       )
@@ -136,12 +137,15 @@ run_all_procedures = function(nsims){
   ########################## RUN FIXED SAMPLING PROCEDURE ##########################
   
   ########################## SET PARAMETERS ##########################
-  d_actual = rep(c(0.2, 0.4, 0.5, 0.6, 0.8), each = 3) 
-  target_power = rep(c(0.7, 0.8, 0.9), 5)
+  target_power = rep(c(0.7, 0.8, 0.9), each = 84)
+  d_actual = rep(rep(seq(0, 1, 0.05), 4), 3)
[…]

### 2021-11-05 supplemental/figures/figures_supplemental.R
created figures S1, S2, and S3

@@ -1,7 +1,4 @@
 # Supplemental Materials
-# Add figures comparing ISP to Fixed (code for these sims in 'FixedvsISPsims.Rmd')
-# Add Google Drive > Independent Segments Procedure > Simulations > Figure 7 (Bias-Variance in Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing).  Code in 'Figures.Rmd'
-# Add Google Drive > Independent Segments Procedure > Simulations > Figure 6 (Estimated Cohen's d using Fixed versus Independent Segments Hypothesis Testing).  Code in 'Figures Appendix.Rmd'
 
 # Set up ------------------------------------------------------------------
 library(ggplot2)
@@ -10,8 +7,18 @@ library(ggtext)
 library(grid)
 library(gridExtra)
 
+
+# Read data ---------------------------------------------------------------
 zipped_df = gzfile("supplemental/simulations/data_supplemental.csv.gz", "rt")
 df = read.csv(zipped_df, header = T) 
+
+df_summary = df %>% 
+  group_by(d_forpower, power, proc, d_actual) %>%  
+  summarize(mean_ES = mean(ES),
+            mse = mean((ES - d_actual)^2),
+            var = mean((ES - mean(ES))^2)) %>% 
+  mutate(bias = mean_ES - d_actual,
+         bias2 = (mean_ES - d_actual)^2) 
 
 # Figure S1 ----------------------------------------------------------------
 # Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing
@@ -49,27 +56,28 @@ p = lapply(split(df_expected, list(df_expected$power, df_expected$d_actual), dro
 p[[1]] = p[[1]] + #manually set legend in first facet
   annotate("segment", x = -0.2, y = 4.5, xend = -.1, yend = 4.5) +
   annotate("segment", x = -0.2, y = 3.5, xend = -.1, yend = 3.5, linetype = "dashed") +
-  annotate(geom = "text", x = -.08, y = 4.3, label = "ISP", hjust = 0) +
+  annotate(geom = "text", x = -.08, y = 4.5, label = "ISP", hjust = 0) +
   annotate(geom = "text", x = -.08, y = 3.5, label = "Fixed", hjust = 0) 
 
-tiff(file="supplemental/figures/figureS1.tiff",width=2100,height=2700, units = "px", res = 300)
-figS1 <- grid.arrange(
[…]

### 2021-11-05 supplemental/figures/figures_supplemental.R
added figure S4

@@ -10,7 +10,7 @@ library(gridExtra)
 
 # Read data ---------------------------------------------------------------
 zipped_df = gzfile("supplemental/simulations/data_supplemental.csv.gz", "rt")
-df = read.csv(zipped_df, header = T) 
+df = read.csv(zipped_df, header = T) #takes a minute
 
 df_summary = df %>% 
   group_by(d_forpower, power, proc, d_actual) %>%  
@@ -23,7 +23,6 @@ df_summary = df %>%
 # Figure S1 ----------------------------------------------------------------
 # Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing
 # Expected effect sizes only (i.e., d powered for == actual d)
-# Google Drive > Independent Segments Procedure > Simulations > Figure 1 (Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing). Code in 'Figures.R'
 
 df_expected <- df %>% 
   filter(d_forpower == d_actual)
@@ -83,8 +82,6 @@ dev.off()
 # Figure S2 ----------------------------------------------------------------
 # Bias in Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing
 # Includes unexpected effect sizes
-# Focus is only on power = 0.8
-# Add Google Drive > Independent Segments Procedure > Simulations > Figure 4 (Bias in Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing). Code in 'Figures.Rmd'
 
 bplot = function(f) {
   ggplot(data = f, 
@@ -142,13 +139,13 @@ dev.off()
 
 # Figure S3 ---------------------------------------------------------------
 # Bias-Variance in Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing
-# Add Google Drive > Independent Segments Procedure > Simulations > Figure 7 (Bias-Variance in Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing).  Code in 'Figures.Rmd'
 
 proc.label = c("Fixed Sample Hypothesis Test", "Independent Segments Procedure")
 names(proc.label) = c("Fixed", "ISP")
 plt = function(f) {
[…]

### 2021-11-05 manuscript.Rmd
Figure 6: Added legend for corrected ES estimate

@@
 
-|  | Group-Sequential Design | Independent Segments Procedure | Sequential Bayes Factor Design |
+|  | Group-Sequential Designs | Independent Segments Procedure | Sequential Bayes Factor Design |
 | ------------- | ------------- | ------------- | ------------- |

### 2021-11-05 figures/figures.R

@@ -1,4 +1,5 @@
-#.rs.restartR() #restart when corrupt
+
+# Set up ------------------------------------------------------------------
 library(beepr)
 library(ggplot2)
 library(tidyverse)
@@ -11,27 +12,36 @@ library(rpact)
 library(BayesFactor)
 library(metafor)
 
-#Insert code section with Ctrl+Shift+R
+# Read data ---------------------------------------------------------------
+df = read.csv("simulations/data.csv")
+
+df = df %>% 
+  mutate(
+    proc = fct_relevel(proc, "Fixed", after = 0),
+    facet = case_when(
+      proc == "Fixed" ~ 1,
+      proc == "Bayes" ~ 2,
+      proc == "ISP" ~ 3,
+      proc == "asP" ~ 4,
+      proc == "asOF"~ 5)
+  )
+
+facet.label = c("Fixed Hypothesis and Equivalence Test", 
+                "Sequential Bayes Factor", 
+                "Independent Segments Procedure", 
+                "Pocock-like GS design",
+                "O'Brien-Fleming-like GS design")
+names(facet.label) = c(1, 2, 3, 4, 5)
 
 # TO DO -------------------------------------------------------------------
-# Figure 6: Add legend for corrected ES estimate 
-# Add line for correct ES estimate BFDA?
 # Check why citations don't render correctly (perhaps add re-add those papers to Zotero?)
-# Change figures after taking out MSPRT (height/width/labels with \n)
-    # Set Fixed Sample Figure to left, and the Sequential Figures to right in a 2x2
-    # OR: Add extra panel for Fixed Sample Test only?
-# Change group-sequential designs figures (binding futility bounds + different rates for OBF changed the N). # Change this^ in manuscript too
-# Change table 2a and 2b code in manuscript
+# Change group-sequential designs figures (binding futility bounds + different rates for OBF changed the N). 
+# Change this^ in manuscript too
 
 # MAYBE -------------------------------------------------------------------
 # Run meta-analysis and include meta-analytic effect size figure 
-# Take figure 9 out of paper
-# Rerun simulations on Sherlock?
-# Run with 90% power (ideally, both?)
-# Make density by segment plot
-# Also vary number of looks 
-# Add a table that show results for all effect sizes (maybe for supplemental?)
[…]

### 2021-11-05 manuscript.Rmd
Changed subscripts in the Bayes plots (fig 2a and 2b)

@@
 
-documentclass     : "apa6"
+documentclass     : "apa7"
 classoption       : "man"
@@
 The goal of GS procedures is to efficiently reject a null hypothesis in favor of an alternative hypothesis.
-In a prototypical group-sequential procedure [e.g., the Pocock correction or the O'Brien-Fleming correction\; @fleming1984; @pocock1977], a researcher is allowed to look at the data several times during the course of a study.
+In a prototypical group-sequential procedure [e.g., the Pocock correction or the O'Brien-Fleming correction, @fleming1984; @pocock1977], a researcher is allowed to look at the data several times during the course of a study.
 
@@
 First, it is symmetric: it allows researchers to examine whether the obtained data is better predicted by, for example, $H_+$ than $H_0$, or vice versa.
-Second, it is highly flexible: it allows researchers to take as many looks at incoming data as desired [even after every study participant; @schonbrodt2017].
+Second, it is highly flexible: it allows researchers to take as many looks at incoming data as desired [even after every study participant\; @schonbrodt2017].
 Third, it does not allow type I and type II error rates to be defined a priori [although simulations can be run to choose an optimal design that controls expected error rates, @schonbrodt2018; @schonbrodt2017; @stefan2020].
@@
 
-To reflect the expectation of a "medium" effect size, the researcher can, for example, choose the default prior distribution [i.e., a Cauchy distribution with scale parameter *r* = $\sqrt2/2$; @morey2011; @schnuerch2020; @schonbrodt2017].
-In addition, the researcher could (for example, on the basis of simulations) choose $BF_{+0}$ = 3 as a decision threshold [commonly referred to as "*moderate evidence*"; @schonbrodt2017].
+To reflect the expectation of a "medium" effect size, the researcher can, for example, choose the default prior distribution [i.e., a Cauchy distribution with scale parameter *r* = $\sqrt2/2$\; @morey2011; @schnuerch2020; @schonbrodt2017].
+In addition, the researcher could (for example, on the basis of simulations) choose $BF_{+0}$ = 3 as a decision threshold [commonly referred to as "*moderate evidence*"\; @schonbrodt2017].
 This decision threshold _t_* implies that the researcher is content to stop sampling once the data is 3 times more likely under $H_+$ than under $H_0$, or vice versa.
@@
 For example, medical trials often have multiple "*endpoints*" that need to be monitored at the same time: a primary endpoint (e.g., treatment success or survival) and multiple secondary -- but exceedingly important -- efficacy or safety endpoints [@wassmer2016].
-An enlightening example can be found in the FDA approval process for carvedilol (a drug developed for congestive heart failure), in which the endpoints were -- among other things -- mortality, distance that can be walked on a treadmill, exercise tolerance, cardiovascular hospitalizations, quality of life, and the patient and physician's global assessment [@fisher1999; @fisher1999a].
+An enlightening example can be found in the FDA approval process for carvedilol (a drug developed for congestive heart failure), in which the endpoints were -- among other things -- mortality, distance that can be walked on a treadmill, exercise tolerance, cardiovascular hospitalizations, quality of life, and the patient and physician's global assessment [@fisher1999].
 
@@
 Controlling (long-run) error rates and quantifying (short-run) evidential strength are both worthy scientific goals.
-However, you can't have one without implicitly abstaining from the other [unless you follow the Likelihood paradigm of statistical inference, which is beyond the scope of this paper; @blume2002; @goodman1988; @schnuerch2020].
-Ultimately, the question is what is more important to the researcher [@lakens2020a]: controlling long-run error rates or quantifying short-run evidential strength [@dienes2011].
+However, you can't have one without implicitly abstaining from the other [unless you follow the Likelihood paradigm of statistical inference, which is beyond the scope of this paper\; @blume2002; @goodman1988; @schnuerch2020].
+Ultimately, the question is what is more important to the researcher [@lakens2021]: controlling long-run error rates or quantifying short-run evidential strength [@dienes2011].
 
@@
 It will be noted, however, that it is possible to overcome this limitation of conventional practice by testing against range predictions rather than a nil-null hypothesis.
-Instead of testing the nil-null hypothesis that the effect is exactly zero, researchers can test whether effects more extreme than a range of predicted or practically important effect sizes can be rejected -- for example, by virtue of an equivalence test or a minimum-effect test [@lakens2018; @lakens2020a].
+Instead of testing the nil-null hypothesis that the effect is exactly zero, researchers can test whether effects more extreme than a range of predicted or practically important effect sizes can be rejected -- for example, by virtue of an equivalence test or a minimum-effect test [@lakens2018; @lakens2021].
 
@@
 ## Figure 2A
-## *Bayes Favor in Favor of the Alternative Hypothesis* 
+## *Bayes Factor in Favor of the Alternative Hypothesis* 
 \noindent

### 2021-11-05 figures/figures.R

@@ -1,4 +1,3 @@
-
 # Set up ------------------------------------------------------------------
 library(beepr)
 library(ggplot2)
@@ -34,13 +33,11 @@ facet.label = c("Fixed Hypothesis and Equivalence Test",
 names(facet.label) = c(1, 2, 3, 4, 5)
 
 # TO DO -------------------------------------------------------------------
-# Check why citations don't render correctly (perhaps add re-add those papers to Zotero?)
 # Change group-sequential designs figures (binding futility bounds + different rates for OBF changed the N). 
 # Change this^ in manuscript too
 
 # MAYBE -------------------------------------------------------------------
 # Run meta-analysis and include meta-analytic effect size figure 
-# Change subscript in the Bayes plots (fig 2a and 2b) (don't remember what I meant by this.. maybe +0?)
 # Change superscript in figures S2 and S3
 
 #=============================================================================#
@@ -380,7 +377,7 @@ dev.off()
 #============================================================================#
 # Adapted from https://www.shinyapps.org/apps/RGraphCompendium/index.php#prior-and-posterior
 # Big thank you to Eric-Jan Wagenmakers & Quentin Gronau for maintaining this site!
-# I took the code from the side above and recoded it from base R to ggplot
+# I took the code from the site above and recoded it from base R to ggplot
 
 .likelihoodShiftedT <- function(par, data) {
   -sum(log(dt((data - par[1])/par[2], par[3])/par[2]))
@@ -490,6 +487,10 @@ dev.off()
   xlim <- c(min(CIlow, range(xticks)[1]), max(range(xticks)[2], CIhigh))
   
   ##ggplot 
+  
+  label1 <- sprintf("BF[1] == '%0.2f'", round(BF10, 2))
+  label2 <- sprintf("BF[0] == '%0.2f'", round(BF01, 2))
+
   ggplot(mapping = aes(x = seq(xlim[1], xlim[2], length.out = 1000), 
                        y = posteriorLine)) +
     geom_line(size = 1) +
@@ -508,10 +509,10 @@ dev.off()
          y = "Density") +
     theme(text = element_text(size = 25),
           axis.ticks.length = unit(.45, "cm")) +
-    annotate("text", x = xticks[2], y = ylim[2] - 0.3, size = 6,
-             label = paste0("BF0+ = ", format(round(BF01, 2), nsmall = 2))) +
[…]

### 2021-11-05 manuscript.Rmd
tried fixing ref errors - close enough

@@
 For example, medical trials often have multiple "*endpoints*" that need to be monitored at the same time: a primary endpoint (e.g., treatment success or survival) and multiple secondary -- but exceedingly important -- efficacy or safety endpoints [@wassmer2016].
-An enlightening example can be found in the FDA approval process for carvedilol (a drug developed for congestive heart failure), in which the endpoints were -- among other things -- mortality, distance that can be walked on a treadmill, exercise tolerance, cardiovascular hospitalizations, quality of life, and the patient and physician's global assessment [@fisher1999].
+An enlightening example can be found in the FDA approval process for carvedilol (a drug developed for congestive heart failure), in which the endpoints were -- among other things -- mortality, distance that can be walked on a treadmill, exercise tolerance, cardiovascular hospitalizations, quality of life, and the patient and physician's global assessment [@fisher1999a].
 
@@
 
-The goal of sequential procedures is to efficiently test hypotheses; the goal is not to properly estimate effect sizes [@lakens2020; @schonbrodt2017].
+The goal of sequential procedures is to efficiently test hypotheses; the goal is not to properly estimate effect sizes [@schonbrodt2017].
 It is well known that early stopping can seriously inflate effect sizes [@fan2004; @miller2020; @pocock1989].

### 2021-11-06 manuscript.Rmd
Updated writing and figures to reflect changes in simulations

@@
 ---
-title             : "Group-Wise Sequential Hypothesis Testing"
+title             : "Group-Wise Sequential Hypothesis Testing: Approaches and Trade-offs"
 shorttitle        : "Sequential Hypothesis Testing"
@@
 I illustrate how these procedures can be implemented in a group-wise manner.
-Using Monte-Carlo simulations, I compare the performance of these procedures in terms of their error control (i.e., false-negative and false-positive decision rates), efficiency (i.e., reduction in required sample size), and accuracy (i.e., bias and variance in effect size estimation).
+Using Monte Carlo simulations, I compare the performance of these procedures in terms of their error control (i.e., false-negative and false-positive decision rates), efficiency (i.e., reduction in required sample size), and accuracy (i.e., bias and variance in effect size estimation).
 I conclude with practical recommendations for researchers who wish to utilize the benefits of sequential hypothesis testing.
@@
 ```{r, include = FALSE}
-n_fixed = ceiling(power.t.test(delta = 0.5, sig.level = .05, power = 0.8, 
-                                 type = "two.sample", alternative = "one.sided")$n)
+n_fixed <- ceiling(
+  power.t.test(
+    delta = 0.5, 
+    sig.level = .05,
+    power = 0.8, 
+    type = "two.sample", 
+    alternative = "one.sided"
+    )$n
+  )
 
-gs = function(proc, return = "n") {
-  bs = ifelse(proc == "asOF", "bsOF", "bsP")
+gs  <- function(proc, return = "n") {
+  bs = ifelse(proc == "asOF", "bsOF", "bsP") #specify beta-spending function
+  if(proc == "asOF") {rates = c(0.50, 0.75, 1)} else{rates = c(1/3, 2/3, 1)} #specify information rates for OBF vs Pocock
 
   # Specify the design
-  design = getDesignGroupSequential(sided = 1,
+  design = getDesignGroupSequential(
+    sided = 1,
     alpha = 0.05, 
@@
     typeOfDesign = proc,
-                                    typeBetaSpending= bs)
-
-  # Add binding futility bounds and the different rates
+    typeBetaSpending= bs,
+    informationRates = rates,
+    bindingFutility = TRUE
+    )
 
   # Get parameters
-  parameters = getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
-  n_gs = ceiling(parameters$maxNumberOfSubjects/3/2) # n per look per group
-  alpha = parameters$criticalValuesPValueScale
-  futility = parameters$futilityBoundsPValueScale %>% round(2)
+  parameters <- getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
+  n_gs = ceiling(c(parameters$numberOfSubjects[1], diff(parameters$numberOfSubjects))/2) # n per look per group
+  alpha <- parameters$criticalValuesPValueScale
+  futility <- parameters$futilityBoundsPValueScale %>% round(2)
 
@@
 
-n_p = gs(proc = "asP")
-alpha_p = gs(proc = "asP", return = "alpha")
-futility_p = gs(proc = "asP", return = "futility")
-n_of = gs(proc = "asOF")
-alpha_of = gs(proc = "asOF", return = "alpha")
-futility_of = gs(proc = "asOF", return = "futility")
+n_p <- gs(proc = "asP")[1]
+alpha_p <- gs(proc = "asP", return = "alpha")
+futility_p <- gs(proc = "asP", return = "futility")
+n_of <- gs(proc = "asOF")
+alpha_of <- gs(proc = "asOF", return = "alpha")
+futility_of <- gs(proc = "asOF", return = "futility")
 ```
@@
 A general, sophisticated approach is provided by an "$\alpha$*-spending function*" [@lan1983; @wassmer2016].
-This approach allows a researcher to have unequally spaced looks at their data (for instance, after the first 20% and the first 75% of the maximum sample size) and to adapt the number of looks as the experiment continues.
+This approach allows a researcher to have unequally spaced looks at their data (for instance, after the first 50% and the first 75% of the maximum sample size) and to adapt the number of looks as the experiment continues.
 
 For example, a researcher who wishes to control their overall type I error rate at 5% and take $k_{max}$ = 3 looks (i.e., two interim looks and one final look) at the data can use the common Pocock-like $\alpha$-spending function to determine what the corrected $\alpha$ level should be for each look (referred to as $\alpha_{1,k}$).
-In this case, $\alpha_{1,1:2}$ = `r alpha_p[1]` for the first two looks and $\alpha_{1,3}$ = `r alpha_p[3]` for the last look (i.e., for a one-tailed test).
+In this case, $\alpha_{1,1}$ = `r alpha_p[1]` for the first look, $\alpha_{1,2}$ = `r alpha_p[2]` for the second look, and $\alpha_{1,3}$ = `r alpha_p[3]` for the last look (i.e., for a one-tailed test).
 If during any of the interim looks $p \le \alpha_1$, the researcher can reject the null hypothesis and end the experiment; this is known as "*stopping for efficacy*".
 
-Because the use of group-sequential procedures allows for early stopping, it generally reduces the average sample size needed (with certain exceptions; e.g., when the true effect size is equal to or close to zero, the use of sequential procedures increases the average sample size needed).
-For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_+:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null-Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group.
-Using the Pocock-like GS design outlined above, the same study can be conducted with only `r n_p` participants per group per look.
-
 In addition to stopping early for efficacy, a researcher can "*stop for futility*" if the observed effect size is too small to warrant further investigation.
-This can be done by manually specifying a lower bound (for example, a lower bound at *d* $\le$ 0 or, equivalently, at *p* $\ge$ .50) or by using a "$\beta$*-spending function*" to determine the critical lower bound for each look (referred to as $\alpha_{0,k}$).
+This can be done by manually specifying a binding or non-binding lower bound (for example, a lower bound at *d* $\le$ 0 or, equivalently, at *p* $\ge$ .50) or by using a "$\beta$*-spending function*" to determine the critical lower bound for each look (referred to as $\alpha_{0,k}$).
 In the remainder of the text, I will use $\alpha_1$ to refer to the critical value that allows a researcher to stop for efficacy, and $\alpha_0$ to refer to the critical value that allows a researcher to stop for futility [@wassmer2016].
@@
 The logic of the $\beta$-spending function is similar to that of the $\alpha$-spending function: critical values that warrant stopping for futility are determined a priori while controlling the overall type II error rate at the desired level (for instance, at a $\beta$ of 20%).
-In other words, the bounds are determined such that the probability of a type II error (i.e., given the number of subjects, the long-run probability of failing to reject the null when the true population effect $\delta$ = 0.5) is $\beta$ %.
+In other words, the bounds are determined such that the probability of a type II error (i.e., given the number of subjects, the long-run probability of failing to reject the null when the true population effect is as hypothesized) is $\beta$ %.
 
@@
 The main difference between the Pocock and the O'Brien-Fleming-like procedure is that the O'Brien-Fleming-like procedure sets a stricter $\alpha$ threshold in earlier looks, with the last look having a threshold closer to the overall $\alpha$.
-Consider the $\alpha$ levels for our example Pocock-like design described above: $\alpha_{1,1:2}$ = `r alpha_p[1]` for the first two looks and $\alpha_{1,3}$ = `r alpha_p[3]` for the last look.
-For the O'Brien-Fleming-like procedure, the critical levels are $\alpha_{1,1}$ = `r alpha_of[1]` for the first look, $\alpha_{1,2}$ = `r alpha_of[2]` for the second look, and $\alpha_{1,3}$ = `r alpha_of[3]` for the last look.
+Consider the $\alpha$ levels for our example Pocock-like design with binding futility bounds: $\alpha_{1,1}$ = `r alpha_p[1]` for the first look, $\alpha_{1,2}$ = `r alpha_p[2]` for the second look, and $\alpha_{1,3}$ = `r alpha_p[3]` for the last look.
+For the O'Brien-Fleming-like procedure with binding futility bounds, the critical levels are $\alpha_{1,1}$ = `r alpha_of[1]` for the first look, $\alpha_{1,2}$ = `r alpha_of[2]` for the second look, and $\alpha_{1,3}$ = `r alpha_of[3]` for the last look.
 
-The futility bounds for our example Pocock-like design are $\alpha_{0,1}$ = `r futility_p[1]` for the first look and $\alpha_{0,2}$ = `r futility_p[2]` for the second look.
-For the O'Brien-Fleming-like design, the futility bounds are $\alpha_{0,1}$ = `r futility_of[1]` for the first look and $\alpha_{0,2}$ = `r futility_of[2]` for the second look.
+The binding futility bounds for our example Pocock-like design are $\alpha_{0,1}$ = `r futility_p[1]` for the first look and $\alpha_{0,2}$ = `r futility_p[2]` for the second look.
+For the O'Brien-Fleming-like design, the binding futility bounds are $\alpha_{0,1}$ = `r futility_of[1]` for the first look and $\alpha_{0,2}$ = `r futility_of[2]` for the second look.
 The researcher continues their study until the observed *p*-value crosses one of the prespecified $\alpha_1$ or $\alpha_0$ thresholds, or until the maximum *N* has been reached.
+
+Because the OBF-like procedure sets such a strict $\alpha$ threshold for the first look, a researcher employing the O'Brien-Fleming-like procedure will have relatively low power, and consequently a low probability of stopping for efficacy, in the first look.
+Thus, when a researcher uses the O'Brien Fleming-like procedure and wishes to be well-powered during the earlier looks, it is often recommended to have the first look take place later in the data collection process (rather than spacing the looks equally; Daniël Lakens, personal communication, December 19, 2020).
+For example, when $k_{max} = 3$, a researcher can decide to equally space their looks when employing a Pocock-like design (i.e., analyze the data after 33%,  67%, and 100% of the data are collected, respectively), while taking unequally spaced looks when employing an O'Brien-Fleming-like design (e.g., analyze the data after 50%, 75%, and 100% of the data are collected, respectively).
+
+Because the use of group-sequential procedures allows for early stopping, it generally reduces the average sample size needed (with certain exceptions; e.g., when the true effect size is equal to or close to zero, the use of sequential procedures tends to increase the average sample size needed).
+For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_+:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group.
+Using the Pocock-like GS design outlined above, the same study can be conducted with only `r n_p` participants per group per look.
+Using the O'Brien-Fleming like design outlined above, this same study can be conducted with `r n_of[1]` participants per group during the first look, and `r n_of[2]` participants per group during the last two looks.
 Below, I briefly outline the decision-theoretic procedure of GS designs.
@@
   *
-1. Randomly sample $N_{gs}$ observations from the specified population(s) 
+1. Randomly sample $N_{\text{gs}}$ observations from the specified population(s) 
 2. Run the test(s) of interest (e.g., a one-tailed, two-sample *t* test) on the obtained sample data
@@
   a. if $p \le \alpha_1$: decide to reject $H_0$ and end the experiment ("stop for efficacy")
-  b. if $p > \ \alpha_0$: decide not to reject $H_0$ and end the experiment  ("stop for futility")
-  c. if $\alpha_1 < p~\le~\alpha_0$: collect $N_{gs}$ additional observations and repeat step 2 on the cumulative data
+  b. if $p > \alpha_0$: decide not to reject $H_0$ and end the experiment  ("stop for futility")
+  c. if $\alpha_1 < p~\le~\alpha_0$: collect $N_{\text{gs}}$ additional observations and repeat step 2 on the cumulative data
 
@@
   a. if $p \le \alpha_1$: decide to reject $H_0$ and end the experiment
-  b. if $p > \ \alpha_1$: decide not to reject $H_0$ and end the experiment
+  b. if $p > \alpha_1$: decide not to reject $H_0$ and end the experiment
 
@@
 
-To reflect the expectation of a "medium" effect size, the researcher can, for example, choose the default prior distribution [i.e., a Cauchy distribution with scale parameter *r* = $\sqrt2/2$\; @morey2011; @schnuerch2020; @schonbrodt2017].
+To reflect the expectation of a "medium" effect size, a researcher can, for example, choose the default prior distribution [i.e., a Cauchy distribution with scale parameter *r* = $\sqrt2/2$\; @morey2011; @schnuerch2020; @schonbrodt2017].
+Alternatively, a researcher can rely on an informed prior distribution that incorporates prior knowledge about the effect size, gathered from expert knowledge, previous work, or theoretical considerations [@stefan2020].
 In addition, the researcher could (for example, on the basis of simulations) choose $BF_{+0}$ = 3 as a decision threshold [commonly referred to as "*moderate evidence*"\; @schonbrodt2017].
@@
   *
-1. Randomly sample $N_{Bayes}$ observations from the specified population(s)
+1. Randomly sample $N_{\text{Bayes}}$ observations from the specified population(s) 
 2. Run the test(s) of interest (e.g., a one-sided default Bayes Factor two-sample *t* test)
@@
 Figures 3A and 3B provide a visual overview of the independent segments procedure, compared to the conventional fixed-sample Neyman-Pearson hypothesis test (Figure 3A shows the conceptual differences; Figure 3B shows our worked example).
-In order to achieve 80% power with an overall $\alpha$ of 5% and a hypothesized population effect size $\delta$ = 0.5, the required sample size $N_{ISP}$ = 25 per group per segment.
+In order to achieve 80% power with an overall $\alpha$ of 5% and a hypothesized population effect size $\delta$ = 0.5, the required sample size $N_{\text{ISP}}$ = 25 per group per segment.
 If the researcher deviates in any way from having exactly 25 participants per group per segment, inferences on the basis of the ISP will be affected.
@@
 
-In sum, the ISP applies only to those settings in which a researcher is comfortable with discarding inconclusive data; can guarantee an exact number of participants, allocated equally across segments; and wishes to test a single, key hypothesis.
+In addition, unlike the Sequential Bayes Factor and the group-sequential designs with futility bounds, the ISP does not incorporate a principled decision rule to reject the alternative hypothesis and/or conclude support for the null hypothesis [@lakens2021a].
+The Bayes Factor is an inherently comparative measure of evidence, and can be used to accept the null hypothesis based on the crossing of a prespecified threshold.
+It is well known that the threshold for concluding evidence of absence on the basis of a Bayes Factor is quite stringent: concluding support for the null demands a relatively large number of observations [@stefan2019].
+Group-sequential designs also allow researchers to reject the alternative hypothesis in a principled manner, by setting futility bounds on the basis of a $\beta$-spending function that controls the overall type II error rate.
+Unlike the SBF and group-sequential procedures, researchers relying on the ISP cannot reject the alternative hypothesis in a manner that either demands crossing a (rather stringent) threshold for absence of evidence or controls overall type II error rates.
+Indeed, Miller and Ulrich [-@miller2020, p.
+6] admit that the Independent Segments Procedure "would not be useful for studies designed to produce evidence for the absence of an effect."
+
+In sum, the ISP applies only to those settings in which a researcher is comfortable with discarding inconclusive data; can guarantee an exact number of participants, allocated equally across segments; wishes to test a single, key hypothesis; and does not desire a principled decision rule to reject the alternative hypothesis.
 Below, I briefly outline the decision-theoretic procedure of the ISP.
@@
   *
-1. Randomly sample $N_{ISP}$ observations from the specified population(s)
+1. Randomly sample $N_{\text{ISP}}$ observations from the specified population(s)
 2. Run the test(s) of interest (e.g., a one-tailed, two-sample *t* test) on the obtained sample data
@@
   b. if $p > \alpha_{weak}$: decide not to reject $H_0$ and end the experiment
-  c. if $\alpha_{strong} < p \le \alpha_{weak}$: discard the data, collect a new batch of $N_{ISP}$ observations, and repeat step 2 on this independent batch of data
+  c. if $\alpha_{strong} < p \le \alpha_{weak}$: discard the data, collect a new batch of $N_{\text{ISP}} observations, and repeat step 2 on this independent batch of data
 
@@
 Thus, it is interesting to evaluate the practical implications of the philosophical differences between frequentist and Bayesian hypothesis testing.
-I will return to this issue in the following section of the paper, in which I compare the error rates of the Sequential Bayes Factor to the fixed-sample Neyman-Pearson hypothesis test, the Pocock-like and O'Brien-Fleming like group-sequential procedures, and the independent segments procedure.
+I will return to this issue in the following section of the paper, in which I compare the error rates of the Sequential Bayes Factor to the fixed-sample Neyman-Pearson hypothesis test, the Pocock-like and O'Brien-Fleming like group-sequential procedures, and the Independent Segments Procedure.
 
@@
 It is well known that early stopping can seriously inflate effect sizes [@fan2004; @miller2020; @pocock1989].
-However, when taking all studies into account (including those that did not terminate early), the bias inherent in both group-sequential designs [@goodman2007; @fan2004] and Sequential Bayes Factors [@schonbrodt2017] is small.
+However, when taking all studies into account (including those that did not terminate early), the bias inherent in both group-sequential designs [@goodman2007; @fan2004] and Sequential Bayes Factors [@schonbrodt2017] tends to be small.
 Thus, group-sequential designs and Sequential Bayes Factors can increase research efficiency at only a slight cost to our ability to estimate the true effect size.
-No such investigation, however, has been conducted for the Independent Segments Procedure [@miller2020].
+No such investigation, however, has been conducted for the Independent Segments Procedure.
 
-In the following section, I use Monte-Carlo simulations to compare group-sequential designs, Sequential Bayes Factors, and the Independent Segments Procedure to the fixed-sample Neyman-Pearson procedure, focusing on their relative error rates, efficiency, and accuracy.
+Of note, because the properties of group-sequential designs are well-studied and understood, bias correction procedures are now widely available for group-sequential designs.
+That is, rather than relying on a naive, uncorrected effect size estimate stemming from a sequential design (e.g., Cohen's *d*), researchers can use a median unbiased estimator (which can be easily implemented in widely used sequential analysis software such as `rpact`).
+A similar bias-correction procedure can be followed for the Sequential Bayes Factor: instead of relying on a naive, uncorrected effect size estimate, a researcher employing a SBF design can use the mean of the posterior distribution as the effect size estimator, which shrinks the estimate in early terminations and thus counteracts the effect size inflation caused by early stopping [@schonbrodt2017].
+No such bias correction procedure, however, exists for the Independent Segments Procedure.
+Thus, it is of great interest to study the accuracy of effect size estimation when relying on the Independent Segments Procedure.
+
+In the following section, I use Monte Carlo simulations to compare group-sequential designs, Sequential Bayes Factors, and the Independent Segments Procedure to the fixed-sample Neyman-Pearson procedure, focusing on their relative error rates, efficiency, and accuracy.
 In a previous contribution, the Sequential Bayes Factor was compared to GS designs in terms of error rates, efficiency, and accuracy [@schonbrodt2017].
@@
 In this paper, I implement the SBF with $k_{max}$ equally spaced looks, for two reasons: (a) to allow for maximum comparability between the sequential procedures under discussion, controlling for as many confounds as possible; and (b) to account for practical considerations: it may very well be impractical for researchers to implement the maximum-flexibility SBF approach and analyze data after every participant.
-Thus, I focus on a slightly different, pragmatic question: For a fixed number of equally spaced looks at data (i.e., what I term a "group-wise" implementation), how do these respective procedures perform in terms of their error rates, efficiency, and accuracy?
+Thus, I focus on a slightly different, pragmatic question: For a fixed number of looks at data (i.e., a "group-wise" implementation), how do these respective procedures perform in terms of their error rates, efficiency, and accuracy?
 
@@
 The procedures will be compared in terms of three main properties: (a) error rates, (b) efficiency, and (c) accuracy in effect size estimation.
-In the following section, I briefly outline how each property is evaluated.
+We will compare the properties of the three sequential procedures to a fixed-sample hypothesis test (i.e., a conventional Null Hypothesis Significance Test) paired with an equivalence test.
+
+For maximum comparability to the sequential procedures that allow some principled form of support for the null hypothesis (i.e., the Sequential Bayes Factor and the Group-Sequential Designs with binding futility bounds), we combine the fixed-sample hypothesis test with an equivalence test.
+Conventionally, the Null Hypothesis Significance test is not able to distinguish between failures to reject the null hypothesis due to evidence of absence or absence of evidence.
+Pairing the fixed sample hypothesis test with an equivalence test overcomes this limitation, by allowing us to reject the presence of a difference more extreme than a smallest effect size of interest [@lakens2018].
+In the following section, I briefly outline how each property under study -- error rates, efficiency, and accuracy in effect size estimation -- is evaluated.
 
@@
 
-1. For the fixed-sample hypothesis test, $\delta$ > 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha$)
+1. For the fixed sample hypothesis test paired with an equivalence test, $\delta$ > 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha$)
 2. For the Pocock and O'Brien-Fleming procedures, $\delta$ > 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha_{1,k}$)
@@
 
-1. For the fixed-sample hypothesis test, $\delta$ = 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha$)
+1. For the fixed-sample hypothesis test paired with an equivalence test, $\delta$ = 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha$)
 2. For the Pocock and O'Brien-Fleming procedures, $\delta$ = 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha_{1,k}$)
@@
 
-1. For the Pocock and O'Brien-Fleming procedures, $\delta$ = 0 and the decision was made to stop for futility (*p* $\ge \alpha_{0,k}$)
-2. For the SBF procedure, $\delta$ = 0 and the decision was made to support the null hypothesis ($BF_{0+} \ge$ _t_*)
-3. For the Independent Segments Procedure, $\delta$ = 0 and the decision was made to stop in an early segment (*p* $> \alpha_{weak}$)
+1.
+For the fixed-sample hypothesis test paired with an equivalence test, $\delta$ = 0 and the decision was made to conclude statistical equivalence [*p* $\le \alpha$ for both *p*-values of a Two One-Sided Tests [TOST] procedure\; @lakens2018]
+2. For the Pocock and O'Brien-Fleming procedures, $\delta$ = 0 and the decision was made to stop for futility (*p* $\ge \alpha_{0,k}$)
+3. For the SBF procedure, $\delta$ = 0 and the decision was made to support the null hypothesis ($BF_{0+} \ge$ _t_*)
+4. For the Independent Segments Procedure, $\delta$ = 0 and the decision was made to stop in an early segment (*p* $> \alpha_{weak}$)
 
@@
 
-1. For the Pocock and O'Brien-Fleming procedures, $\delta$ > 0 and the decision was made to stop for futility (*p* $\ge \alpha_{0,k}$)
-2. For the SBF procedure, $\delta$ > 0 and the decision was made to support the null hypothesis ($BF_{0+} \ge$ _t_*)
-3. For the Independent Segments Procedure, $\delta$ > 0 and the decision was made to stop in an early segment (*p* $> \alpha_{weak}$)
+1. For the fixed-sample hypothesis test paired with an equivalence test, $\delta > 0$ and the decision was made to conclude statistical equivalence ($p_{\text{TOST}, 1:2}$ $\le \alpha$)
+2. For the Pocock and O'Brien-Fleming procedures, $\delta$ > 0 and the decision was made to stop for futility (*p* $\ge \alpha_{0,k}$)
+3. For the SBF procedure, $\delta$ > 0 and the decision was made to support the null hypothesis ($BF_{0+} \ge$ _t_*)
+4. For the Independent Segments Procedure, $\delta$ > 0 and the decision was made to stop in an early segment (*p* $> \alpha_{weak}$)
 
@@
 
-1. For the fixed-sample hypothesis test, the procedure ended in a failure to reject the null hypothesis 
+1.
+For the fixed-sample hypothesis test paired with an equivalence test, the procedure failed to reject the null hypothesis or conclude statistical equivalence ($p_{\text{fixed}} > \alpha$ and $p_{\text{TOST},1:2} > \alpha$ )
 2. For the Pocock and O'Brien-Fleming procedures, $k_{max}$ was reached and ended in a failure to reject the null hypothesis ($p > \ \alpha_1$)
@@
 
-Conventionally, the Null-Hypothesis Significance test is not able to distinguish between failures to reject the null hypothesis due to evidence of absence (a True Negative result) or absence of evidence (an Inconclusive result).
-Therefore, we labeled all failures to reject the null on the basis of a fixed-sample hypothesis test as Inconclusive.
-It will be noted, however, that it is possible to overcome this limitation of conventional practice by testing against range predictions rather than a nil-null hypothesis.
-Instead of testing the nil-null hypothesis that the effect is exactly zero, researchers can test whether effects more extreme than a range of predicted or practically important effect sizes can be rejected -- for example, by virtue of an equivalence test or a minimum-effect test [@lakens2018; @lakens2021].
-
 Procedural efficiency is evaluated by means of the average (expected) sample size across several true effect sizes for each procedure.
-Accuracy in effect size estimation is evaluated by means of the density of empirical effect size estimates, the bias in effect size estimates (i.e., the median estimated effect size compared to the true effect size), the mean squared error of effect size estimates (i.e., the sum of the squared bias and the variance of effect size estimates), and the meta-analytic effect size estimate derived from each procedure.
+Accuracy in effect size estimation is evaluated by means of the density of empirical effect size estimates, the bias in effect size estimates (i.e., the median estimated effect size compared to the true effect size), and the mean squared error of effect size estimates (i.e., the sum of the squared bias and the variance of effect size estimates) derived from each procedure.
+For all procedures, accuracy is evaluated for the obtained, uncorrected effect size estimates --- in addition, for procedures that have a bias-adjusted estimator available (i.e., the Sequential Bayes Factor and the group-sequential designs), I also evaluate the accuracy of the bias-corrected estimates.
 
@@
 
-I simulate populations with a specific standardized mean difference $\delta$ and examine the efficiency (i.e., average sample size required), error rates (i.e., rates of false-positive and false-negative evidence), and accuracy (i.e., of the simulation estimate of the population effect size) of five hypothesis testing procedures: the fixed-sample Neyman-Pearson hypothesis test, the independent segments procedure [@miller2020], a group-sequential design with Pocock-like $\alpha$- and $\beta$-spending functions [@pocock1977; @wassmer2016], a group-sequential design with O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions [@fleming1984; @wassmer2016], and the sequential Bayes Factor [@schonbrodt2017].
+I simulate populations with a specific standardized mean difference $\delta$ and examine the efficiency (i.e., average sample size required), error rates (i.e., rates of false-positive and false-negative evidence), and accuracy (i.e., of the simulation estimate of the population effect size) of five hypothesis testing procedures: the fixed-sample Neyman-Pearson hypothesis test paired with an equivalence test [@lakens2014], the Independent Segments Procedure [@miller2020], a group-sequential design with Pocock-like $\alpha$- and $\beta$-spending functions [@pocock1977; @wassmer2016], a group-sequential design with O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions [@fleming1984; @wassmer2016], and the sequential Bayes Factor [@schonbrodt2017].
 
 To facilitate discussion, I focus on one typical scenario in reporting the simulation results.
-Full results that vary additional parameters are reported in the online supplementary material, and reproducible analysis scripts are available at [link].
-For the frequentist procedures (i.e., the fixed-sample N-P procedure, the ISP, and the Pocock-like and O'Brien-Fleming-like GS designs, respectively) under discussion, the chosen typical scenario corresponds to a one-tailed, two-sample hypothesis test powered to 80% ($\beta$ = 0.2) to detect a hypothesized population effect size of $\delta$ = 0.5 with $\alpha$ = .05.
+Full results that vary additional parameters are reported in the online supplementary material at [link] and reproducible analysis scripts are available at [link].
+For the frequentist procedures (i.e., the fixed-sample N-P procedure paired with an equivalence test, the ISP, and the Pocock-like and O'Brien-Fleming-like GS designs, respectively) under discussion, the chosen typical scenario corresponds to a one-tailed, two-sample hypothesis test powered to 80% ($\beta$ = 0.2) to detect a hypothesized population effect size of $\delta$ = 0.5 with $\alpha$ = .05.
 In the fixed-sample case, these specifications require a sample size of `r n_fixed` observations per group.
+The equivalence bounds to achieve 80% power with N = `r n_fixed` are a lower equivalence bound of $\Delta_L$ = -0.58 and an upper equivalence bound of $\Delta_U$ = 0.58.
 
-For maximum comparability, all sequential procedures (ISP, Pocock-like GS design, O'Brien-Fleming-like GS design, and the SBF) were set to a maximum number of 3 equally spaced looks.
-For the independent segments procedure with $\alpha_{total}$ = .05, $\alpha_{strong}$ = .025, $\alpha_{weak}$ = .28, $\beta_{total}$ = .2 and $k_{max}$ = 3 segments, the required sample size $N_{ISP}$ = 25 per group per segment.
+All sequential procedures, with the exception of the O'Brien-Fleming-like GS design, are set to a maximum number of 3 equally spaced looks (i.e., the interim data are looked at twice, after 33% and 67% of the data are collected, and the final look takes place when all data are collected).
+In contrast, because the O'Brien-Fleming-like GS design sets such a stringent $\alpha$ threshold in the first look, it is often recommended to have the first look take place on a larger fraction of the data (Daniël Lakens, personal communication, December 19, 2020).
+In this case, we implemented the O'Brien-Fleming-like GS design with the two interim looks taking place after 50% and 75% of the data, respectively, are collected.
 
+For the independent segments procedure with $\alpha_{total}$ = .05, $\alpha_{strong}$ = .025, $\alpha_{weak}$ = .28, $\beta_{total}$ = .2 and $k_{max}$ = 3 segments, the required sample size $N_{\text{ISP}}$ = 25 per group per segment.
 For the GS designs with $\alpha_{total}$ = .05, $\beta_{total}$ = .20, $k_{max}$ = 3 looks, an $\alpha$-spending approach was used to determine critical $\alpha_{1,k}$ levels.
@@
 For our worked example, the adjusted $\alpha_{0,k}$ levels correspond to $\alpha_{0,1}$ = `r futility_p[1]` and $\alpha_{0,2}$ = `r futility_p[2]` at the first and second look for the Pocock-like design and $\alpha_{0,1}$ = `r futility_of[1]` and $\alpha_{0,2}$ = `r futility_of[2]` at the first and second look for the O'Brien-Fleming-like design.
-For the Pocock-like design, the required sample size $N_{pocock}$ = `r n_p` per group per look; for the O'Brien-Fleming correction, the required sample size $N_{O'Brien-Fleming}$ = `r n_of` per group per look.
+For the Pocock-like design, the required sample size $N_{\text{Pocock}}$ = `r n_p` per group per look; for the O'Brien-Fleming-like design, the required sample size $N_{\text{O'Brien-Fleming}}$ = `r n_of[1]` per group for the first look, and `r n_of[2]` per group for the last two looks.
 
-For the SBF, I ran simulations to choose an optimal sample size (with $k_{max}$ = 3 equally spaced looks), Bayes Factor threshold, and prior distribution, such that the overall error rate would approach that of the frequentist procedures (i.e., a false negative rate of 20% and a false positive rate of 5%).
-On the basis of these simulations, the minimum sample size was set to $N_{Bayes}$ = 22 per group per look with $k_{max}$ = 3 looks, and the Bayes Factor threshold was set to 3 for $H_+$ and 1/3 for $H_0$.
+The parameters for the Sequential Bayes Factor procedure were chosen to allow for maximum comparability with the frequentist sequential procedures, which correspond to a one-tailed hypothesis test powered to detect a hypothesized population effect size of $\delta$ = 0.5 with $\alpha$ = .05 and $k_{max} = 3$ looks.
+To calculate the Bayes Factor, I used an informed prior (a normal distribution centered at the hypothesized effect size, $\delta = 0.5$).
+To incorporate information about the sidedness of the hypothesized effect, the prior distribution was truncated at $\delta = 0$ to include only positive values.
+Thus, like in the one-sided *t*-tests described above, we compare $H_0:~\delta = 0$ to $H_+:~\delta > 0$.
 
+For the SBF procedure, I ran simulations to choose an optimal sample size (paired to the other procedures under study, with $k_{max}$ = 3 equally spaced looks) and Bayes Factor threshold, such that the overall error rate would approach that of the frequentist procedures (i.e., a false negative rate of 20% and a false positive rate of 5%).
+In the main paper, I report the results of a max-SBF procedure with the Bayes Factor threshold set to 3 for $H_+$ and 1/3 for $H_0$, the sample size set to $N_{\text{Bayes}}$ = 25 per group per look with $k_{max}$ = 3 looks, and a normal prior distribution centered at $\delta = 0.5$ with $\sigma^2 = 0.3$.
+In the supplemental materials at [link], results can be found for simulations that vary additional parameters for the max-SBF procedure.
 Of note, as discussed above, the Sequential Bayes Factor can be implemented much more flexibly (i.e., with as many looks as desired, without a maximum sample size).
-However, to make the sequential procedures as comparable as possible (and to account for practical considerations that researchers may have), each procedure is each implemented with $k_{max}$ = 3 equally spaced looks.
+However, to make the sequential procedures as comparable as possible and to account for practical considerations that researchers may have, each procedure is each implemented with $k_{max}$ = 3 looks.
 
-To calculate the Bayes Factor, I used the one-tailed default Bayes Factor for *t* tests [@morey2011] to compare $H_0:~\delta = 0$ to $H_+:~\delta > 0$.
-A default prior distribution was chosen for the standardized effect size $\delta$ under $H_+$ (i.e., a Cauchy distribution with scale parameter *r* = $\sqrt2/2$).
-This prior was chosen to reflect the expectation of a "medium" effect size [@morey2011; @schnuerch2020; @schonbrodt2017].
-
-I drew random samples from two normal distributions with common variance $\sigma^2$ = 1 and means $\mu_1$ = 0 and $\mu_2 = \delta$ ($\delta$ = -0.2, 0, 0.2, 0.4, 0.5, 0.6, 0.8, 1).
+For the simulations, I drew random samples from two normal distributions with common variance $\sigma^2$ = 1 and means $\mu_1$ = 0 and $\mu_2 = \delta$ ($\delta$ = -0.2, 0, 0.2, 0.4, 0.5, 0.6, 0.8, 1).
 For each hypothesis testing procedure and true effect size $\delta$, 10,000 iterations were simulated (i.e., 5 procedures with 8 distinct true effect sizes and 10,000 replications each, leading to 400,000 observations in total).
@@
   *
-1. Randomly sample $N_{fixed}$ observations from each of the two specified populations
+1. Randomly sample $N_{text{fixed}}$ observations from each of the two specified populations
 2. Run a one-tailed, two-sample *t* test on the mean difference between the two samples
@@
   b. If $p > \alpha$: decide not to reject $H_0$
-3. To estimate the standardized effect size, calculate Cohen's *d* (the difference in sample means divided by their pooled standard deviation)
-4. Repeat 10,000 times
-5. Compute the meta-analytic effect size estimate
+3. Run a one-tailed, two-sample equivalence test (TOST) on the mean difference between the two samples
+  a. If $p_{\text{TOST},1:2} \le \alpha$: decide to reject $H_1$ 
+  b. If $p_{\text{TOST},1:2} > \alpha$: decide not to reject $H_1$ (inconclusive)
+4. To estimate the standardized effect size, calculate Cohen's *d* (the difference in sample means divided by their pooled standard deviation)
+5. Repeat 10,000 times
 
 ## *Pocock-like (O'Brien-Fleming-like) Group-sequential procedures*
-1. Randomly sample $N_{Pocock}$ ($N_{O'Brien-Fleming}$) observations from each of the two specified populations 
+1. Randomly sample $N_{\text{Pocock}}$ ($N_{\text{O'Brien-Fleming}}$) observations from each of the two specified populations 
 2. Run a one-tailed, two-sample *t* test on the mean difference between the two samples
@@
   a. if $p \le \alpha_{1}$: decide to reject $H_0$ and end the experiment ("stop for efficacy")
-  b. if $p > \ \alpha_{0}$: decide not to reject $H_0$ and end the experiment  ("stop for futility")
-  c. if $\alpha_{1} < p~\le~\alpha_{0}$: collect $N_{Pocock}$ ($N_{O'Brien-Fleming}$) additional observations and repeat step 2 on the cumulative data
+  b. if $p > \ \alpha_{0}$: decide to reject $H_1$ and end the experiment  ("stop for futility")
+  c. if $\alpha_{1} < p~\le~\alpha_{0}$: collect $N_{\text{Pocock}}$ ($N_{\text{O'Brien-Fleming}}$) additional observations and repeat step 2 on the cumulative data
 
@@
   a. if $p \le \alpha_{1}$: decide to reject $H_0$ and end the experiment
-  b. if $p > \ \alpha_{1}$: decide not to reject $H_0$ and end the experiment
-5. To estimate the standardized effect size, calculate Cohen's *d*
-6. Repeat 10,000 times
-7. Compute the meta-analytic effect size estimate
+  b. if $p > \ \alpha_{1}$: decide not to reject $H_0$ and end the experiment (inconclusive)
+5. To estimate the naive (i.e., uncorrected) standardized effect size, calculate Cohen's *d*
+6. To estimate the bias-adjusted standardized effect size, calculate the median unbiased effect estimate
+7. Repeat 10,000 times
 
 ## *Sequential-Bayes Factor*
-1. Randomly sample $N_{Bayes}$ observations from each of the two specified populations
-2. Run a one-sided, two-tailed default Bayes Factor *t* test on the mean difference between the two samples
-  a. if $BF_{+0} \ge 3$: decide to reject $H_0$ (support $H_+$) and end the experiment
-  b. if $BF_{+0} \le 1/3$: decide to reject $H_+$ (support $H_0$) and end the experiment
-  c. if $1/3 < BF_{10} < 3$: collect $N_{Bayes}$ additional observations and repeat step 2 on the cumulative data
-3. Continue until the Bayes Factor crosses one of the thresholds, or until the maximum *N* ($N_{Bayes} * k_{max}$) has been reached
-4. To estimate the standardized effect size, calculate the mean of the posterior distribution using a two-sided model [@vandoorn2020]
-5. Repeat 10,000 times 
-6. Compute the meta-analytic effect size estimate
+1. Randomly sample $N_{\text{Bayes}}$ observations from each of the two specified populations
+2. Run a one-sided, two-sample Bayes Factor *t* test on the mean difference between the two samples
+  a. if $BF_{+0} \ge 3$: decide to reject $H_0$ (support $H_1$) and end the experiment
+  b. if $BF_{+0} \le 1/3$: decide to reject $H_1$ (support $H_0$) and end the experiment
+  c. if $1/3 < BF_{10} < 3$: collect $N_{\text{Bayes}}$ additional observations and repeat step 2 on the cumulative data
+3. Continue until the Bayes Factor crosses one of the thresholds, or until the maximum *N* ($N_{\text{Bayes}} * k_{max}$) has been reached
+4. To estimate the uncorrected standardized effect size, calculate Cohen's *d*
+5. To estimate the bias-adjusted standardized effect size, calculate the mean of the posterior distribution using a two-sided model [@vandoorn2020]
+6. Repeat 10,000 times
 
 ## *Independent segments procedure* 
-1. Randomly sample $N_{ISP}$ observations from each of the two specified populations
+1. Randomly sample $N_{\text{ISP}}$ observations from each of the two specified populations
 2. Run a one-tailed, two-sample *t* test on the mean difference between the two samples
@@
   b. if $p > \alpha_{weak}$: decide not to reject $H_0$ and end the experiment
-  c. if $\alpha_{strong} < p \le \alpha_{weak}$: discard the data, collect $N_{ISP}$ new observations, and repeat step 2 on this independent batch of data
+  c. if $\alpha_{strong} < p \le \alpha_{weak}$: discard the data, collect $N_{\text{ISP}}$ new observations, and repeat step 2 on this independent batch of data
 
@@
   b. if $p > \alpha_{weak}$: decide not to reject $H_0$ and end the experiment
-5. To estimate the standardized effect size, calculate Cohen's *d*
+5. To estimate the uncorrected standardized effect size, calculate Cohen's *d*
 6. Repeat 10,000 times
-7. Compute the meta-analytic effect size estimate
 
@@
 As a result, it is near impossible to obtain a credible effect size estimate from the Independent Segments Procedure.
+We are forced to conclude that the Independent Segments Procedure is a novel method for efficiently rejecting the most plausible effect sizes.
 
@@
 <!-- Incorporate points from three new Psych Methods papers (Lakens, Sneurch/Erdfelder, Ulrich/Miller) -->
-<!-- Add writing on the combination of fixed sample NHST and equivalence test. -->
-<!-- Rephrase decision wording for the Fixed Sample Test combined with equivalence test: refer to rejecting the null, rejecting the alternative, and inconclusive. -->
-<!-- Rewrite settings for O'Brien-Fleming: look after 50%, 75%, 100%. -->
-<!-- Add writing on the implemented bias corrections for GS designs - median unbiased ES from rpact -->
-<!-- Add writing on naive effect size estimate Bayes -->
-<!-- Change writing modified SBF prior from Stefan et al. 2020 - informed normal distribution centered on the population effect size (mean 0.5, variance 0.3).  -->
-<!-- Explain oracle prior setup: the population effect size is equal to the mode of the prior distribution on effect size in the alternative model [@stefan2020:p.15]  -->
-<!-- In this paper, the example we use throughout looks at the performance of these procedures under ideal conditions (an oracle prior setup), i.e.
-the performance of these procedures when the true effect size matches the model expectations as closely as possible: the population effect size is either equal to zero or it matches the effect size parameter (SPRT, Fixed, GS-designs) or the mode of the prior distribution on effect size (SBF) in the alternative model [@stefan2020: p.
-15].
--->
-
 <!-- (To make the point that the setup of the simulations is more practical and realistic than previous ones), Show some data on what scientists actually do, e.g., how often they look at their data when they use the Sequential Bayes Factor.
@@
 
-*Note.* The table shows the median effect size estimate across simulations.
-For the Fixed, Pocock, O'Brien-Fleming, and ISP procedure, the effect size measure is Cohen's d.
-For the Sequential Bayes Factor, the effect size measure is the mean of the posterior distribution.
+*Note.* The table shows the median effect size estimate across simulations. The effect size measure is Cohen's d.
 

### 2021-11-06 figures/figures.R

@@ -12,9 +12,9 @@ library(BayesFactor)
 library(metafor)
 
 # Read data ---------------------------------------------------------------
-df = read.csv("simulations/data.csv")
+df <- read.csv("simulations/data.csv")
 
-df = df %>% 
+df <- df %>% 
   mutate(
     proc = fct_relevel(proc, "Fixed", after = 0),
     facet = case_when(
@@ -25,16 +25,17 @@ df = df %>%
       proc == "asOF"~ 5)
   )
 
-facet.label = c("Fixed Hypothesis and Equivalence Test", 
+facet.label <- c("Fixed Hypothesis and Equivalence Test", 
                 "Sequential Bayes Factor", 
                 "Independent Segments Procedure", 
                 "Pocock-like GS design",
                 "O'Brien-Fleming-like GS design")
-names(facet.label) = c(1, 2, 3, 4, 5)
+names(facet.label) <- c(1, 2, 3, 4, 5)
 
 # TO DO -------------------------------------------------------------------
-# Change group-sequential designs figures (binding futility bounds + different rates for OBF changed the N). 
-# Change this^ in manuscript too
+# Fix p-values in Fig 1AB 
+# Reconcile H1 H+ in the paper
+# Include both bias-correct and uncorrected estimates in the manuscript
 
 # MAYBE -------------------------------------------------------------------
 # Run meta-analysis and include meta-analytic effect size figure 
@@ -44,193 +45,318 @@ names(facet.label) = c(1, 2, 3, 4, 5)
 ########################## FIGURE 1A: POCOCK EXAMPLE ########################## 
 #=============================================================================#
 
-alpha_gs = function(proc, return = "n") {
-  bs = ifelse(proc == "asOF", "bsOF", "bsP")
+gs <- function(proc, return = "n") {
+  bs <- ifelse(proc == "asOF", "bsOF", "bsP")
+  if(proc == "asOF") {rates = c(0.50, 0.75, 1)} else{rates = c(1/3, 2/3, 1)} #specify information rates for OBF vs Pocock
+  
   
   # Specify the design
[…]

### 2021-11-06 simulations/simulations.R

@@ -475,7 +475,7 @@ run_all_procedures = function(nsims){
         prior = list("normal", list(prior.mean = 0.5, prior.variance = 0.3)),
         n.min = 25, 
         n.max = 75, 
-        alternative = "greater", 
+        alternative = "greater", #directionality of the prior
         boundary = 3, 
         B = nsims,
         verbose = TRUE, 
@@ -513,9 +513,6 @@ run_all_procedures = function(nsims){
           n2 = n_cumulative)$value 
         ) 
     
-        #ie 2*t/sqrt(degrees of freedom)
-    
-    
     data.table(df)[, .(id, 
                        proc, 
                        segment, 
@@ -525,7 +522,7 @@ run_all_procedures = function(nsims){
                        d_actual = true.ES, 
                        power, 
                        decision, 
-                       ES = emp.ES, #ES = 2*t/sqrt(degrees of freedom) 
+                       ES = emp.ES, #ES = Cohen's d (2*t/sqrt(degrees of freedom)) 
                        ES_corrected)]
     
   }

### 2021-11-06 manuscript.Rmd
typographic changes to text+figures

@@
 Because the use of group-sequential procedures allows for early stopping, it generally reduces the average sample size needed (with certain exceptions; e.g., when the true effect size is equal to or close to zero, the use of sequential procedures tends to increase the average sample size needed).
-For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_+:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group.
+For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_1:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group.
 Using the Pocock-like GS design outlined above, the same study can be conducted with only `r n_p` participants per group per look.
@@
 The SBF procedure has several properties of interest.
-First, it is symmetric: it allows researchers to examine whether the obtained data is better predicted by, for example, $H_+$ than $H_0$, or vice versa.
+First, it is symmetric: it allows researchers to examine whether the obtained data is better predicted by, for example, $H_1$ than $H_0$, or vice versa.
 Second, it is highly flexible: it allows researchers to take as many looks at incoming data as desired [even after every study participant\; @schonbrodt2017].
@@
 
-Again, we consider the example of a researcher running a one-sided, two-sample *t* test between $H_0:~\delta$ = 0 and $H_+:~\delta > 0$.
-To follow the SBF procedure, the researcher needs to prespecify two things: (a) the prior distribution for the standardized effect size $\delta$ under $H_+$ and (b) the decision threshold _t_* for the Bayes Factor.
+Again, we consider the example of a researcher running a one-sided, two-sample *t* test between $H_0:~\delta$ = 0 and $H_1:~\delta > 0$.
+To follow the SBF procedure, the researcher needs to prespecify two things: (a) the prior distribution for the standardized effect size $\delta$ under $H_1$ and (b) the decision threshold _t_* for the Bayes Factor.
 Excellent guidance on these important decisions can be found in @schonbrodt2017.
@@
 Alternatively, a researcher can rely on an informed prior distribution that incorporates prior knowledge about the effect size, gathered from expert knowledge, previous work, or theoretical considerations [@stefan2020].
-In addition, the researcher could (for example, on the basis of simulations) choose $BF_{+0}$ = 3 as a decision threshold [commonly referred to as "*moderate evidence*"\; @schonbrodt2017].
-This decision threshold _t_* implies that the researcher is content to stop sampling once the data is 3 times more likely under $H_+$ than under $H_0$, or vice versa.
-Figure 2A (2B) provides a visual example of a situation in which the Bayes Factor indicates that the data is 3 times more likely under $H_+$ ($H_0$).
+In addition, the researcher could (for example, on the basis of simulations) choose $BF_{10}$ = 3 as a decision threshold [commonly referred to as "*moderate evidence*"\; @schonbrodt2017].
+This decision threshold _t_* implies that the researcher is content to stop sampling once the data is 3 times more likely under $H_1$ than under $H_0$, or vice versa.
+Figure 2A (2B) provides a visual example of a situation in which the Bayes Factor indicates that the data is 3 times more likely under $H_1$ ($H_0$).
 
 The Sequential Bayes Factor and the Group-Sequential Procedures address different questions, each interesting in their own right.
-Group-sequential procedures (as well as the traditional Null Hypothesis Significance Test and the Independent Segments Procedure described below) raise the question "*On the basis of a decision rule under which the long-run probability of rejecting a true null hypothesis will be no more than $\alpha$%, can the null hypothesis be rejected?*" The Sequential Bayes Factor procedure, on the other hand, considers "*On the basis of a prespecified threshold and desired error rates under the assumption of a specific $H_0$ and $H_+$, can we conclude that there is sufficient evidence for $H_0$ or $H_+$?*"
+Group-sequential procedures (as well as the traditional Null Hypothesis Significance Test and the Independent Segments Procedure described below) raise the question "*On the basis of a decision rule under which the long-run probability of rejecting a true null hypothesis will be no more than $\alpha$%, can the null hypothesis be rejected?*" The Sequential Bayes Factor procedure, on the other hand, considers "*On the basis of a prespecified threshold and desired error rates under the assumption of a specific $H_0$ and $H_1$, can we conclude that there is sufficient evidence for $H_0$ or $H_1$?*"
 
@@
 2. Run the test(s) of interest (e.g., a one-sided default Bayes Factor two-sample *t* test)
-  a. if $BF_{+0} \ge t*$: decide to reject $H_0$ (support $H_+$) and end the experiment
-  b. if $BF_{+0} \le 1/t*$: decide to reject $H_+$ (support $H_0$) and end the experiment
-  c. if $1/t*~< BF_{+0} < t*$: collect additional observations and repeat step 2 on the cumulative data
+  a. if $BF_{10} \ge t*$: decide to reject $H_0$ (support $H_1$) and end the experiment
+  b. if $BF_{10} \le 1/t*$: decide to reject $H_1$ (support $H_0$) and end the experiment
+  c. if $1/t*~< BF_{10} < t*$: collect additional observations and repeat step 2 on the cumulative data
 3. Continue until the Bayes Factor crosses one of the thresholds, or until the maximum *N* has been reached
@@
   b. if $p > \alpha_{weak}$: decide not to reject $H_0$ and end the experiment
-  c. if $\alpha_{strong} < p \le \alpha_{weak}$: discard the data, collect a new batch of $N_{\text{ISP}} observations, and repeat step 2 on this independent batch of data
+  c. if $\alpha_{strong} < p \le \alpha_{weak}$: discard the data, collect a new batch of $N_{\text{ISP}}$ observations, and repeat step 2 on this independent batch of data
 
@@
 2. For the Pocock and O'Brien-Fleming procedures, $\delta$ > 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha_{1,k}$)
-3. For the SBF procedure, $\delta$ > 0 and the decision was made to reject the null hypothesis ($BF_{+0}~\ge$ _t_*)
+3. For the SBF procedure, $\delta$ > 0 and the decision was made to reject the null hypothesis ($BF_{10}~\ge$ _t_*)
 4.
@@
 2. For the Pocock and O'Brien-Fleming procedures, $\delta$ = 0 and the decision was made to reject the null hypothesis (*p* $\le \alpha_{1,k}$)
-3. For the SBF procedure, $\delta$ = 0 and the decision was made to reject the null hypothesis ($BF_{+0} \ge$ _t_*)
+3. For the SBF procedure, $\delta$ = 0 and the decision was made to reject the null hypothesis ($BF_{10} \ge$ _t_*)
 4.
@@
 2. For the Pocock and O'Brien-Fleming procedures, $\delta$ = 0 and the decision was made to stop for futility (*p* $\ge \alpha_{0,k}$)
-3. For the SBF procedure, $\delta$ = 0 and the decision was made to support the null hypothesis ($BF_{0+} \ge$ _t_*)
+3. For the SBF procedure, $\delta$ = 0 and the decision was made to support the null hypothesis ($BF_{01} \ge$ _t_*)
 4. For the Independent Segments Procedure, $\delta$ = 0 and the decision was made to stop in an early segment (*p* $> \alpha_{weak}$)
@@
 2. For the Pocock and O'Brien-Fleming procedures, $\delta$ > 0 and the decision was made to stop for futility (*p* $\ge \alpha_{0,k}$)
-3. For the SBF procedure, $\delta$ > 0 and the decision was made to support the null hypothesis ($BF_{0+} \ge$ _t_*)
+3. For the SBF procedure, $\delta$ > 0 and the decision was made to support the null hypothesis ($BF_{01} \ge$ _t_*)
 4. For the Independent Segments Procedure, $\delta$ > 0 and the decision was made to stop in an early segment (*p* $> \alpha_{weak}$)
@@
 2. For the Pocock and O'Brien-Fleming procedures, $k_{max}$ was reached and ended in a failure to reject the null hypothesis ($p > \ \alpha_1$)
-3. For the SBF procedure, $k_{max}$ was reached and neither of the thresholds was met ($1/t*~< BF_{+0} < t*$)
+3. For the SBF procedure, $k_{max}$ was reached and neither of the thresholds was met ($1/t*~< BF_{10} < t*$)
 4. For the Independent Segments Procedure, $k_{max}$ was reached and ended in a failure to reject the null hypothesis (*p* > $\alpha_{weak}$)
@@
 To incorporate information about the sidedness of the hypothesized effect, the prior distribution was truncated at $\delta = 0$ to include only positive values.
-Thus, like in the one-sided *t*-tests described above, we compare $H_0:~\delta = 0$ to $H_+:~\delta > 0$.
+Thus, like in the one-sided *t*-tests described above, we compare $H_0:~\delta = 0$ to $H_1:~\delta > 0$.
 
 For the SBF procedure, I ran simulations to choose an optimal sample size (paired to the other procedures under study, with $k_{max}$ = 3 equally spaced looks) and Bayes Factor threshold, such that the overall error rate would approach that of the frequentist procedures (i.e., a false negative rate of 20% and a false positive rate of 5%).
-In the main paper, I report the results of a max-SBF procedure with the Bayes Factor threshold set to 3 for $H_+$ and 1/3 for $H_0$, the sample size set to $N_{\text{Bayes}}$ = 25 per group per look with $k_{max}$ = 3 looks, and a normal prior distribution centered at $\delta = 0.5$ with $\sigma^2 = 0.3$.
+In the main paper, I report the results of a max-SBF procedure with the Bayes Factor threshold set to 3 for $H_1$ and 1/3 for $H_0$, the sample size set to $N_{\text{Bayes}}$ = 25 per group per look with $k_{max}$ = 3 looks, and a normal prior distribution centered at $\delta = 0.5$ with $\sigma^2 = 0.3$.
 In the supplemental materials at [link], results can be found for simulations that vary additional parameters for the max-SBF procedure.
@@
 2. Run a one-sided, two-sample Bayes Factor *t* test on the mean difference between the two samples
-  a. if $BF_{+0} \ge 3$: decide to reject $H_0$ (support $H_1$) and end the experiment
-  b. if $BF_{+0} \le 1/3$: decide to reject $H_1$ (support $H_0$) and end the experiment
+  a. if $BF_{10} \ge 3$: decide to reject $H_0$ (support $H_1$) and end the experiment
+  b. if $BF_{10} \le 1/3$: decide to reject $H_1$ (support $H_0$) and end the experiment
   c. if $1/3 < BF_{10} < 3$: collect $N_{\text{Bayes}}$ additional observations and repeat step 2 on the cumulative data
@@
 Table 2A describes the rates of True Negative, Inconclusive, and False Positive results when the true population effect size $\delta$ = 0; Table 2B describes the rates of True Positive, Inconclusive, and False Negative results when the true population effect size $\delta$ = 0.50.
-As can be read from the tables, the frequentist procedures all perform as expected: 4-5% of simulations rejected the null hypothesis when $\delta$ = 0 (i.e., a type I error rate of 4-5%; see the False Positive column Table 2A) and 80-81% of simulations rejected the null hypothesis when $\delta$ = 0.5 (i.e., a type II error rate of 19-20%; see the True Positive column of Table 2B).
+As can be read from the tables, the frequentist procedures all perform as expected: 5% of simulations rejected the null hypothesis when $\delta$ = 0 (i.e., a type I error rate of 5%; see the False Positive column Table 2A) and 80-82% of simulations rejected the null hypothesis when $\delta$ = 0.5 (i.e., a type II error rate of 18-20%; see the True Positive column of Table 2B).
 
 In addition, the results speak to the ability of using simulations to design an efficient Bayesian hypothesis testing procedure with attractive long-run properties.
-In this case, the False Positive rate of the SBF procedure is 4% and the True Positive rate of the SBF procedure is 78%.
-In addition, when a true population effect size exists, the SBF procedure consistently leads to a lower False Negative rate (and a higher rate of Inconclusive results) than its frequentist counterparts.
-This is in part due to the high(er) bar that the Bayes Factor requires to support the Null Hypothesis.
-In this case, the alternative hypothesis ($H_+:~\delta > 0$) has a within-model prior centered on zero (i.e., a Cauchy distribution with scale parameter *r* = $\sqrt2/2$).
-For this model to be dominated by the null hypothesis ($H_0:~\delta = 0$) when in fact a true population effect exists, it takes a large number of misleading samples in which the effect size is close to zero.
-Thus, a Bayesian will have a harder time concluding evidence in favor of the null (a False Negative) when a population effect exists.
-
+In this case, the False Positive rate of the SBF procedure is 6% and the True Positive rate of the SBF procedure is 84%.
 In line with previous findings [@schonbrodt2017; @schonbrodt2018; @stefan2020], I find that simulation-based approaches allow researchers to choose an optimal Sequential Bayes Factor design that controls expected error rates.
@@
 Across the procedures, efficiency gains appear to be rather similar.
+When the true population effect size is relatively small, the max-SBF procedure is less efficient than its frequentist counterparts.
 When the true population effect size is relatively large, the O'Brien-Fleming approach is slightly less efficient than its counterparts.
 This can be explained by the relatively stringent $\alpha_1$ levels of the O'Brien-Fleming procedure in earlier segments (recall that, for the O'Brien-Fleming-like procedure, $\alpha_{1,1}$ = `r alpha_of[1]` and $\alpha_{1,2}$ = `r alpha_of[2]`).
+Across all true effect sizes, the Pocock-like procedure, with our chosen parameters, provides the greatest efficiency gains.
 
@@
 
-Figure 6 shows the empirical distribution of effect size estimates stemming from the five hypothesis testing procedures when the true effect size $\delta$ = 0.5.
+Figure 6 shows the empirical distribution of effect size estimates, along with the available bias-adjusted estimates, stemming from the five hypothesis testing procedures when the true effect size $\delta$ = 0.5.
 Effect size estimates stemming from the fixed sample hypothesis test are, as expected, normally distributed around the true effect size ($\delta$ = 0.5).
@@
 
-For the SBF, the median effect size estimate (*d* = 0.51) is very close to the true value ($\delta$ = 0.5).
-The GS procedures each have a slight bias in the median effect size estimate: *d* = 0.54 for the O'Brien-Fleming-like procedure and *d* = 0.57 for the Pocock-like procedure.
+For the group-sequential designs, we can see that the available median-unbiased estimators provide a great improvement over the naive (Cohen's *d*) estimator.
+For both the Pocock-like and the O'Brien-Fleming-like designs, the bias-adjusted effect size estimate is very close to the true population effect size: *d* = 0.49 compared to $\delta$ = 0.50.
+For the SBF procedure, we see that the using the mean of the posterior distribution (which shrinks the effect size estimate in earlier terminations) also provides a slight improvement in parameter estimation (by reducing the estimated effect size from *d* = 0.57 to *d* = 0.55).
 The bias in the median effect size estimate of the ISP, however, is severe (*d* = 0.66; an exaggeration ratio of 1.33).
-The distribution of effect size estimates stemming from the ISP is bi-modal.
+This biased estimate remains uncorrected due to the absence of a bias-adjusted estimator accompanying the ISP.
+A we can see, the distribution of effect size estimates stemming from the ISP is bimodal.
 
@@
 
-In the supplemental files, readers can find that this observation is not merely idiosyncratic to the test chosen (a two-sample *t* test with $\alpha$ = .05 and $\beta = .20$ for a hypothesized population effect size $\delta$ = 0.5).
-In contrast, across dozens of different test specifications, effect size estimates stemming from the ISP are consistently bi-modal and systematically exclude the true population effect size from being observed.
+In figures S1-S4 in the supplemental files, readers can find that this observation is not merely idiosyncratic to the test chosen (a two-sample *t* test with $\alpha$ = .05 and $\beta = .20$ for a hypothesized population effect size $\delta$ = 0.5).
+In contrast, across dozens of different test specifications, effect size estimates stemming from the ISP are consistently bimodal and systematically exclude the true population effect size from being observed.
 
@@
 <!-- Add in discussion/limitation section: We do not discuss adaptive designs, which extend GS designs and allow for even more flexibility while controlling error rates.  -->
+<!-- List recommendations: for OBF, set unequally spaced looks. For SBF, run simulations to choose an optimal sample size, decision threshold, and prior distribution. -->
 <!-- Discuss the power of the individual segments.
@@
 <!-- In the language of decision theory, the ISP is inadmissible: its total squared error risk exceeds that of other hypothesis testing procedures no matter what the true effect size may be.  -->
-<!-- Change writing on Bayes True Negative. Because I switched to an oracle prior setup, true negative rates actually look great -->
 <!-- Also refer to MSE as reducible error  -->
-<!-- Maybe add the figure of bimodality for all ISP parameters? Look through previous commits for this. -->
-<!-- Emphasize that there is no bias correction available for the ISP, while there is for the SBF and GS.
-Also emphasize that the ISP is the only procedure that is *not* median unbiased.
-Check if Lakens' comment makes this point?
--->
-<!-- Incorporate points from three new Psych Methods papers (Lakens, Sneurch/Erdfelder, Ulrich/Miller) -->
+<!-- Emphasize that there is no bias correction available for the ISP, while there is for the SBF and GS.  -->
 <!-- (To make the point that the setup of the simulations is more practical and realistic than previous ones), Show some data on what scientists actually do, e.g., how often they look at their data when they use the Sequential Bayes Factor.
@@
 
-```{r include = F}
+```{r, echo = FALSE}
 E_n = df %>% 
@@
   pivot_wider(names_from = d_actual, names_prefix = "*d* = ", values_from = E_n)
+
+knitr::kable(t3)
 ```
 
-```{r, echo = FALSE}
-knitr::kable(t3)
-```
+*Note.*
 
@@
 
-*Note.* $BF_{+0}$ = 3 indicates that the alternative hypothesis ($H_+:~\delta > 0$) performs three times better at predicting the observed data than the null hypothesis ($H_0:~\delta = 0$).
+*Note.* $BF_{10}$ = 3 indicates that the alternative hypothesis ($H_1:~\delta > 0$) performs three times better at predicting the observed data than the null hypothesis ($H_0:~\delta = 0$).
 Figure based on Jeffreys' Amazing Statistics Program (JASP), adapted from Wagenmakers & Gronau [-@wagenmakers]
@@
 
-*Note.* $BF_{0+}$ = 3 indicates that the null hypothesis ($H_0:~\delta = 0$) performs three times better at predicting the observed data than the  alternative hypothesis ($H_+:~\delta > 0$).
+*Note.* $BF_{01}$ = 3 indicates that the null hypothesis ($H_0:~\delta = 0$) performs three times better at predicting the observed data than the  alternative hypothesis ($H_1:~\delta > 0$).
 Figure based on JASP, adapted from Wagenmakers & Gronau [-@wagenmakers]

### 2021-11-06 figures/figures.R

@@ -33,13 +33,15 @@ facet.label <- c("Fixed Hypothesis and Equivalence Test",
 names(facet.label) <- c(1, 2, 3, 4, 5)
 
 # TO DO -------------------------------------------------------------------
-# Fix p-values in Fig 1AB 
-# Reconcile H1 H+ in the paper
-# Include both bias-correct and uncorrected estimates in the manuscript
+# Include both bias-corrected and uncorrected estimates in the manuscript
+# Write up final results
+# Wrap up discussion/practical recommendations
+# Change Fig 3B according to the changes in 1AB
+# Add equivalence test to Fig 3B
+# Update Notes for figures
 
 # MAYBE -------------------------------------------------------------------
-# Run meta-analysis and include meta-analytic effect size figure 
-# Change superscript in figures S2 and S3
+# Fix p-values in Fig 1AB 
 
 #=============================================================================#
 ########################## FIGURE 1A: POCOCK EXAMPLE ########################## 

### 2021-11-06 simulations/simulations.R

@@ -540,95 +540,7 @@ run_all_procedures = function(nsims){
   
   bayes = data.table::rbindlist(dt)
   
-  # MSPRT with adjusted information rates -----------------------------------
-  
-  ########################### MSPRT ###########################
-  
-  # MSPRT = function(d_actual) {
-  #   
-  #   # Run all data in batches of minN
-  #   set.seed(1234*(d_forpower*target_power))
-  #   l = lapply(1:length(ns), function(i){generate_data(nsims = nsims, n = ns[i], d = d_actual)})
-  #   
-  #   df = data.table::rbindlist(l)[order(nsim)] %>% 
-  #     mutate(segment = rep(lapply(1:length(ns), function(i) {rep(i, times = ns[i])}) %>%  unlist(), nsims),
-  #            n_cumulative = case_when(
-  #              segment == 1 ~ ns[1],
-  #              segment == 2 ~ ns[1] + ns[2],
-  #              segment == 3 ~ sum(ns)
-  #            ))
-  #   
-  #   # Run LR tests
-  #   lr = df %>% 
-  #     split(.$nsim) %>%
-  #     map(~ implement.MSPRT(
-  #       obs1 = .$m1,
-  #       obs2 = .$m2,
-  #       design.MSPRT.object = design_SPRT,
-  #       plot.it = 0, 
-  #       verbose = F
-  #     ))
-  #   
-  #   #Keep all data until we've hit the boundary (if no boundary is hit, keep all data)
-  #   df %>% 
-  #     mutate(n_final = lr %>% 
-  #              map("n1") %>%
-  #              unlist() %>% 
-  #              rep(each = maxN)) %>% 
-  #     filter(n_cumulative <= n_final) %>% 
-  #     group_by(nsim) %>% 
-  #     summarize(sd1 = sd(m1),
-  #               sd2 = sd(m2),
-  #               m1 = mean(m1),
-  #               m2 = mean(m2),
-  #               segment = max(segment),
-  #               n = case_when(
-  #                 segment == 1 ~ ns[1],
-  #                 segment == 2 ~ ns[2],
-  #                 segment == 3 ~ ns[3]),
-  #               n_cumulative = max(n_cumulative)) %>% 
-  #     ungroup() %>% 
-  #     mutate(id = row_number(),
-  #            proc = "SPRT",
-  #            d_forpower = d_forpower, 
-  #            d_actual = d_actual, 
-  #            power = target_power,
-  #            decision = lr %>% 
-  #              map("decision") %>% 
-  #              unlist(),
[…]

### 2021-11-06 supplemental/figures/figures_supplemental.R

@@ -7,7 +7,6 @@ library(ggtext)
 library(grid)
 library(gridExtra)
 
-
 # Read data ---------------------------------------------------------------
 zipped_df = gzfile("supplemental/simulations/data_supplemental.csv.gz", "rt")
 df = read.csv(zipped_df, header = T) #takes a minute
@@ -101,7 +100,7 @@ bplot = function(f) {
                        labels = c(round(-max(abs(f$bias)), 2), 0, round(max(abs(f$bias)), 2)),
                        limits = c(-max(abs(f$bias)), max(abs(f$bias)))) +
     scale_linetype_manual(values=c("dashed", "solid")) + 
-    labs(title = paste0("H\u00b2: \u03B4 = ", 
+    labs(title = paste0("H<sub>1</sub>: \u03B4 = ", 
                         mean(f$d_forpower), 
                         ", 1 - *\u03b2* = ", 
                         mean(f$power))) +
@@ -145,7 +144,7 @@ names(proc.label) = c("Fixed", "ISP")
 plt = function(f) {
   D = mean(f$d_forpower) #get separate plot for each d powered for
   #t = paste0(sprintf("H\u00b2: \u03b4 = %0.1f, 1 - *\u03b2* = ", D), mean(f$power))
-  t = sprintf("H\u00b2: \u03b4 = %0.1f, 1 - *\u03b2* = %0.1f", D, mean(f$power))
+  t = sprintf("H<sub>1</sub>: \u03b4 = %0.1f, 1 - *\u03b2* = %0.1f", D, mean(f$power))
   
   ggplot(data = f, mapping = aes(x = d_actual, y = bias2)) +
     geom_smooth(method = "gam", formula = y ~ s(x, k = 12), se = F, color = "black") +

### 2021-11-06 manuscript.Rmd
cleaning up repo

@@
 
-note: "\\clearpage"
+#note: "\\clearpage"
 
-author_note:
+author_note: |
+  All simulation data, analysis code, and research materials have been made publicly available on the OSF and GitHub and can be accessed at [link] and [link].
 
@@
 library(papaja)
-library(beepr)
 library(tidyverse)
@@
 library(grid)
+library(gridExtra)
 library(rpact)
@@
 library(parallel)
-library(metafor)
 library(citr)
@@
 
+## Transparency and openness
+
+All data, analysis code, and research materials are available at [link]. The study was not preregistered. This manuscript was created using `r cite_r("r-refs.bib")`.
+
 # Results
@@
 Here's some examples on how SBFs are used in practice https://github.com/nicebread/BFDA/blob/master/BFDA_examples.md  -->
-
-\newpage
-
-## Acknowledgement
-This manuscript was created using `r cite_r("r-refs.bib")`.
 

### 2021-11-06 figures/figures.R

@@ -1,15 +1,12 @@
 # Set up ------------------------------------------------------------------
-library(beepr)
 library(ggplot2)
 library(tidyverse)
 library(ggtext)
 library(grid)
 library(papaja)
 library(gridExtra)
-library(metafor) #For an overview and introduction to the package please type: help(metafor).
 library(rpact)
 library(BayesFactor)
-library(metafor)
 
 # Read data ---------------------------------------------------------------
 df <- read.csv("simulations/data.csv")
@@ -39,6 +36,9 @@ names(facet.label) <- c(1, 2, 3, 4, 5)
 # Change Fig 3B according to the changes in 1AB
 # Add equivalence test to Fig 3B
 # Update Notes for figures
+# Put figures S1-S4 in the appendix
+# Write cover letter (note open data + materials badges)
+# Write non-technical abstract
 
 # MAYBE -------------------------------------------------------------------
 # Fix p-values in Fig 1AB 
@@ -359,7 +359,8 @@ p3 <- d.plot(
 
 tiff(file="figures/figure1a.tiff",width=2500,height=800, units = "px", res = 300)
 grid.arrange(p1, p2, p3, nrow = 1, 
-             left = textGrob("Density under H0: \u03b4 = 0", rot = 90, hjust = 0.57,
+             left = textGrob("Density under H0: \u03b4 = 0", 
+                             rot = 90, hjust = 0.57,
                              gp = gpar(fontsize = 10)))
 dev.off()
 
@@ -1162,7 +1163,8 @@ ggplot(data = df_summary,
   theme(legend.position = "none") +
   scale_x_continuous(breaks = c(-0.2, 0, 0.2, 0.4, 0.6, 0.8, 1),
                      limits = c(-0.2, 1)) +
-  labs(x = "True Effect Size (Cohen's d)", y = "Mean Estimated Effect Size - True Effect Size")
+  labs(x = "True Effect Size (Cohen's d)", 
+       y = "Mean Estimated Effect Size - True Effect Size")
 dev.off()
 
 # Median bias
@@ -1246,11 +1248,9 @@ ggplot(data = df_summary, mapping = aes(x = d_actual, y = bias.sq)) +
             size = 2.7)
 dev.off()
 
-#==============================================================================================#
-########################## FIGURE 9: MSE CONDITIONAL ON STOPPING TIME ########################## 
[…]

### 2021-11-07 manuscript.Rmd
wrote discussion+practical considerations, added appendix

@@
 In this practical guide, I outline the commonalities and differences between three sequential hypothesis testing procedures: group-sequential designs, the sequential Bayes Factor, and the Independent Segments Procedure.
-I illustrate how these procedures can be implemented in a group-wise manner.
-Using Monte Carlo simulations, I compare the performance of these procedures in terms of their error control (i.e., false-negative and false-positive decision rates), efficiency (i.e., reduction in required sample size), and accuracy (i.e., bias and variance in effect size estimation).
-I conclude with practical recommendations for researchers who wish to utilize the benefits of sequential hypothesis testing.
+I illustrate how these sequential procedures can be implemented in a group-wise manner, and present two trade-offs of importance to the procedures: long-run error control vs.
+short-run evidence and efficiency in hypothesis testing vs.
+accuracy in parameter estimation.
+Using Monte Carlo simulations, I compare the performance of the procedures in terms of their error control (i.e., false-negative and false-positive decision rates), efficiency (i.e., reduction in required sample size), and accuracy (i.e., bias and variance in effect size estimation).
+I find that effect size estimates stemming from the Independent Segments Procedure are highly inaccurate, and systematically exclude the true population effect size.
+I conclude with practical recommendations for researchers who wish to utilize the benefits of sequential hypothesis testing while retaining their ability to obtain a credible effect size estimate.
 
@@
   )
+
+isp_power <- 
+  power.t.test(
+    delta = 0.5, 
+    sig.level = .025, #alpha_strong
+    n = 25,
+    type = "two.sample", 
+    alternative = "one.sided"
+    )$power
 
@@
 To facilitate discussion, I focus on one typical scenario in reporting the simulation results.
-Full results that vary additional parameters are reported in the online supplementary material at [link] and reproducible analysis scripts are available at [link].
+Full results that vary additional parameters and reproducible analysis scripts are available at [link].
 For the frequentist procedures (i.e., the fixed-sample N-P procedure paired with an equivalence test, the ISP, and the Pocock-like and O'Brien-Fleming-like GS designs, respectively) under discussion, the chosen typical scenario corresponds to a one-tailed, two-sample hypothesis test powered to 80% ($\beta$ = 0.2) to detect a hypothesized population effect size of $\delta$ = 0.5 with $\alpha$ = .05.
@@
 
-For the SBF procedure, I ran simulations to choose an optimal sample size (paired to the other procedures under study, with $k_{max}$ = 3 equally spaced looks) and Bayes Factor threshold, such that the overall error rate would approach that of the frequentist procedures (i.e., a false negative rate of 20% and a false positive rate of 5%).
+For the SBF procedure, I ran simulations to choose an optimal sample size (paired to the other procedures under study; a group-wise implementation with $k_{max}$ = 3 equally spaced looks) and Bayes Factor threshold, such that the overall error rate would approach that of the frequentist procedures (i.e., a false negative rate of 20% and a false positive rate of 5%).
 In the main paper, I report the results of a max-SBF procedure with the Bayes Factor threshold set to 3 for $H_1$ and 1/3 for $H_0$, the sample size set to $N_{\text{Bayes}}$ = 25 per group per look with $k_{max}$ = 3 looks, and a normal prior distribution centered at $\delta = 0.5$ with $\sigma^2 = 0.3$.
-In the supplemental materials at [link], results can be found for simulations that vary additional parameters for the max-SBF procedure.
 Of note, as discussed above, the Sequential Bayes Factor can be implemented much more flexibly (i.e., with as many looks as desired, without a maximum sample size).
-However, to make the sequential procedures as comparable as possible and to account for practical considerations that researchers may have, each procedure is each implemented with $k_{max}$ = 3 looks.
+However, to make the sequential procedures as comparable as possible and to account for practical considerations that researchers may have, each procedure is each implemented in a group-wise manner, with $k_{max}$ = 3 looks.
 
@@
 Here, I show that this observation holds when we use a restricted version of the SBF with $k_{max}$ = 3 equally spaced looks.
+Continuous monitoring may often be impractical [@jennison2000].
+Thus, it is of great interest that the SBF may be implemented in a group-wise manner, with a guaranteed upper limit on sample size, while retaining the ability to control expected error rates.
 
@@
 
-In figures S1-S4 in the supplemental files, readers can find that this observation is not merely idiosyncratic to the test chosen (a two-sample *t* test with $\alpha$ = .05 and $\beta = .20$ for a hypothesized population effect size $\delta$ = 0.5).
+In figures A1-A4 in Appendix A, the reader can find that this observation is not merely idiosyncratic to the test chosen (a two-sample *t* test with $\alpha$ = .05 and $\beta = .20$ for a hypothesized population effect size $\delta$ = 0.5).
 In contrast, across dozens of different test specifications, effect size estimates stemming from the ISP are consistently bimodal and systematically exclude the true population effect size from being observed.
+Because the ISP involves running three independent mini-experiments (rather than accumulating data over time), the individual segments -- with the exception of the last segment, for which a very weak $\alpha$ is set -- are severely underpowered.
+The first $k_{max} - 1$ segments of the ISP tend to have less than 50% power.
+As a result, the critical effect size for these segments will always be higher than the hypothesized effect size.
+In our worked example, the first two individual segments have a power of `r round(isp_power*100)`%.
+To make sure that the hypothesized effect size is not weeded out by the procedure's significance filter, the power of each individual segment must be set to at least 50%.
 
@@
 
+```{r, echo = FALSE}
+df <- read.csv("simulations/data.csv")
+
+mean_mse <- df %>% 
+  group_by(proc, d_actual) %>% 
+  summarize(
+    mse = mean((ES_corrected - d_actual) ^2),
+    var = mean((ES_corrected - mean(ES_corrected))^2),
+    median.ES = median(ES_corrected),
+    ES = mean(ES_corrected)
+    ) %>%
+  mutate(
+    bias = ES - d_actual,
+    median.bias = median.ES - d_actual,
+    bias.sq = (ES - d_actual)^2,
+    median.bias.sq = (median.ES - d_actual)^2
+    ) %>% 
+  group_by(proc) %>% 
+  summarize(mean.mse = mean(mse)) %>% 
+  arrange(desc(mean.mse)) %>% 
+  pull(mean.mse) %>% 
+  round(2)
+```
+
 Figure 8 shows the mean squared error of the four procedures.
@@
 As discussed before, the variance in the effect size estimates stemming from the ISP is caused by the process of conducting three independent mini-experiments, rather than a cumulative study.
+In the language of decision theory, the ISP is inadmissible: its total squared error risk exceeds that of other hypothesis testing procedures (in our worked example, mean MSE = `r mean_mse[1]` for the ISP, compared to `r mean_mse[2]` for the Pocock-like design, `r mean_mse[3]` for the O'Brien-Fleming design, `r mean_mse[4]` for the SBF, and `r mean_mse[5]` for the fixed-sample procedure).
 
@@
 
-In this final section, I present a practical guide for researchers who wish to choose between the sequential procedures under discussion.
+In this final section, I present several practical considerations for researchers who wish to choose between the sequential procedures under discussion.
 Group-sequential designs, Sequential Bayes Factors, and the Independent Segments Procedure each have advantages and disadvantages.
 In deciding between sequential hypothesis testing procedures, researchers need to carefully weigh the advantages and disadvantages of the procedures of interest, with regard to the context in which they are conducting their research [@stefan2020].
-As discussed earlier, the SBF allows researchers to specify their uncertainty regarding the effect size by means of a prior distribution.
-Thus, for researchers who wish to incorporate prior beliefs (e.g., because they work in a subject area with high uncertainty about expected effect sizes), the SBF presents an appealing research strategy.
-On the other hand, if a researcher wants robust (i.e., guaranteed) error control, GS designs present an appealing research strategy.
 
-Although the ISP has some advantages, none of these advantages are unique: efficiency and explicit error control can be achieved through alternative methods, none of which pose the disadvantages unique to the ISP.
-The ISP applies only to settings in which a researcher can guarantee an exact number of participants -- allocated equally across segments -- and wishes to test a single hypothesis.
-More importantly, the researcher in question should be willing to discard valuable data and abstain from the ability to obtain a credible estimate of effect size.
+A great advantage to the procedures under discussion is that they can all be implemented in a group-wise manner, with a guaranteed upper limit on sample size.
+Continuous assessment of data can be a serious practical burden.
+In the field of clinical trial design, early proposals for fully sequential methods that require analysis after every observation -- such as the Sequential Probability Ratio Test -- did not receive widespread acceptance [@pramanik2021].
+It was only when methods were extended to more realistic settings, and allowed for the evaluation of outcomes after groups of participants were observed, that sequential methods became widely used [@jennison2000, p.
+24; @pramanik2021, pp.
+4-5].
+
+Because the O'Brien-Fleming procedure uses very conservative stopping boundaries at earlier looks, the decision rule at the last stage of the experiment (if reached) will be very similar to that of the fixed-sample hypothesis test.
+In our worked example in Figure 1B, we see that the boundary for rejecting $H_0$ at the final look is $\alpha_{1,3}$ = `r alpha_of[3]`.
+To practitioners, this has turned out to be an attractive feature [@jennison2000, p.
+25].
+When using the O'Brien-Fleming correction, it is recommended to collect more than $\displaystyle \frac{1}{k_{max}}$% of the observations before taking the first look, to increases the probability of stopping for efficacy.
+
+Although the ISP has some advantages, none of these advantages are unique.
+Efficiency and explicit error control can be achieved through alternative methods, none of which pose the disadvantages unique to the ISP.
+The ISP applies only to settings in which a researcher wishes to test a single null hypothesis and does not desire a principled decision rule to reject an alternative hypothesis.
+In addition, users of the procedure must guarantee an exact number of participants, allocated equally across segments.
+In realistic settings, however, group sizes and increments in information tend to be unequal and unpredictable [@jennison2000, p.
+26].
+Most importantly, the researcher in question should be willing to discard valuable data and abstain from the ability to obtain a credible estimate of effect size.
 
@@
 Based on our simulations, however, we find that the information loss inherent to the procedure is a bug, not a feature, that leads to severe difficulty in effect size estimation.
+Users of the Independent Segments Procedure are advised to choose their $\alpha_{strong}$ and overall statistical power such that the effect size of interest will not be systematically excluded.
+Ideally, each individual segment should be powered to at least 50% to detect the hypothesized effect size; otherwise, the critical effect size (the effect size needed to reject the null and end the experiment) will be higher than the hypothesized effect size, and the user will thus systematically discard the effect size which they intended to detect.
+This likely means, however, that the users will have to power the procedure to well over 90% (see Figure A1 in the Appendix); consequently, much of the desired efficiency gain will be lost.
 Given the existence of appealing alternatives that have the same advantages (of efficiently testing hypotheses with explicit error control) as the ISP, without its associated disadvantages (of severe difficulty in effect size estimation), the reader is advised to think twice before utilizing the Independent Segments Procedure.
@@
 I show here -- in line with previous work -- that with appropriate design choices, the SBF procedure (implemented with a lower degree of flexibility than in previous investigations) can have comparable error rates to its frequentist alternatives.
-Moreover, despite the trade-off between accurate effect size estimation and efficient hypothesis testing [@goodman2007; @schonbrodt2017], I find that both the SBF procedure and the GS procedure can provide relatively accurate estimates of effect size.
+When using the Sequential Bayes Factor procedure, however, researchers are recommended to run simulations [e.g., using the R package `BFDA`, @R-BFDA] to choose an optimal sample size, decision threshold, and prior distribution.
+
+As discussed earlier, the SBF allows researchers to specify their uncertainty regarding the effect size by means of a prior distribution.
+Thus, for researchers who wish to incorporate prior beliefs (e.g., because they work in a subject area with high uncertainty about expected effect sizes), the SBF presents an appealing research strategy [@stefan2020].
+On the other hand, if a researcher wants robust (i.e., guaranteed) error control, group-sequential designs present an appealing research strategy.
+
+Despite the trade-off between accurate effect size estimation and efficient hypothesis testing [@goodman2007; @schonbrodt2017], I find that both the SBF procedure and the GS procedure can provide relatively accurate estimates of effect size.
 Thus, the Sequential Bayes Factor and Group-Sequential procedures can substantially increase research efficiency at only a slight cost to our ability to estimate an effect size.
@@
 In these contexts, sequential hypothesis testing can provide substantial efficiency gains.
-Group-Sequential designs and the Sequential Bayes Factor procedure are valuable tools for psychological scientists to add to their statistical toolbox.
-
-<!-- # Practical considerations section  -->
-
-<!-- A great advantage to group-sequential designs (compared to the SPRT or the SBF) is that, unlike the SPRT and the SBF, it is not open-ended: you know exactly how many participants you at most will need to sample.
-How many interim looks can researchers actually take?
-Look at how SBF/SPRT are applied in practice.
-In SBF designs, how often is data analyzed in practice?
-Look at papers that use it.
-GS-designs are closed - there is a guaranteed upper limit on sample size.
-"Whereas continuous monitoring is desirable, it is often impractical" [@jennison2000: p.21] Ethical considerations: when an experiment is highly costly or may be unsafe, it is an ethical imperative to terminate a trial as soon as possible - in this case, "ethical considerations prescribe that accumulating data be evaluated frequently" [@jennison2000: p.19] -->
-
-<!-- "Armitage (1954, 1958b, 1975) and Bross (1952, 1958) pioneered the use of sequential methods in the medical field, particularly for comparative clinical trials.
-Initially these plans were fully sequential and did not receive widespread acceptance, perhaps because continuous assessment of study results was often impractical."[@jennison2000: p.24] -->
-
-<!-- Re.
-O'Brien-Fleming: discuss the very conservative stopping boundary at early analysis, which also makes the decision rule similar to the fixed sample test if the last stage is reached -- "features which have turned out to be very appealing to practitioners"[@jennison2000: p.25] -->
-
-<!-- Methods were "extended to the often more realistic setting in which group sizes, or more generally increments in information, are unequal and unpredictable" [@jennison2000: p.26] (ISP can not be applied in realistic settings.) -->
-
-<!-- "From a practical perspective, there are many hypothesis testing contexts where it is not feasible to implement a SPRT.
-For instance, a SPRT cannot be applied when data are not collected sequentially.
-Similarly, it cannot be applied when it is not possible to perform the evaluation as soon as participants are treated...
-To address these challenges, group sequential designs have been developed to allow for the evaluation of patient outcomes only after groups of patients have been observed or at scheduled interim analysis times" [@pramanik2021: p.
-4-5] -->
-
-<!-- "Continuous data monitoring can be a serious practical burden, and the introduction of group sequential tests has led to much wider use of sequential methods.
-(...) Not only are group sequential tests convenient to conduct, they also provide ample opportunity for early stopping and can achieve most of the benefits of fully sequential tests in terms of lower expected sample sizes and shorter average study lengths." [@jennison2000: p.
-38] -->
-
-<!-- # To add -->
-
-<!-- Add in discussion/limitation section: We do not discuss adaptive designs, which extend GS designs and allow for even more flexibility while controlling error rates.  -->
-<!-- List recommendations: for OBF, set unequally spaced looks. For SBF, run simulations to choose an optimal sample size, decision threshold, and prior distribution. -->
-<!-- Discuss the power of the individual segments.
-Practical advice (can lead to guidance on choosing alpha strong; make sure that the alpha strong you set does not exclude the effect size you are interested in.
-Write reveal.effectsize function): You have to make sure that the individual segments are each powered to at least 50 percent to detect the hypothesized effect size: otherwise, the critical effect size will be higher than the hypothesized effect size, and you will thus systematically discard the effect sizes that you intended to detect!
-Show that with the example used, the first two independent segments are powered to ~30%.
-In all but the last segment, the critical effect size (the effect size you need to observe to claim significance) is a factor larger than the actual population effect that you powered your test for.
-It is thus impossible to get an unbiased estimate out of anything but the last segment.
-When you obtain an effect size close to the true effect, the ISP procedure forces you to continue to the last segment, and *throw out that data*.
-(You can even graph this exaggeration ratio: create a table that shows the exact critical effect size for all possible combinations, and then you can just read off that plot that you're setting yourself up for overestimation).
--->
-<!-- Discuss conditional bias (bias in effect size estimates conditional on stopping early) and unconditional bias (bias in effect size estimates across all studies).
-All other procedures have low unconditional bias, when you take the late termination studies into account, because they do cumulative analyses — the late termination studies are able to compensate for the bias in the early stage terminations.
-However, because the independent segments procedure is independent, each individual segment is biased and this is never accounted or compensated for.
-Thus, the ISP is unique among sequential procedures in that it not only has strong conditional bias (like most sequential procedures): it also has strong unconditional bias.
-[In this paper, I focused on two properties: conditional bias (bias in effect size estimates conditional on stopping early) and unconditional bias (bias in effect size estimates across all studies).
-I show that the independent segments procedure is unique among sequential procedures in that it not only has strong conditional bias (like all sequential procedures); it also has strong unconditional bias.] Emphasize that this is also what we see in Figure 3B: we only reject when d > 0.57.
-(This is not the case for, e.g., Pocock)   -->
-<!-- In the language of decision theory, the ISP is inadmissible: its total squared error risk exceeds that of other hypothesis testing procedures no matter what the true effect size may be.  -->
-<!-- Also refer to MSE as reducible error  -->
-<!-- Emphasize that there is no bias correction available for the ISP, while there is for the SBF and GS.  -->
-<!-- (To make the point that the setup of the simulations is more practical and realistic than previous ones), Show some data on what scientists actually do, e.g., how often they look at their data when they use the Sequential Bayes Factor.
-Here's some examples on how SBFs are used in practice https://github.com/nicebread/BFDA/blob/master/BFDA_examples.md  -->
+Group-sequential designs and the Sequential Bayes Factor procedure are valuable tools for psychological scientists to add to their statistical toolbox.
 
@@
 ```{r include = F}
-df = read.csv("simulations/data.csv")
 
@@
 *Note.*
+
+# Appendix A
+
+## Figure A1 
+## *Distribution of Obtained Effect Sizes Using Fixed versus Independent Segments Hypothesis Testing*
+\noindent
+```{r, fig.heigth = 11, fig.width = 5}
+knitr::include_graphics("appendix/figures/figureA1.tiff", dpi = 300)
+```
+
+*Note.* Distributions of effect sizes for the ISP (compared to fixed-sample hypothesis testing), across twelve different combinations of statistical power ($1 - \beta$ = 0.7, 0.8, and 0.9) and population effect size ($\delta$ = 0.2, 0.4, 0.6, 0.8).
+
+\newpage
+
+## Figure A2 
+## *Bias in Obtained Effect Sizes Using Fixed versus Independent Segments Hypothesis Testing*
+\noindent
+```{r}
+knitr::include_graphics("appendix/figures/figureA2.tiff")
+```
+
+*Note.* Bias in obtained effect sizes for the ISP (compared to fixed-sample hypothesis testing), across twelve different combinations of statistical power ($1 - \beta$ = 0.7, 0.8, and 0.9) and hypothesized population effect size ($H_1: \delta$ = 0.2, 0.4, 0.6, 0.8), for a range of true population effect sizes ($0 \le \delta \le 1$).
+
+\newpage
+
+## Figure A3 
+## *Bias-Variance in Obtained Effect Sizes Using Fixed versus Independent Segments Hypothesis Testing*
+\noindent
+```{r, fig.height = 6, fig.width=11}
+knitr::include_graphics("appendix/figures/figureA3.tiff")
+```
+
+*Note.* Bias and variance in obtained effect sizes for the ISP (compared to fixed-sample hypothesis testing), across twelve different combinations of statistical power ($1 - \beta$ = 0.7, 0.8, and 0.9) and hypothesized population effect size ($H_1: \delta$ = 0.2, 0.4, 0.6, 0.8), for a range of true population effect sizes ($0 \le \delta \le 1$).
+
+\newpage
+
+## Figure A4
+## *Exaggeration of Significant Results using Fixed versus Independent Segments Hypothesis Testing*
+\noindent
+```{r, fig.width=9}
+knitr::include_graphics("appendix/figures/figureA4.tiff")
+```
+
+*Note.* The figure shows obtained significant results for the two different hypothesis testing procedures, across twelve different combinations of statistical power ($1 - \beta$ = 0.7, 0.8, and 0.9) and hypothesized population effect size ($H_1: \delta$ = 0.2, 0.4, 0.6, 0.8).
+The number of significant results is held constant across conditions (*N* = 5,000), and the significant results are sorted by the observed Cohen's *d*.
+Results that deviate more than 50% from the true population value are shown in red.
+Each panel denotes the probability of obtaining a Cohen's *d* that is 50% larger than the true population effect size $\delta$.

### 2021-11-07 appendix/figures/figures_appendix.R

@@ -1,0 +1,339 @@
+# Appendix A 
+
+# Set up ------------------------------------------------------------------
+library(ggplot2)
+library(tidyverse)
+library(ggtext)
+library(grid)
+library(gridExtra)
+
+# Read data ---------------------------------------------------------------
+zipped_df = gzfile("appendix/simulations/data_appendix.csv.gz", "rt")
+df = read.csv(zipped_df, header = T) #takes a minute
+
+df_summary = df %>% 
+  group_by(d_forpower, power, proc, d_actual) %>%  
+  summarize(mean_ES = mean(ES),
+            mse = mean((ES - d_actual)^2),
+            var = mean((ES - mean(ES))^2)) %>% 
+  mutate(bias = mean_ES - d_actual,
+         bias2 = (mean_ES - d_actual)^2) 
+
+# Figure A1 ----------------------------------------------------------------
+# Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing
+# Expected effect sizes only (i.e., d powered for == actual d)
+
+df_expected <- df %>% 
+  filter(d_forpower == d_actual)
+
+plt = function(f){
+  D = mean(f$d_actual) #get separate plot for each combo of power and d
+  target_power = mean(f$power)
+  t = sprintf("\u03b4 = %0.1f, 1 - *\u03b2* = %0.1f", D, target_power)
+  
+  ggplot(data = f, aes(x = ES, group = proc, linetype = proc)) +
+    geom_density() +
+    theme_classic() +
+    theme(
+      legend.position = "none", 
+      plot.title = ggtext::element_markdown(size = 11, hjust = 0.5, face = "bold"),
+      axis.title = element_blank(), 
+      text = element_text(size = 11), 
+      axis.text.y = element_blank(),
+      axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1), 
+      axis.ticks.y = element_blank()
+      ) +
+    scale_x_continuous(breaks = c(0, 0.5*D, D, 1.5*D),
+                       limits = c(-D, 2.5*D)) +
+    scale_linetype_manual(values=c("dashed", "solid"))+
+    labs(title = t)
+}
+
+p = lapply(split(df_expected, list(df_expected$power, df_expected$d_actual), drop = TRUE), plt)
+
[…]

### 2021-11-07 appendix/simulations/simulations_appendix.R

@@ -1,0 +1,291 @@
+########################## SET UP ########################## 
+library(data.table)
+library(rpact)
+library(pracma)
+library(tidyverse)
+library(parallel)
+library(TOSTER)
+library(here)
+
+########################## GLOBAL HELPER FUNCTIONS ##########################
+generate_data = function(nsims, n, d) { 
+  # Generate raw data
+  raw = lapply(1:nsims, function(i) {
+    data.table(nsim = i, n = n, d = d,
+               m2 = rnorm(n = n, mean = 0),
+               m1 = rnorm(n = n, mean = 0) + d)})
+  data.table::rbindlist(raw)
+}
+
+run_tests = function(df, n, proc, segment = 1, alpha) {
+  
+  # Run t-tests and create data-frame
+  df %>% 
+    group_by(nsim) %>% 
+    summarize(
+      p = t.test(m1, m2, alternative = "greater", var.equal = T)$p.value,
+      proc = proc, 
+      segment = segment, 
+      alpha = alpha, 
+      n = mean(n), 
+      sd1 = sd(m1), 
+      sd2 = sd(m2),
+      m1 = mean(m1), 
+      m2 = mean(m2)
+    ) %>% 
+    mutate(sdpooled = sqrt((sd1^2 +sd2^2)/2)) %>% 
+    mutate(ES = (m1 - m2) / sdpooled) #ES estimate = Cohen's d
+  
+} 
+
+########################## GLOBAL PARAMETERS ##########################
+max_n_segments = 3
+alpha_total = 0.05
+
+# Simulations for appendix data ---------------------------------------
+
+########################## FULL PROCEDURE FUNCTION ##########################
+
+run_all_procedures = function(nsims){
+  
+  ########################## FIXED SAMPLING PROCEDURE FUNCTION ########################## 
+  
+  run_fixed = function(d_actual, d_forpower, target_power) {
+    
+    ########################## FIXED SAMPLE ##########################
+    
+    # Calculate fixed n per group for the one-tailed two-sample t-test
+    n_fixed = ceiling(
+      power.t.test(
+        delta = d_forpower, 
+        #delta = d_actual, 
+        sig.level = alpha_total, 
+        power = target_power, 
+        type = "two.sample", 
+        alternative = "one.sided")$n)
+    
+    # RUN FIXED SAMPLE NHST
+    set.seed(1234*(d_actual*target_power))
+    fixed = generate_data(
[…]

### 2021-11-07 figures/figures.R

@@ -1,3 +1,9 @@
+# TO DO -------------------------------------------------------------------
+# Include both bias-corrected and uncorrected estimates in the manuscript
+# Write up final results
+# Update Notes for figures
+# Write translational abstract
+
 # Set up ------------------------------------------------------------------
 library(ggplot2)
 library(tidyverse)
@@ -28,20 +34,6 @@ facet.label <- c("Fixed Hypothesis and Equivalence Test",
                 "Pocock-like GS design",
                 "O'Brien-Fleming-like GS design")
 names(facet.label) <- c(1, 2, 3, 4, 5)
-
-# TO DO -------------------------------------------------------------------
-# Include both bias-corrected and uncorrected estimates in the manuscript
-# Write up final results
-# Wrap up discussion/practical recommendations
-# Change Fig 3B according to the changes in 1AB
-# Add equivalence test to Fig 3B
-# Update Notes for figures
-# Put figures S1-S4 in the appendix
-# Write cover letter (note open data + materials badges)
-# Write non-technical abstract
-
-# MAYBE -------------------------------------------------------------------
-# Fix p-values in Fig 1AB 
 
 #=============================================================================#
 ########################## FIGURE 1A: POCOCK EXAMPLE ########################## 
@@ -159,7 +151,7 @@ p1 <- d.plot(n = n_p, alpha = alpha_p[1], t = "Pocock (First Look)") +
       " < d < ",
       alpha_pd[1],
       "\n",
-      signif(alpha_p[1], 2),
+      signif(alpha_p[1], 1),
       " < p < ",
       signif(futility_p[1], 2))
     ) +
@@ -171,7 +163,8 @@ p1 <- d.plot(n = n_p, alpha = alpha_p[1], t = "Pocock (First Look)") +
       "REJECT H0 \n d > ", 
       alpha_pd[1],
       "\n p < ", 
-      gsub("0.", ".", as.character(signif(alpha_p[1], 2)), fixed = T)), 
+      signif(alpha_p[1], 1)
+      ), 
     fontface = "bold", 
     size = 2.2
     ) +
[…]

### 2021-11-07 simulations/simulations.R

@@ -559,14 +559,3 @@ write.csv(df, "simulations/data.csv", row.names = F)
 
 ########################## READ IN DATA ##########################
 df = read.csv("simulations/data.csv")
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
-############################################################################
-############################################################################
-############################################################################
-############################################################################
-############################################################################
-############################################################################
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

### 2021-11-07 supplemental/figures/figures_supplemental.R

@@ -1,337 +1,0 @@
-# Supplemental Materials
-
-# Set up ------------------------------------------------------------------
-library(ggplot2)
-library(tidyverse)
-library(ggtext)
-library(grid)
-library(gridExtra)
-
-# Read data ---------------------------------------------------------------
-zipped_df = gzfile("supplemental/simulations/data_supplemental.csv.gz", "rt")
-df = read.csv(zipped_df, header = T) #takes a minute
-
-df_summary = df %>% 
-  group_by(d_forpower, power, proc, d_actual) %>%  
-  summarize(mean_ES = mean(ES),
-            mse = mean((ES - d_actual)^2),
-            var = mean((ES - mean(ES))^2)) %>% 
-  mutate(bias = mean_ES - d_actual,
-         bias2 = (mean_ES - d_actual)^2) 
-
-# Figure S1 ----------------------------------------------------------------
-# Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing
-# Expected effect sizes only (i.e., d powered for == actual d)
-
-df_expected <- df %>% 
-  filter(d_forpower == d_actual)
-
-plt = function(f){
-  D = mean(f$d_actual) #get separate plot for each combo of power and d
-  target_power = mean(f$power)
-  t = sprintf("\u03b4 = %0.1f, 1 - *\u03b2* = %0.1f", D, target_power)
-  
-  ggplot(data = f, aes(x = ES, group = proc, linetype = proc)) +
-    geom_density() +
-    theme_classic() +
-    theme(
-      legend.position = "none", 
-      plot.title = ggtext::element_markdown(size = 11, hjust = 0.5, face = "bold"),
-      axis.title = element_blank(), 
-      text = element_text(size = 11), 
-      axis.text.y = element_blank(),
-      axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1), 
-      axis.ticks.y = element_blank()
-      ) +
-    scale_x_continuous(breaks = c(0, 0.5*D, D, 1.5*D),
-                       limits = c(-D, 2.5*D)) +
-    scale_linetype_manual(values=c("dashed", "solid"))+
-    labs(title = t)
-}
-
-p = lapply(split(df_expected, list(df_expected$power, df_expected$d_actual), drop = TRUE), plt)
-
[…]

### 2021-11-07 supplemental/simulations/simulations_supplemental.R

@@ -1,296 +1,0 @@
-########################## SET UP ########################## 
-library(data.table)
-library(rpact)
-library(pracma)
-library(tidyverse)
-library(parallel)
-library(TOSTER)
-library(here)
-
-########################## GLOBAL HELPER FUNCTIONS ##########################
-generate_data = function(nsims, n, d) { 
-  # Generate raw data
-  raw = lapply(1:nsims, function(i) {
-    data.table(nsim = i, n = n, d = d,
-               m2 = rnorm(n = n, mean = 0),
-               m1 = rnorm(n = n, mean = 0) + d)})
-  data.table::rbindlist(raw)
-}
-
-run_tests = function(df, n, proc, segment = 1, alpha) {
-  
-  # Run t-tests and create data-frame
-  df %>% 
-    group_by(nsim) %>% 
-    summarize(
-      p = t.test(m1, m2, alternative = "greater", var.equal = T)$p.value,
-      proc = proc, 
-      segment = segment, 
-      alpha = alpha, 
-      n = mean(n), 
-      sd1 = sd(m1), 
-      sd2 = sd(m2),
-      m1 = mean(m1), 
-      m2 = mean(m2)
-    ) %>% 
-    mutate(sdpooled = sqrt((sd1^2 +sd2^2)/2)) %>% 
-    mutate(ES = (m1 - m2) / sdpooled) #ES estimate = Cohen's d
-  
-} 
-
-########################## GLOBAL PARAMETERS ##########################
-max_n_segments = 3
-alpha_total = 0.05
-
-# Simulations for supplemental data ---------------------------------------
-
-########################## FULL PROCEDURE FUNCTION ##########################
-
-run_all_procedures = function(nsims){
-  
-  ########################## FIXED SAMPLING PROCEDURE FUNCTION ########################## 
-  
-  run_fixed = function(d_actual, d_forpower, target_power) {
-    
-    ########################## FIXED SAMPLE ##########################
-    
-    # Calculate fixed n per group for the one-tailed two-sample t-test
-    n_fixed = ceiling(
-      power.t.test(
-        delta = d_forpower, 
-        #delta = d_actual, 
-        sig.level = alpha_total, 
-        power = target_power, 
-        type = "two.sample", 
-        alternative = "one.sided")$n)
-    
-    # RUN FIXED SAMPLE NHST
-    set.seed(1234*(d_actual*target_power))
-    fixed = generate_data(
[…]

### 2021-11-07 manuscript.Rmd
results write-up, included uncorrected+corrected estimates in Table 4

@@
 In our worked example, the first two individual segments have a power of `r round(isp_power*100)`%.
-To make sure that the hypothesized effect size is not weeded out by the procedure's significance filter, the power of each individual segment must be set to at least 50%.
+To make sure that the hypothesized effect size is not consistently weeded out by the procedure's significance filter, the power of each individual segment must be set to at least 50%.
 
 It is enlightening to consider that the Sequential Bayes Factor and Group-Sequential procedures are similarly biased in earlier segments.
-Early stopping acts as a filter that weeds out moderately sized effects: the only effect sizes that are maintained are the ones that are relatively small or in the wrong direction (e.g., when stopping for futility); or the ones that are really large (e.g., when stopping for efficacy).
+Early stopping acts as a filter that weeds out moderately sized effects: the only effect sizes that are maintained are the ones that are relatively small or in the wrong direction (e.g., when stopping for futility), or the ones that are really large (e.g., when stopping for efficacy).
 However, due to the cumulative nature of data collection and analysis, the bias inherent in early terminations tends to be compensated for by studies that terminate at later stages [@goodman2007; @fan2004; @schonbrodt2017].
@@
 As expected, the median effect size estimates stemming from fixed-sample hypothesis tests are equal to the true population effect size.
-Median effect size estimates stemming from the Pocock-like design deviate only slightly from the true population effect size.
-In line with previous findings [@schonbrodt2017], median effect size estimates stemming from the Sequential Bayes Factor slightly underestimate the true population effect size for certain true effect sizes.
-This is because we use the mean of the posterior distribution as the effect size estimator, which shrinks the estimate in early terminations.
+The uncorrected median effect size estimates stemming from the Pocock-like and O'Brien-Fleming-like designs deviate only slightly from the true population effect size.
+When the population effect size is relatively small ($\delta < 0.2$), the Pocock-like and O'Brien-Fleming-like effect size estimates slightly underestimate the true effect; when the population effect size is relatively large ($\delta > 0.4$), the Pocock-like and O'Brien-Fleming-like effect size estimates slightly overestimate the true population effect size.
+As can be read in the Table, the group-sequential median-unbiased estimator, used to adjust for bias in the effect size estimates, provides a remarkable improvement in the accuracy of the parameter estimates.
+In all cases, the group-sequential bias-adjusted effect size estimate is equal to or no more than $d = 0.02$ removed from the true population parameter.
+
+For the unadjusted Bayesian effect size estimates, a similar patterns holds: smaller effect sizes are slightly underestimated, while larger effect sizes are slightly overestimated.
+The bias-adjusted estimate (the mean of the posterior distribution) increases the estimates for smaller population effect sizes (i.e., for early terminations in which the null was accepted), and shrinks the estimates for larger population effect sizes (i.e., for early terminations in which the alternative was accepted).
+Although this provides a slight improvement in the parameter estimate for certain population effect sizes ($\delta$ = 0.2, 0.5, and 0.6), the Bayesian bias-corrected estimates do not perform as well as the group-sequential bias-corrected estimates.
+
 Again, we see that the median effect size estimates stemming from the Independent Segments Procedure deviate most from the true population effect size.
-
 Consider the columns of Table 4 that show the median effect size estimates for the respective procedures when the true population effect size $\delta$ = 0.2, 0.4, or 0.6.
-When the true population effect size $\delta$ = 0.2, the median effect size estimate on the basis of the ISP is *d* = 0.06 (compared to 0.13, 0.13, 0.16, and 0.20 for the SBF, Pocock-like, O'Brien-Fleming-like, and fixed sample procedures, respectively).
-When the true population effect size $\delta$ = 0.4, the median effect size estimate on the basis of the ISP is *d* = 0.57 (compared to 0.45, 0.42, 0.42, and 0.40 for the Pocock-like, O'Brien-Fleming-like procedure, SBF, and fixed sample procedures, respectively).
-When the true population effect size $\delta$ = 0.6, the median effect size estimate on the basis of the ISP is *d* = 0.74 (compared to 0.66, 0.62, 0.60, and 0.58 for the Pocock-like, O'Brien-Fleming-like procedure, fixed sample, and SBF procedures, respectively).
+When the true population effect size $\delta$ = 0.2, the median effect size estimate on the basis of the ISP is *d* = 0.05, compared to the following unadjusted (adjusted) estimates for the Pocock-like, SBF, and O'Brien-Fleming-like procedures: 0.13 (0.20), 0.14 (0.17), and 0.16 (0.20).
+The bias-corrected estimates are no further than *d* = 0.03 removed from the true population value (with the group-sequential estimates not at all deviating from the true value), while the ISP underestimates the population value by *d* = 0.15.
+
+When the true population effect size $\delta$ = 0.4, the median effect size estimate on the basis of the ISP is *d* = 0.57, compared to the following unadjusted (adjusted) estimates for the SBF, Pocock-like, and O'Brien-Fleming-like procedures: 0.47 (0.47), 0.45 (0.39), and 0.45 (0.40).
+The bias-corrected estimates are no further than *d* = 0.07 removed from the true population value (with the group-sequential estimates deviating only *d* = 0.01 from the true value), while the ISP overestimates the population value by *d* = 0.17.
+
+When the true population effect size $\delta$ = 0.6, the median effect size estimate on the basis of the ISP is *d* = 0.74, compared to the following unadjusted (adjusted) estimates for the Pocock-like, SBF, and O'Brien-Fleming-like procedures: 0.67 (0.58), 0.65 (0.62), and 0.64 (0.59).
+The bias-corrected estimates are no further than *d* = 0.02 removed from the true population value, while the ISP overestimates the population value by *d* = 0.14.
 
@@
 Despite the trade-off between accurate effect size estimation and efficient hypothesis testing [@goodman2007; @schonbrodt2017], I find that both the SBF procedure and the GS procedure can provide relatively accurate estimates of effect size.
-Thus, the Sequential Bayes Factor and Group-Sequential procedures can substantially increase research efficiency at only a slight cost to our ability to estimate an effect size.
+Thus, the Sequential Bayes Factor and group-sequential procedures can substantially increase research efficiency at only a slight cost to our ability to estimate an effect size.
+For Sequential Bayes Factors, the mean of the posterior distribution shrinks the effect size estimate in case of early terminations when the population effect size is large.
+However, the mean of the posterior distribution does not always provide an improvement: at times, the adjusted estimate is further removed from the true value than the unadjusted estimate.
+In the case of group-sequential designs, however, we find a remarkable increase in the accuracy of parameter estimation when relying on available median unbiased estimators.
+Thus, users of group-sequential designs are recommended to always use bias-adjusted estimates for parameter estimation.
 
@@
 | **Flexibility** | Moderately flexible | Highly restrictive | Highly flexible |
-| **Acceptance of the null** | Possible | Impossible | Possible |
+| **Rejection of the alternative** | Possible | Impossible | Possible |
 
@@
 ```{r include = F}
+#df <- read.csv("simulations/data.csv")
 
@@
 ```{r, include = FALSE}
-med = df %>% 
-  group_by(name, d_actual) %>% 
-  summarize(medianES = median(ES)) %>% 
+df_long <- df %>% 
+  pivot_longer(
+    cols = c(ES, ES_corrected),
+    names_to = c("ES_type"),
+    values_to = "ES"
+    )
+
+median <- df_long %>% 
+  group_by(name, d_actual, ES_type) %>% 
+  summarize(medianES = round(median(ES), 2)) %>% 
+  rename(Procedure = name) %>% 
+  filter(!((ES_type == "ES_corrected") & (Procedure == "Fixed" | Procedure == "ISP"))) %>% 
+  rowwise() %>% 
+  mutate(
+    name = case_when(
+      ES_type == "ES" ~ as.character(Procedure),
+      ES_type == "ES_corrected" ~ paste(Procedure, "adjusted")
+      )
+    ) %>% 
+  ungroup() %>% 
+  select(name, d_actual, medianES) %>% 
   rename(Procedure = name)
 
-t4 = med %>% mutate(medianES = medianES %>% round(2)) %>%  
-  pivot_wider(names_from = d_actual, names_prefix = "*d* = ", values_from = medianES)
+t4 <- median %>% 
+  pivot_wider(
+    names_from = d_actual, 
+    names_prefix = "*d* = ", 
+    values_from = medianES
+    )
+
+t4
 ```
@@
 ```{r, echo = FALSE}
-knitr::kable(t4) #perhaps add median fold inflation to the tables (i.e., exaggeration ratio)
-```
-
-*Note.* The table shows the median effect size estimate across simulations. The effect size measure is Cohen's d.
+knitr::kable(t4) 
+```
+
+*Note.* The table shows the median effect size estimate across simulations.
+The naive effect size measure is Cohen's *d*.
+For the group-sequential designs, a median unbiased estimator (available in statistical software such as `rpact`) was used to calculate the bias-adjusted estimate; for the Sequential Bayes Factor, the bias-adjusted estimate is the mean of the posterior distribution.
 

### 2021-11-07 manuscript.Rmd
continued results section, added corrected+uncorrected estimates for figures 7-8

@@
 Here, I show that this observation holds when we use a restricted version of the SBF with $k_{max}$ = 3 equally spaced looks.
-Continuous monitoring may often be impractical [@jennison2000].
+Continuous monitoring is often impractical [@jennison2000].
 Thus, it is of great interest that the SBF may be implemented in a group-wise manner, with a guaranteed upper limit on sample size, while retaining the ability to control expected error rates.
@@
 
-Figure 7 visualizes the information captured by Table 4, illustrating the differences between the median estimated effect sizes and true effect sizes.
-We can see that the Sequential Bayes Factor underestimates the true effect size for certain parameters and the Group-Sequential procedures slightly underestimate the true effect when it is smaller ($\delta \le$ 0.2) and slightly overestimate it when it is larger ($\delta \ge$ 0.4).
-As before, the ISP shows the greatest deviations from the true effect size.
+Figure 7 visualizes the information captured by Table 4, illustrating the differences between the median estimated effect sizes and the true effect sizes.
+For the Sequential Bayes Factor, we can see that the uncorrected estimated slightly underestimate the true population effect size when it is small, and slightly overestimate the true population effect size when it is large.
+The bias-adjusted estimate, the mean of the posterior distribution, corrects for this phenomenon -- but it is important to note that the Bayesian bias-adjusted estimates only provide a slight improvement or, for certain parameters, an overcorrection.
+Contrast this with the two bottom row panels for the group-sequential designs.
+Here, we witness that the group-sequential median unbiased estimator lives up to its name.
+As before, the ISP shows the greatest overall deviation from the true effect size.
 
@@
 mean_mse <- df %>% 
+  filter(d_actual == d_forpower) %>% 
   group_by(proc, d_actual) %>% 
@@
 
-Figure 8 shows the mean squared error of the four procedures.
+Figure 8 shows the mean squared error of the five procedures for the naive effect size estimates and the available bias-adjusted estimates.
 The mean squared error is a measure of the accuracy of an estimator, composed of two elements: variance (the spread of effect size estimates from one sample to the other) and squared bias (the difference between the mean estimated effect size and the true population effect size).
@@
 As discussed before, the variance in the effect size estimates stemming from the ISP is caused by the process of conducting three independent mini-experiments, rather than a cumulative study.
-In the language of decision theory, the ISP is inadmissible: its total squared error risk exceeds that of other hypothesis testing procedures (in our worked example, mean MSE = `r mean_mse[1]` for the ISP, compared to `r mean_mse[2]` for the Pocock-like design, `r mean_mse[3]` for the O'Brien-Fleming design, `r mean_mse[4]` for the SBF, and `r mean_mse[5]` for the fixed-sample procedure).
+In the language of decision theory, the ISP is inadmissible: its total squared error risk exceeds that of other hypothesis testing procedures (in our worked example of $\delta$ = 0.5, MSE = `r mean_mse[1]` for the ISP, compared to `r mean_mse[2]` for the Pocock-like design, `r mean_mse[3]` for the O'Brien-Fleming design, `r mean_mse[4]` for the SBF, and `r mean_mse[5]` for the fixed-sample procedure).
 
@@
 | **Control of error rates** | Direct | Direct | Indirect (e.g., through simulation) |
-| **Bias correction procedure**| Available | Does not exist | Available |
+| **Bias-adjusted estimator**| Available | Does not exist | Available |
 | **Flexibility** | Moderately flexible | Highly restrictive | Highly flexible |
-| **Rejection of the alternative** | Possible | Impossible | Possible |
+| **Principled rejection of the alternative** | Possible | Impossible | Possible |
 
@@
 
-*Note.*
+*Note.* Average sample size required for five hypothesis testing procedures across a range of unexpected effect sizes.
+All frequentist procedures were powered to detect an effect of $\delta$ = 0.5.
+For the Sequential Bayes Factor procedure, simulations were run to choose an optimal sample size, decision threshold, and prior distribution, with the goal of approach similar error rates as the frequentist procedures.
 
@@
 The naive effect size measure is Cohen's *d*.
-For the group-sequential designs, a median unbiased estimator (available in statistical software such as `rpact`) was used to calculate the bias-adjusted estimate; for the Sequential Bayes Factor, the bias-adjusted estimate is the mean of the posterior distribution.
+For the group-sequential designs, a median unbiased estimator (available in statistical software such as `rpact`, Wassmer & Pahlke, 2020) was used to calculate the bias-adjusted estimate; for the Sequential Bayes Factor, the bias-adjusted estimate is the mean of the posterior distribution.
 
@@
 
-*Note.* Pocock-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 looks.
-Critical values are determined on the basis of Pocock-like $\alpha$- and $\beta$-spending functions in *R* package *rpact* (Wassmer & Pahlke, 2020).
+*Note.* Pocock-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 equally spaced looks.
+Critical values are determined on the basis of Pocock-like $\alpha$- and $\beta$-spending functions with bindings futility bounds.
 
@@
 
-*Note.* O'Brien-Fleming-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 looks.
-Critical values are determined on the basis of O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions in *R* package *rpact* (Wassmer & Pahlke, 2020).
+*Note.* O'Brien-Fleming-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 unequally spaced looks (the looks were set to 50%, 75%, and 100%, respectively).
+Critical values are determined on the basis of O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions with binding futility bounds.
 
@@
 
-*Note.* Comparing the general procedures.
+*Note.* A general comparison of the fixed sample hypothesis test and the independent segments procedure.
 
@@
 \noindent
-```{r, fig.heigth = 4.72, fig.width = 5}
+```{r, fig.heigth = 4.72, fig.width = 6}
 knitr::include_graphics("figures/figure4.tiff", dpi = 1200)
@@
 \noindent
-```{r, fig.heigth = 4.72, fig.width = 5}
+```{r, fig.heigth = 4.72, fig.width = 6}
 knitr::include_graphics("figures/figure6.tiff", dpi = 300)
@@
 
-*Note.* One-tailed, two-sample *t* tests.
-Fixed Sample, Independent Segments, and Group-Sequential (Pocock- and O'Brien-Fleming-like) Hypothesis Tests are based on $\alpha$ = .05, 1 - $\beta$ = 0.8 for $\delta$ = 0.5.
-*N*s per group per look are 51, 25, 24, and 19, respectively, with a maximum number of 3 looks.
+*Note.* One-tailed, two-sample *t* tests using, for the frequentist procedures,  $\alpha$ = .05, 1 - $\beta$ = 0.8 for $\delta$ = 0.5.
+*N*s per group per look are `r n_fixed`, 25, and `r n_p` for the Fixed Sample, Independent Segments, and Pocock-like hypothesis tests.
+For the O'Brien Fleming test, *N* = `r n_of[1]` per group for the first look, and `r n_of[2]` per group for the last two looks.
+All sequential procedures were set to a maximum number of $k_{max}$ = 3 looks.
 For the Sequential Bayes Factor, data was collected in (at most three) batches of 22 subjects per group, and the procedure was terminated when the one-tailed Bayes Factor was greater than or equal to 3 in favor of the null or the alternative hypothesis.
-Figure includes the median effect size estimate.
-True effect size $\delta$ = 0.5
+Figure includes the median effect size estimate (the solid line is the naive estimate; the dashed line is the bias-adjusted estimate).
+The true effect size $\delta$ = 0.5.
 
@@
 
-## *Bias in Empirical Effect Size Estimates*   
-\noindent
-```{r, fig.heigth = 4.72, fig.width = 5}
-knitr::include_graphics("figures/figure7a.tiff", dpi = 300)
-```
-
-*Note.* Figure displays the mean estimated effect size minus the true effect size.
+## *Median Bias in Empirical Effect Size Estimates*   
+\noindent
+```{r, fig.heigth = 4.72, fig.width = 6}
+knitr::include_graphics("figures/figure7b.tiff", dpi = 300)
+```
+
+*Note.* Figure displays the median estimated effect size minus the true effect size for the five different procedures.
+The median uncorrected effect size estimate is shown in red; the median bias-adjusted effect size estimate, if available, is shown in blue.
 
@@
 \noindent
-```{r, fig.heigth = 4.72, fig.width = 5}
-knitr::include_graphics("figures/figure8.tiff", dpi = 300)
-```
-
-*Note.*
+```{r, fig.heigth = 4.72, fig.width = 7}
+knitr::include_graphics("figures/figure8b.tiff", dpi = 300)
+```
+
+*Note.* Figure displays the squared bias, variance, and mean squared error (the sum of the squared bias and the variance) in empirical effect size estimates stemming from the five procedures.
+Results are split into two panels for the three procedures that have bias corrections available.
+For these three procedures -- the Sequential Bayes Factor, Pocock-like group-sequential design, and O'Brien-Fleming-like group-sequential design -- the top panel shows the bias and variance in the naive effect size estimates; the bottom panel shows the bias and variance in the bias-adjusted estimates.
 
@@
 \noindent
-```{r, fig.heigth = 11, fig.width = 5}
-knitr::include_graphics("appendix/figures/figureA1.tiff", dpi = 300)
+```{r, fig.heigth = 14, fig.width = 6}
+knitr::include_graphics("appendix/figures/figureA1.tiff", dpi = 1200)
 ```

### 2021-11-07 figures/figures.R

@@ -1173,77 +1173,181 @@ dev.off()
 #========================================================================================#
 # How does the procedure perform when d is unexpected?
 
-df_summary = df %>% 
-  group_by(proc, facet, d_actual) %>% 
+df_long <- df %>% 
+  pivot_longer(
+    cols = c(ES, ES_corrected),
+    names_to = c("ES_type"),
+    values_to = "ES"
+  ) 
+
+df_summary <- df_long %>% 
+  group_by(proc, facet, d_actual, ES_type) %>% 
   summarize(
-    mse = mean((ES_corrected - d_actual) ^2),
-    var = mean((ES_corrected - mean(ES_corrected))^2),
-    median.ES = median(ES_corrected),
-    ES = mean(ES_corrected)
-    ) %>%
+    mse = mean((ES - d_actual) ^2),
+    var = mean((ES - mean(ES))^2),
+    median.ES = median(ES),
+    ES = mean(ES)
+  ) %>% 
   mutate(
     bias = ES - d_actual,
     median.bias = median.ES - d_actual,
     bias.sq = (ES - d_actual)^2,
     median.bias.sq = (median.ES - d_actual)^2
-    )
+  ) %>% 
+  filter(!((ES_type == "ES_corrected") & (proc == "Fixed" | proc == "ISP"))) #%>% 
+  
 
-df_summary %>% 
-  group_by(proc) %>% 
-  summarize(mean.mse = mean(mse))
-# Bias
+# Mean Bias
+
+legend <- data.frame(
+  facet = rep(1),
+  label = c("Uncorrected", "Corrected"),
+  y = c(0.08, 0.05),
+  x = rep(0.1),
+  ES_type = c("ES", "ES_corrected")
+)
+
 tiff(file="figures/figure7a.tiff",width=2300,height=1200, units = "px", res = 300)
 ggplot(data = df_summary, 
-       mapping = aes(x = d_actual, y = bias)) +
+       mapping = aes(x = d_actual, y = bias, color = ES_type)) +
   theme_bw() +
   theme(strip.background = element_rect(fill = "white")) +
   geom_hline(yintercept = 0, linetype = "dashed", color = "gray") + 
-  geom_segment(aes(x = d_actual, xend = d_actual, y = 0, yend = bias),
-               color = "red", linetype = "dashed") + 
+  geom_segment(
+    aes(
+      x = d_actual, 
+      xend = d_actual, 
+      y = 0, 
+      yend = bias, 
+      color = ES_type
+      ),
[…]

### 2021-11-08 manuscript.Rmd
final edits before making repo public

@@
   Sequential analysis allows researchers to perform well-powered experiments with smaller sample sizes.
-In this practical guide, I outline the commonalities and differences between three sequential hypothesis testing procedures: group-sequential designs, the sequential Bayes Factor, and the Independent Segments Procedure.
+I outline the commonalities and differences between three sequential hypothesis testing procedures: group-sequential designs, the sequential Bayes Factor, and the Independent Segments Procedure.
 I illustrate how these sequential procedures can be implemented in a group-wise manner, and present two trade-offs of importance to the procedures: long-run error control vs.
@@
 keywords          : "sequential analysis, hypothesis testing, efficiency, estimation, bias"
-wordcount         : "-"
+wordcount         : "11,387"
 
@@
 
-Because the OBF-like procedure sets such a strict $\alpha$ threshold for the first look, a researcher employing the O'Brien-Fleming-like procedure will have relatively low power, and consequently a low probability of stopping for efficacy, in the first look.
+Because the OBF-like procedure sets such a strict $\alpha$ threshold for the first look, a researcher employing the O'Brien-Fleming-like procedure will have relatively low power, and consequently a low probability of stopping for efficacy, during the first look.
 Thus, when a researcher uses the O'Brien Fleming-like procedure and wishes to be well-powered during the earlier looks, it is often recommended to have the first look take place later in the data collection process (rather than spacing the looks equally; Daniël Lakens, personal communication, December 19, 2020).
-For example, when $k_{max} = 3$, a researcher can decide to equally space their looks when employing a Pocock-like design (i.e., analyze the data after 33%,  67%, and 100% of the data are collected, respectively), while taking unequally spaced looks when employing an O'Brien-Fleming-like design (e.g., analyze the data after 50%, 75%, and 100% of the data are collected, respectively).
-
-Because the use of group-sequential procedures allows for early stopping, it generally reduces the average sample size needed (with certain exceptions; e.g., when the true effect size is equal to or close to zero, the use of sequential procedures tends to increase the average sample size needed).
+For example, when $k_{max} = 3$, a researcher can decide to space their looks equally when employing a Pocock-like design (i.e., analyze the data after 33%,  67%, and 100% of the data are collected, respectively), while taking unequally spaced looks when employing an O'Brien-Fleming-like design (e.g., analyze the data after 50%, 75%, and 100% of the data are collected, respectively).
+
+Because the use of group-sequential procedures allows for early stopping, it generally reduces the average sample size needed (with certain exceptions; e.g., when the true effect size is equal to or close to zero, the use of sequential procedures can increase the average sample size needed).
 For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_1:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group.
@@
 ## *Independent segments procedure* 
-Another sequential procedure recently proposed for use in psychological research is the Independent Segments Procedure [ISP, @miller2020].
+Another sequential procedure recently proposed for use in psychological research is the Independent Segments Procedure [ISP\; @miller2020].
 A central property of the ISP is that data is collected in a series of independent experiments ('segments').
-Using other sequential procedures (such as the GS design and the Sequential Bayes Factor), data collection and analysis is cumulative: when an experiment is continued (because the interim data did not cross the pre-specified boundaries for early stopping), a researcher collects more observations, adds these to the observations from previous looks, and calculates an interim test statistic on the basis of all the data collected so far.
+Using other sequential procedures (such as the GS designs and the Sequential Bayes Factor), data collection and analysis is cumulative: when an experiment is continued (because the interim data did not cross the pre-specified boundaries for early stopping), a researcher collects more observations, adds these to the observations from previous looks, and calculates an interim test statistic on the basis of all the data collected so far.
 In contrast, when an independent segments experiment is continued, the researcher discards the previous batch of observations and calculates an interim test statistic on the basis of a new, completely independent, batch of observations.
@@
 For example, medical trials often have multiple "*endpoints*" that need to be monitored at the same time: a primary endpoint (e.g., treatment success or survival) and multiple secondary -- but exceedingly important -- efficacy or safety endpoints [@wassmer2016].
-An enlightening example can be found in the FDA approval process for carvedilol (a drug developed for congestive heart failure), in which the endpoints were -- among other things -- mortality, distance that can be walked on a treadmill, exercise tolerance, cardiovascular hospitalizations, quality of life, and the patient and physician's global assessment [@fisher1999a].
 
@@
 Of note, because the properties of group-sequential designs are well-studied and understood, bias correction procedures are now widely available for group-sequential designs.
-That is, rather than relying on a naive, uncorrected effect size estimate stemming from a sequential design (e.g., Cohen's *d*), researchers can use a median unbiased estimator (which can be easily implemented in widely used sequential analysis software such as `rpact`).
+That is, rather than relying on a naive, uncorrected effect size estimate stemming from a sequential design (e.g., Cohen's *d*), researchers can use a median unbiased estimator [which can be easily implemented in widely used sequential analysis software such as `rpact`\; @R-rpact].
 A similar bias-correction procedure can be followed for the Sequential Bayes Factor: instead of relying on a naive, uncorrected effect size estimate, a researcher employing a SBF design can use the mean of the posterior distribution as the effect size estimator, which shrinks the estimate in early terminations and thus counteracts the effect size inflation caused by early stopping [@schonbrodt2017].
@@
 @schonbrodt2017 show that a highly flexible implementation of the Sequential Bayes Factor -- i.e., the SBF in all its glory, with as many looks as is desired -- is at least as efficient as a default group-sequential design (i.e., a group-sequential approach that does not incorporate binding futility bounds on the basis of a $\beta$-spending approach).
-In this paper, I implement the SBF with $k_{max}$ equally spaced looks, for two reasons: (a) to allow for maximum comparability between the sequential procedures under discussion, controlling for as many confounds as possible; and (b) to account for practical considerations: it may very well be impractical for researchers to implement the maximum-flexibility SBF approach and analyze data after every participant.
+In this paper, I offer a more balanced comparison between the two procedures.
+First, because the group-sequential approach also offers a principled decision rule to reject an alternative hypothesis, I implement the group-sequential designs with binding futility bounds.
+Second, I implement the SBF with $k_{max}$ equally spaced looks, for two reasons: (a) to allow for maximum comparability between the sequential procedures under discussion, and (b) to account for practical considerations; it may very well be impractical for researchers to implement the maximum-flexibility SBF approach and analyze data after every participant.
 Thus, I focus on a slightly different, pragmatic question: For a fixed number of looks at data (i.e., a "group-wise" implementation), how do these respective procedures perform in terms of their error rates, efficiency, and accuracy?
@@
 
-For maximum comparability to the sequential procedures that allow some principled form of support for the null hypothesis (i.e., the Sequential Bayes Factor and the Group-Sequential Designs with binding futility bounds), we combine the fixed-sample hypothesis test with an equivalence test.
+For maximum comparability to the sequential procedures that allow some principled form of support for the null hypothesis (i.e., the Sequential Bayes Factor and the group-sequential designs with binding futility bounds), we combine the fixed-sample hypothesis test with an equivalence test.
 Conventionally, the Null Hypothesis Significance test is not able to distinguish between failures to reject the null hypothesis due to evidence of absence or absence of evidence.
@@
 Accuracy in effect size estimation is evaluated by means of the density of empirical effect size estimates, the bias in effect size estimates (i.e., the median estimated effect size compared to the true effect size), and the mean squared error of effect size estimates (i.e., the sum of the squared bias and the variance of effect size estimates) derived from each procedure.
-For all procedures, accuracy is evaluated for the obtained, uncorrected effect size estimates --- in addition, for procedures that have a bias-adjusted estimator available (i.e., the Sequential Bayes Factor and the group-sequential designs), I also evaluate the accuracy of the bias-corrected estimates.
+For all procedures, accuracy is evaluated for the obtained, uncorrected effect size estimates.
+In addition, for procedures that have a bias-adjusted estimator available (i.e., the Sequential Bayes Factor and the group-sequential designs), I also evaluate the accuracy of the bias-corrected estimates.
 
@@
 
-I simulate populations with a specific standardized mean difference $\delta$ and examine the efficiency (i.e., average sample size required), error rates (i.e., rates of false-positive and false-negative evidence), and accuracy (i.e., of the simulation estimate of the population effect size) of five hypothesis testing procedures: the fixed-sample Neyman-Pearson hypothesis test paired with an equivalence test [@lakens2014], the Independent Segments Procedure [@miller2020], a group-sequential design with Pocock-like $\alpha$- and $\beta$-spending functions [@pocock1977; @wassmer2016], a group-sequential design with O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions [@fleming1984; @wassmer2016], and the sequential Bayes Factor [@schonbrodt2017].
+I simulate populations with a specific standardized mean difference $\delta$ and examine the efficiency (i.e., average sample size required), error rates (i.e., rates of false-positive and false-negative evidence), and accuracy (i.e., of the simulation estimate of the population effect size) of five hypothesis testing procedures: the fixed-sample Neyman-Pearson hypothesis test paired with an equivalence test [@lakens2014], a group-sequential design with Pocock-like $\alpha$- and $\beta$-spending functions [@pocock1977; @wassmer2016], a group-sequential design with O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions [@fleming1984; @wassmer2016], the sequential Bayes Factor [@schonbrodt2017], and the Independent Segments Procedure [@miller2020].
 
 To facilitate discussion, I focus on one typical scenario in reporting the simulation results.
-Full results that vary additional parameters and reproducible analysis scripts are available at [link].
 For the frequentist procedures (i.e., the fixed-sample N-P procedure paired with an equivalence test, the ISP, and the Pocock-like and O'Brien-Fleming-like GS designs, respectively) under discussion, the chosen typical scenario corresponds to a one-tailed, two-sample hypothesis test powered to 80% ($\beta$ = 0.2) to detect a hypothesized population effect size of $\delta$ = 0.5 with $\alpha$ = .05.
@@
 In contrast, because the O'Brien-Fleming-like GS design sets such a stringent $\alpha$ threshold in the first look, it is often recommended to have the first look take place on a larger fraction of the data (Daniël Lakens, personal communication, December 19, 2020).
-In this case, we implemented the O'Brien-Fleming-like GS design with the two interim looks taking place after 50% and 75% of the data, respectively, are collected.
+In this case, I implemented the O'Brien-Fleming-like GS design with the two interim looks taking place after 50% and 75% of the data, respectively, are collected.
 
@@
   *
-1. Randomly sample $N_{text{fixed}}$ observations from each of the two specified populations
+1. Randomly sample $N_{\text{fixed}}$ observations from each of the two specified populations
 2. Run a one-tailed, two-sample *t* test on the mean difference between the two samples
@@
 
-All data, analysis code, and research materials are available at [link]. The study was not preregistered. This manuscript was created using `r cite_r("r-refs.bib")`.
+All data, code, and research materials are available at [link]. The study was not preregistered. This manuscript was created using `r cite_r("r-refs.bib")`.
 
@@
 Table 2A describes the rates of True Negative, Inconclusive, and False Positive results when the true population effect size $\delta$ = 0; Table 2B describes the rates of True Positive, Inconclusive, and False Negative results when the true population effect size $\delta$ = 0.50.
+Figure 4 shows the True Positive, Inconclusive, and False Negative rates for all population effect sizes greater than zero ($\delta$ = 0.2, 0.4, 0.6, 0.8, 1).
 As can be read from the tables, the frequentist procedures all perform as expected: 5% of simulations rejected the null hypothesis when $\delta$ = 0 (i.e., a type I error rate of 5%; see the False Positive column Table 2A) and 80-82% of simulations rejected the null hypothesis when $\delta$ = 0.5 (i.e., a type II error rate of 18-20%; see the True Positive column of Table 2B).
@@
 When the true population effect size is relatively large, the O'Brien-Fleming approach is slightly less efficient than its counterparts.
-This can be explained by the relatively stringent $\alpha_1$ levels of the O'Brien-Fleming procedure in earlier segments (recall that, for the O'Brien-Fleming-like procedure, $\alpha_{1,1}$ = `r alpha_of[1]` and $\alpha_{1,2}$ = `r alpha_of[2]`).
+This can be explained by the relatively stringent $\alpha_1$ levels of the O'Brien-Fleming procedure in the first segment (recall that, for the O'Brien-Fleming-like procedure, $\alpha_{1,1}$ = `r alpha_of[1]`).
 Across all true effect sizes, the Pocock-like procedure, with our chosen parameters, provides the greatest efficiency gains.
@@
 
-In figures A1-A4 in Appendix A, the reader can find that this observation is not merely idiosyncratic to the test chosen (a two-sample *t* test with $\alpha$ = .05 and $\beta = .20$ for a hypothesized population effect size $\delta$ = 0.5).
+In figure A1 in Appendix A, the reader can find that this observation is not merely idiosyncratic to the test chosen (a two-sample *t* test with $\alpha$ = .05 and $\beta = .20$ for a hypothesized population effect size $\delta$ = 0.5).
 In contrast, across dozens of different test specifications, effect size estimates stemming from the ISP are consistently bimodal and systematically exclude the true population effect size from being observed.
@@
 As more observations are collected, the variability in effect size estimates decreases.
+In meta-analyses of findings stemming from cumulative sequential procedures (such as the SBF and the GS designs), less weight will be given to the relatively biased findings stemming from earlier terminations and -- because of their larger sample sizes -- more weight will be given to the relatively unbiased findings stemming from later terminations.
+As a result, the meta-analytic bias in the effect size estimates becomes negligible.
 Unfortunately, this is not the case for the Independent Segments Procedure.
+For the Independent Segments Procedure,  the first $k_{max} - 1$ segments are highly biased due to their low power, which only allows very small or very large effects to filter through.
+The last segment has slightly higher power because a weaker $\alpha$ is set (although this increase in power is less substantial than for cumulative procedures, which receive a marked increase in power due to their accumulating sample size as the study progresses), and thus tends to be slightly less biased.
+However, because each independent segment has equal sample sizes, they are all given equal weight.
+As a result, the overall bias in the effect size estimates remains severe.
 
@@
 As can be read in the Table, the group-sequential median-unbiased estimator, used to adjust for bias in the effect size estimates, provides a remarkable improvement in the accuracy of the parameter estimates.
-In all cases, the group-sequential bias-adjusted effect size estimate is equal to or no more than $d = 0.02$ removed from the true population parameter.
+In all cases, the group-sequential bias-adjusted effect size estimate is equal to or no more than $d$ = 0.02 removed from the true population parameter.
 
@@
 Although this provides a slight improvement in the parameter estimate for certain population effect sizes ($\delta$ = 0.2, 0.5, and 0.6), the Bayesian bias-corrected estimates do not perform as well as the group-sequential bias-corrected estimates.
+The mean of the posterior distribution either slightly improves or, for certain parameters, overcorrects the unadjusted Bayesian effect size estimates.
 
@@
 
+In our simulations, we find that -- for moderate population parameters (0.4 < $\delta$ < 0.6) -- the ISP (powered to detect an effect of $\delta$ = 0.5, with $1 - \beta = 0.8$) severely overestimates effect sizes.
+In Figure A4 in Appendix A, the reader can find that this observation generalizes across many different specifications of statistical power and hypothesized effect size.
+Due to the low power of each independent segment, the critical effect size is higher -- thus, observed significant effects are likely to exaggerate the true population parameter.
+In Figure A4, we find that this exaggeration is severe.
+For example, when $\delta = 0.6$ and power is 80%, the probability of observing a significant effect that is more than 1.5 times the size of the true population parameter (i.e., the probability of observing *d* > 0.9) is 45%, compared to only 15% for a fixed sample hypothesis test.
+
 Figure 7 visualizes the information captured by Table 4, illustrating the differences between the median estimated effect sizes and the true effect sizes.
 For the Sequential Bayes Factor, we can see that the uncorrected estimated slightly underestimate the true population effect size when it is small, and slightly overestimate the true population effect size when it is large.
-The bias-adjusted estimate, the mean of the posterior distribution, corrects for this phenomenon -- but it is important to note that the Bayesian bias-adjusted estimates only provide a slight improvement or, for certain parameters, an overcorrection.
+The bias-adjusted estimate, the mean of the posterior distribution, corrects for this phenomenon -- but, as noted before, the Bayesian bias-adjusted estimates only provide a slight improvement or, for certain parameters, an overcorrection.
 Contrast this with the two bottom row panels for the group-sequential designs.
@@
 As discussed before, the variance in the effect size estimates stemming from the ISP is caused by the process of conducting three independent mini-experiments, rather than a cumulative study.
+
+In Figures A2 and A3 in Appendix A, the reader can observe that our findings regarding the bias, variance, and mean squared error of the Independent Segments Procedure generalize across many different test specifications.
 In the language of decision theory, the ISP is inadmissible: its total squared error risk exceeds that of other hypothesis testing procedures (in our worked example of $\delta$ = 0.5, MSE = `r mean_mse[1]` for the ISP, compared to `r mean_mse[2]` for the Pocock-like design, `r mean_mse[3]` for the O'Brien-Fleming design, `r mean_mse[4]` for the SBF, and `r mean_mse[5]` for the fixed-sample procedure).
 
-To provide more meat to this argument, we separate the error of the procedures by segment in Figure 9.
+To provide more meat to the argument that the error of the ISP is caused by the process of conducting three independent mini-experiments, we separate the error of the procedures by segment in Figure 9.
 Here, we can clearly see that stopping after a first look causes error for all sequential procedures.
-For both the SBF and the Pocock-like procedure, however, error becomes near-negligible in the second and third segment.
+For both the SBF and the group-sequential procedures, however, error becomes near-negligible in the second and third segment.
 As a result, overall error (across all segments) is much lower.
@@
 For the Independent Segments Procedure, as we now know, error remains substantial in every segment, leading to non-negligible overall error.
+We can see that the error for the first $k_{max} -1$ segments is highest, while the error for the last segment is slightly decreased.
+As discussed before, this is due to the slightly higher power of the last segment, for which a weaker $\alpha$ is set.
+This error reduction, however, simply does not compare to the reduction that would have been obtained had the experiment been cumulative.
 
@@
 In deciding between sequential hypothesis testing procedures, researchers need to carefully weigh the advantages and disadvantages of the procedures of interest, with regard to the context in which they are conducting their research [@stefan2020].
+Table 1 presents several of the commonalities and differences between the sequential hypothesis testing procedures.
 
@@
 In the field of clinical trial design, early proposals for fully sequential methods that require analysis after every observation -- such as the Sequential Probability Ratio Test -- did not receive widespread acceptance [@pramanik2021].
-It was only when methods were extended to more realistic settings, and allowed for the evaluation of outcomes after groups of participants were observed, that sequential methods became widely used [@jennison2000, p.
-24; @pramanik2021, pp.
-4-5].
+It was only when methods were extended to more realistic settings, and allowed for the evaluation of outcomes after groups of participants were observed, that sequential methods became widely used [@jennison2000; @pramanik2021].
 
@@
 In our worked example in Figure 1B, we see that the boundary for rejecting $H_0$ at the final look is $\alpha_{1,3}$ = `r alpha_of[3]`.
-To practitioners, this has turned out to be an attractive feature [@jennison2000, p.
-25].
-When using the O'Brien-Fleming correction, it is recommended to collect more than $\displaystyle \frac{1}{k_{max}}$% of the observations before taking the first look, to increases the probability of stopping for efficacy.
+To practitioners, this has turned out to be an attractive feature [@jennison2000].
+When using the O'Brien-Fleming correction, it is recommended to collect more than $\displaystyle \frac{1}{k_{max}}$% of the observations before taking the first look, to increase the probability of stopping for efficacy.
 
 Although the ISP has some advantages, none of these advantages are unique.
-Efficiency and explicit error control can be achieved through alternative methods, none of which pose the disadvantages unique to the ISP.
+Efficiency and explicit error control can be achieved through alternative methods---none of which pose the disadvantages unique to the ISP.
 The ISP applies only to settings in which a researcher wishes to test a single null hypothesis and does not desire a principled decision rule to reject an alternative hypothesis.
 In addition, users of the procedure must guarantee an exact number of participants, allocated equally across segments.
-In realistic settings, however, group sizes and increments in information tend to be unequal and unpredictable [@jennison2000, p.
-26].
+In realistic settings, however, group sizes and increments in information tend to be unequal and unpredictable [@jennison2000].
 Most importantly, the researcher in question should be willing to discard valuable data and abstain from the ability to obtain a credible estimate of effect size.
@@
 Ideally, each individual segment should be powered to at least 50% to detect the hypothesized effect size; otherwise, the critical effect size (the effect size needed to reject the null and end the experiment) will be higher than the hypothesized effect size, and the user will thus systematically discard the effect size which they intended to detect.
-This likely means, however, that the users will have to power the procedure to well over 90% (see Figure A1 in the Appendix); consequently, much of the desired efficiency gain will be lost.
-Given the existence of appealing alternatives that have the same advantages (of efficiently testing hypotheses with explicit error control) as the ISP, without its associated disadvantages (of severe difficulty in effect size estimation), the reader is advised to think twice before utilizing the Independent Segments Procedure.
+Increasing the power of the individual segments likely means, however, that the user will have to power the overall procedure to well over 90% (see Figure A1 in the Appendix).
+Consequently, much of the desired efficiency gains will be lost.
+Given the existence of appealing alternatives that have the same advantages as the ISP (i.e., the ability to efficiently test hypotheses with explicit error control), without its associated disadvantages (i.e., the severe difficulty in effect size estimation), the reader is advised to think twice before utilizing the Independent Segments Procedure.
 
@@
 I show here -- in line with previous work -- that with appropriate design choices, the SBF procedure (implemented with a lower degree of flexibility than in previous investigations) can have comparable error rates to its frequentist alternatives.
-When using the Sequential Bayes Factor procedure, however, researchers are recommended to run simulations [e.g., using the R package `BFDA`, @R-BFDA] to choose an optimal sample size, decision threshold, and prior distribution.
+When using the Sequential Bayes Factor procedure, however, researchers are recommended to run simulations [e.g., using the R package `BFDA`, @R-BFDA] to choose an optimal design.
 
@@
 Despite the trade-off between accurate effect size estimation and efficient hypothesis testing [@goodman2007; @schonbrodt2017], I find that both the SBF procedure and the GS procedure can provide relatively accurate estimates of effect size.
-Thus, the Sequential Bayes Factor and group-sequential procedures can substantially increase research efficiency at only a slight cost to our ability to estimate an effect size.
-For Sequential Bayes Factors, the mean of the posterior distribution shrinks the effect size estimate in case of early terminations when the population effect size is large.
+For Sequential Bayes Factors, the mean of the posterior distribution shrinks the effect size estimate in case of early terminations that rejected the null.
 However, the mean of the posterior distribution does not always provide an improvement: at times, the adjusted estimate is further removed from the true value than the unadjusted estimate.
@@
 Thus, users of group-sequential designs are recommended to always use bias-adjusted estimates for parameter estimation.
+Overall, the Sequential Bayes Factor and group-sequential procedures can substantially increase research efficiency at only a slight cost to our ability to estimate an effect size.
 
@@
 In these contexts, sequential hypothesis testing can provide substantial efficiency gains.
-Group-sequential designs and the Sequential Bayes Factor procedure are valuable tools for psychological scientists to add to their statistical toolbox.
+For researchers who wish to utilize the benefits of sequential hypothesis testing while retaining their ability to obtain a credible effect size estimate, group-sequential designs and Sequential Bayes Factor are an especially good fit.
+Sequential hypothesis testing procedures are valuable tools for psychological scientists to add to their statistical toolbox.
 
@@
 | **Nature of data**  | Cumulative  | Independent  | Cumulative  | Cumulative |
-| **Control of error rates** | Direct | Direct | Indirect (e.g., through simulation) |
-| **Bias-adjusted estimator**| Available | Does not exist | Available |
 | **Flexibility** | Moderately flexible | Highly restrictive | Highly flexible |
 | **Principled rejection of the alternative** | Possible | Impossible | Possible |
+| **Control of error rates** | Direct | Direct | Indirect (e.g., through simulation)|
+| **Bias-adjusted estimator**| Available | Does not exist | Available |
+| **Consideration of prior beliefs**| Absent | Absent | Present and required |
 
@@
 All frequentist procedures were powered to detect an effect of $\delta$ = 0.5.
-For the Sequential Bayes Factor procedure, simulations were run to choose an optimal sample size, decision threshold, and prior distribution, with the goal of approach similar error rates as the frequentist procedures.
+For the Sequential Bayes Factor procedure, simulations were run to choose an optimal sample size, decision threshold, and prior distribution, with the goal of approaching similar error rates as the frequentist procedures.
 
@@
 *Note.* Pocock-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 equally spaced looks.
-Critical values are determined on the basis of Pocock-like $\alpha$- and $\beta$-spending functions with bindings futility bounds.
+Critical values are determined on the basis of Pocock-like $\alpha$- and $\beta$-spending functions with binding futility bounds.
 
@@
 
-*Note.* O'Brien-Fleming-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 unequally spaced looks (the looks were set to 50%, 75%, and 100%, respectively).
+*Note.* O'Brien-Fleming-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 unequally spaced looks (the looks were set to 50%, 75%, and 100% of the data, respectively).
 Critical values are determined on the basis of O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions with binding futility bounds.
@@
 
-*Note.* $BF_{10}$ = 3 indicates that the alternative hypothesis ($H_1:~\delta > 0$) performs three times better at predicting the observed data than the null hypothesis ($H_0:~\delta = 0$).
+*Note.* $BF_{1}$ = 3 indicates that the alternative hypothesis ($H_1:~\delta > 0$) performs three times better at predicting the observed data than the null hypothesis ($H_0:~\delta = 0$).
 Figure based on Jeffreys' Amazing Statistics Program (JASP), adapted from Wagenmakers & Gronau [-@wagenmakers]
@@
 
-*Note.* $BF_{01}$ = 3 indicates that the null hypothesis ($H_0:~\delta = 0$) performs three times better at predicting the observed data than the  alternative hypothesis ($H_1:~\delta > 0$).
+*Note.* $BF_{0}$ = 3 indicates that the null hypothesis ($H_0:~\delta = 0$) performs three times better at predicting the observed data than the  alternative hypothesis ($H_1:~\delta > 0$).
 Figure based on JASP, adapted from Wagenmakers & Gronau [-@wagenmakers]
@@
 
-*Note.*
+*Note.* Figure displays the mean squared error in empirical effect size estimates by segment, and the overall mean squared error across all three segments.
+For the SBF and GS designs, corrected effect size estimates were used to calculate the MSE.
 
@@
 \noindent
-```{r, fig.heigth = 14, fig.width = 6}
+```{r, fig.heigth = 20}
 knitr::include_graphics("appendix/figures/figureA1.tiff", dpi = 1200)

### 2021-11-08 figures/figures.R

@@ -1,8 +1,6 @@
 # TO DO -------------------------------------------------------------------
-# Include both bias-corrected and uncorrected estimates in the manuscript
-# Write up final results
-# Update Notes for figures
 # Write translational abstract
+# Update OSF + GitHub links
 
 # Set up ------------------------------------------------------------------
 library(ggplot2)
@@ -1255,7 +1253,6 @@ ggplot(data = df_summary,
     color = "blue",
     linetype = "dashed"
   )
-  
 dev.off()
 
 # Median bias
@@ -1317,7 +1314,6 @@ ggplot(data = df_summary,
     linetype = "dashed"
   )
 dev.off()
-
 
 #===========================================================================#
 ########################## FIGURE 8: BIAS-VARIANCE ########################## 
@@ -1401,12 +1397,15 @@ dev.off()
 ################ FIGURE 9: MSE CONDITIONAL ON STOPPING TIME #################### 
 #==============================================================================#
 
-sub = df %>% filter(proc != "Fixed" & d_forpower == d_actual) %>% 
+sub = df %>% 
+  filter(proc != "Fixed" & d_forpower == d_actual) %>% 
   group_by(proc, facet, segment, d_actual) %>% 
-  summarize(mse = mean((ES_corrected - d_actual)^2)) %>% ungroup() %>% 
+  summarize(mse = mean((ES_corrected - d_actual)^2)) %>% 
+  ungroup() %>% 
   select(proc, facet, segment, mse) %>% 
   bind_rows(
-    df %>% filter(proc != "Fixed" & d_forpower == d_actual) %>%
+    df %>% #calculate overall MSE (not conditional on segment)
+      filter(proc != "Fixed" & d_forpower == d_actual) %>%
       group_by(proc, facet) %>% 
       summarize(mse = mean((ES_corrected - d_actual)^2)) %>% 
       mutate(segment = rep(4))) #%>% 
@@ -1422,8 +1421,11 @@ names(facet.label) = c(1, 2, 3, 4, 5)
 tiff(file="figures/figure9.tiff",width=2000,height=1500, units = "px", res = 300)
 ggplot(data = sub, mapping = aes(x = as.factor(segment), y = mse)) +
   facet_wrap(vars(facet), labeller = labeller(facet = facet.label), nrow = 2) +
-  geom_segment(aes(x = segment, xend = segment, y = 0, yend = mse),
-               color = "red", linetype = "dashed") +
+  geom_segment(
[…]

### 2021-11-08 manuscript.Rmd
final edits prior to submission, added manuscript PDF and OSF+GitHub links

@@
 author_note: |
-  All simulation data, analysis code, and research materials have been made publicly available on the OSF and GitHub and can be accessed at [link] and [link].
+  All simulation data, analysis code, and research materials have been made publicly available on the OSF and GitHub and can be accessed at osf.io/veczn/ and github.com/shilaan/sequential-testing.
 
@@
 keywords          : "sequential analysis, hypothesis testing, efficiency, estimation, bias"
-wordcount         : "11,387"
+wordcount         : "11,368"
 
@@
 Fast-forward 80 years, and we find ourselves in a decade of unparalleled methods reform in psychological science [@nelson2018].
-At the receiving end of these proposed reforms we often find the statistical power of hypothesis tests [@bakker2012; @button2013]; that is, the long-run probability of observing a statistically significant result ($p < \alpha$) for a given number of subjects and a hypothesized effect size [@morey2016].
+At the receiving end of these proposed reforms we often find the statistical power of hypothesis tests [@bakker2012; @button2013]; that is, the long-run probability of observing a statistically significant result (*p* $< \alpha$) for a given number of subjects and a hypothesized effect size [@morey2016].
 
@@
 
-All data, code, and research materials are available at [link]. The study was not preregistered. This manuscript was created using `r cite_r("r-refs.bib")`.
+All data, code, and research materials are available at osf.io/veczn and github.com/shilaan/sequential-testing.
+The study was not preregistered.
+This manuscript was created using `r cite_r("r-refs.bib")`.
 
@@
 When using the Sequential Bayes Factor procedure, however, researchers are recommended to run simulations [e.g., using the R package `BFDA`, @R-BFDA] to choose an optimal design.
-
 As discussed earlier, the SBF allows researchers to specify their uncertainty regarding the effect size by means of a prior distribution.
@@
 
-*Note.* One-tailed, two-sample *t* tests using, for the frequentist procedures,  $\alpha$ = .05, 1 - $\beta$ = 0.8 for $\delta$ = 0.5.
+*Note.* One-tailed, two-sample *t* tests using, for the frequentist procedures,  $\alpha$ = .05 and 1 - $\beta$ = 0.8 for $\delta$ = 0.5.
 *N*s per group per look are `r n_fixed`, 25, and `r n_p` for the Fixed Sample, Independent Segments, and Pocock-like hypothesis tests.
-For the O'Brien Fleming test, *N* = `r n_of[1]` per group for the first look, and `r n_of[2]` per group for the last two looks.
+For the O'Brien-Fleming test, *N* = `r n_of[1]` per group for the first look, and `r n_of[2]` per group for the last two looks.
 All sequential procedures were set to a maximum number of $k_{max}$ = 3 looks.
-For the Sequential Bayes Factor, data was collected in (at most three) batches of 22 subjects per group, and the procedure was terminated when the one-tailed Bayes Factor was greater than or equal to 3 in favor of the null or the alternative hypothesis.
+For the Sequential Bayes Factor, data was collected in (at most three) batches of 22 subjects per group.
 Figure includes the median effect size estimate (the solid line is the naive estimate; the dashed line is the bias-adjusted estimate).
@@
 ## Figure A1 
-## *Distribution of Obtained Effect Sizes Using Fixed versus Independent Segments Hypothesis Testing*
+## *Distribution of Effect Sizes Using Fixed versus Independent Segments Hypothesis Testing*
 \noindent
@@
 ## Figure A3 
-## *Bias-Variance in Obtained Effect Sizes Using Fixed versus Independent Segments Hypothesis Testing*
+## *Bias-Variance in Effect Sizes Using Fixed versus Independent Segments Hypothesis Testing*
 \noindent

### 2021-11-08 figures/figures.R

@@ -1,7 +1,3 @@
-# TO DO -------------------------------------------------------------------
-# Write translational abstract
-# Update OSF + GitHub links
-
 # Set up ------------------------------------------------------------------
 library(ggplot2)
 library(tidyverse)

### 2021-11-08 manuscript.Rmd
removed redundant figures

@@
 ```{r, fig.heigth = 4.72, fig.width = 6}
-knitr::include_graphics("figures/figure7b.tiff", dpi = 300)
+knitr::include_graphics("figures/figure7.tiff", dpi = 300)
 ```
@@
 ```{r, fig.heigth = 4.72, fig.width = 7}
-knitr::include_graphics("figures/figure8b.tiff", dpi = 300)
+knitr::include_graphics("figures/figure8.tiff", dpi = 300)
 ```
@@
 ```{r, fig.heigth = 20}
-knitr::include_graphics("appendix/figures/figureA1.tiff", dpi = 1200)
+knitr::include_graphics("appendix/figures/figureA1.tiff", dpi = 300)
 ```

### 2021-11-08 figures/figures.R

@@ -1189,67 +1189,6 @@ df_summary <- df_long %>%
     median.bias.sq = (median.ES - d_actual)^2
   ) %>% 
   filter(!((ES_type == "ES_corrected") & (proc == "Fixed" | proc == "ISP"))) #%>% 
-  
-
-# Mean Bias
-
-legend <- data.frame(
-  facet = rep(1),
-  label = c("Uncorrected", "Corrected"),
-  y = c(0.08, 0.05),
-  x = rep(0.1),
-  ES_type = c("ES", "ES_corrected")
-)
-
-tiff(file="figures/figure7a.tiff",width=2300,height=1200, units = "px", res = 300)
-ggplot(data = df_summary, 
-       mapping = aes(x = d_actual, y = bias, color = ES_type)) +
-  theme_bw() +
-  theme(strip.background = element_rect(fill = "white")) +
-  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") + 
-  geom_segment(
-    aes(
-      x = d_actual, 
-      xend = d_actual, 
-      y = 0, 
-      yend = bias, 
-      color = ES_type
-      ),
-    linetype = "dashed"
-    ) + 
-  geom_point() +
-  facet_wrap(vars(facet), labeller = labeller(facet = facet.label)) +
-  theme(legend.position = "none") +
-  scale_x_continuous(
-    breaks = c(-0.2, 0, 0.2, 0.4, 0.6, 0.8, 1),
-    limits = c(-0.2, 1)
-    ) +
-  labs(x = "True Effect Size (Cohen's d)", 
-       y = "Mean Estimated Effect Size - True Effect Size") +
-  scale_color_manual(values = c("red", "blue")) +
-  geom_text(data = legend, 
-            aes(x = x, y = y, label = label, hjust = 0)) + 
-  geom_line(
-    data = data.frame(
-      x = c(-0.2, 0.05),
-      y = rep(0.08),
-      facet = 1
-    ),
-    mapping = aes(x = x, y = y),
-    color = "red",
-    linetype = "dashed"
-  ) +
-  geom_line(
-    data = data.frame(
-      x = c(-0.2, 0.05),
-      y = rep(0.05),
-      facet = 1
-    ),
-    mapping = aes(x = x, y = y),
-    color = "blue",
-    linetype = "dashed"
-  )
-dev.off()
 
 # Median bias
 
@@ -1261,7 +1200,7 @@ legend <- data.frame(
[…]

### 2021-11-13 manuscript.Rmd
coding style

@@
 
-keywords          : "sequential analysis, hypothesis testing, efficiency, estimation, bias"
+keywords          : "sequential analysis, hypothesis testing, efficiency, error rates, parameter estimation, sample size"
 wordcount         : "11,368"
@@
 r_refs(file = "r-refs.bib") #update later
-default_plot_hook = knitr::knit_hooks$get('plot')
+default_plot_hook <- knitr::knit_hooks$get('plot')
 
@@
 gs  <- function(proc, return = "n") {
-  bs = ifelse(proc == "asOF", "bsOF", "bsP") #specify beta-spending function
-  if(proc == "asOF") {rates = c(0.50, 0.75, 1)} else{rates = c(1/3, 2/3, 1)} #specify information rates for OBF vs Pocock
+  bs <- ifelse(proc == "asOF", "bsOF", "bsP") #specify beta-spending function
+  if(proc == "asOF") {
+    rates = c(0.50, 0.75, 1)
+    } else{
+      rates = c(1/3, 2/3, 1)
+      } #specify information rates for OBF vs Pocock
 
@@
   parameters <- getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
-  n_gs = ceiling(c(parameters$numberOfSubjects[1], diff(parameters$numberOfSubjects))/2) # n per look per group
+  n_gs <- ceiling(c(parameters$numberOfSubjects[1], diff(parameters$numberOfSubjects))/2) # n per look per group
   alpha <- parameters$criticalValuesPValueScale
@@
 
-  if(return == "alpha") {return(alpha %>% signif(2) %>% as.character() %>% sub(".", "", .,))}
+  if(return == "alpha") {
+    return(
+      alpha %>% 
+        signif(2) %>% 
+        as.character() %>% 
+        sub(".", "", .,))
+    }
   if(return == "n") {return(n_gs)}
-  if(return == "futility") {return(futility %>% as.character() %>% sub(".", "", .,))}
+  if(return == "futility") {
+    return(
+      futility %>% 
+        as.character() %>% 
+        sub(".", "", .,))
+    }
 }
@@
 This biased estimate remains uncorrected due to the absence of a bias-adjusted estimator accompanying the ISP.
-A we can see, the distribution of effect size estimates stemming from the ISP is bimodal.
+As we can see, the distribution of effect size estimates stemming from the ISP is bimodal.
 
@@
 ```{r include = F}
-#df <- read.csv("simulations/data.csv")
-
-df = df %>% 
+df <- df %>% 
   mutate(
@@
 
-t1 = df %>% filter(d_forpower == d_actual) %>%
+t1 <- df %>% 
+  filter(d_forpower == d_actual) %>%
   group_by(name) %>% 
-  summarize(true_positive = mean(decision == "reject.null") %>% round(2),
+  summarize(
+    true_positive = mean(decision == "reject.null") %>% round(2),
     inconclusive = mean(decision == "inconclusive") %>% round(2),
-            false_negative = mean(decision == "reject.alt") %>% round(2))
-
-t2 = df %>% filter(d_actual == 0) %>% 
+    false_negative = mean(decision == "reject.alt") %>% round(2)
+    )
+
+t2 <- df %>% 
+  filter(d_actual == 0) %>% 
   group_by(name) %>% 
-  summarize(true_negative = mean(decision == "reject.alt") %>% round(2),
+  summarize(
+    true_negative = mean(decision == "reject.alt") %>% round(2),
     inconclusive = mean(decision == "inconclusive") %>% round(2),
-            false_positive = mean(decision == "reject.null") %>% round(2))
+    false_positive = mean(decision == "reject.null") %>% round(2)
+    )
 ```
@@
 ```{r, echo = FALSE}
-knitr::kable(t2, 
+knitr::kable(
+  t2, 
   col.names = c("Procedure", "True Negative", "Inconclusive", "False Positive"), 
-             align = "lccc")
+  align = "lccc"
+  )
 ```
@@
 ```{r, echo = FALSE}
-knitr::kable(t1, 
+knitr::kable(
+  t1, 
   col.names = c("Procedure", "True Positive", "Inconclusive", "False Negative"), 
-             align = "lccc")
+  align = "lccc"
+  )
 ```
@@
 ```{r, echo = FALSE}
-E_n = df %>% 
+E_n <- df %>% 
   group_by(name, d_actual) %>% 
@@
 
-t3 = E_n %>% mutate(E_n = E_n %>% round()) %>% 
-  pivot_wider(names_from = d_actual, names_prefix = "*d* = ", values_from = E_n)
+t3 <- E_n %>%
+  mutate(
+    E_n = E_n %>% round()
+    ) %>% 
+  pivot_wider(
+    names_from = d_actual, 
+    names_prefix = "*d* = ", 
+    values_from = E_n
+    )
 
@@
   group_by(name, d_actual, ES_type) %>% 
-  summarize(medianES = round(median(ES), 2)) %>% 
+  summarize(
+    medianES = round(median(ES), 2)
+    ) %>% 
   rename(Procedure = name) %>% 
@@
     )
-
-t4
 ```

### 2021-11-13 appendix/figures/figures_appendix.R

@@ -8,16 +8,20 @@ library(grid)
 library(gridExtra)
 
 # Read data ---------------------------------------------------------------
-zipped_df = gzfile("appendix/simulations/data_appendix.csv.gz", "rt")
-df = read.csv(zipped_df, header = T) #takes a minute
+zipped_df <- gzfile("appendix/simulations/data_appendix.csv.gz", "rt")
+df <- read.csv(zipped_df, header = T) #takes a minute
 
-df_summary = df %>% 
+df_summary <- df %>% 
   group_by(d_forpower, power, proc, d_actual) %>%  
-  summarize(mean_ES = mean(ES),
-            mse = mean((ES - d_actual)^2),
-            var = mean((ES - mean(ES))^2)) %>% 
-  mutate(bias = mean_ES - d_actual,
-         bias2 = (mean_ES - d_actual)^2) 
+  summarize(
+    mean_ES = mean(ES),
+    mse = mean((ES - d_actual)^2),
+    var = mean((ES - mean(ES))^2)
+    ) %>% 
+  mutate(
+    bias = mean_ES - d_actual,
+    bias2 = (mean_ES - d_actual)^2
+    ) 
 
 # Figure A1 ----------------------------------------------------------------
 # Obtained Effect Sizes using Fixed versus Independent Segments Hypothesis Testing
@@ -26,10 +30,10 @@ df_summary = df %>%
 df_expected <- df %>% 
   filter(d_forpower == d_actual)
 
-plt = function(f){
-  D = mean(f$d_actual) #get separate plot for each combo of power and d
-  target_power = mean(f$power)
-  t = sprintf("\u03b4 = %0.1f, 1 - *\u03b2* = %0.1f", D, target_power)
+plt <- function(f){
+  D <- mean(f$d_actual) #get separate plot for each combo of power and d
+  target_power <- mean(f$power)
+  t <- sprintf("\u03b4 = %0.1f, 1 - *\u03b2* = %0.1f", D, target_power)
   
   ggplot(data = f, aes(x = ES, group = proc, linetype = proc)) +
     geom_density() +
@@ -43,22 +47,49 @@ plt = function(f){
       axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1), 
       axis.ticks.y = element_blank()
       ) +
-    scale_x_continuous(breaks = c(0, 0.5*D, D, 1.5*D),
-                       limits = c(-D, 2.5*D)) +
[…]

### 2021-11-13 appendix/simulations/simulations_appendix.R

@@ -8,16 +8,20 @@ library(TOSTER)
 library(here)
 
 ########################## GLOBAL HELPER FUNCTIONS ##########################
-generate_data = function(nsims, n, d) { 
+generate_data <- function(nsims, n, d) { 
   # Generate raw data
-  raw = lapply(1:nsims, function(i) {
-    data.table(nsim = i, n = n, d = d,
-               m2 = rnorm(n = n, mean = 0),
-               m1 = rnorm(n = n, mean = 0) + d)})
+  raw <- lapply(1:nsims, function(i) {
+    data.table(
+      nsim = i, 
+      n = n, 
+      d = d,
+      m2 = rnorm(n = n, mean = 0),
+      m1 = rnorm(n = n, mean = 0) + d
+      )})
   data.table::rbindlist(raw)
 }
 
-run_tests = function(df, n, proc, segment = 1, alpha) {
+run_tests <- function(df, n, proc, segment = 1, alpha) {
   
   # Run t-tests and create data-frame
   df %>% 
@@ -39,34 +43,34 @@ run_tests = function(df, n, proc, segment = 1, alpha) {
 } 
 
 ########################## GLOBAL PARAMETERS ##########################
-max_n_segments = 3
-alpha_total = 0.05
+max_n_segments <- 3
+alpha_total <- 0.05
 
 # Simulations for appendix data ---------------------------------------
 
 ########################## FULL PROCEDURE FUNCTION ##########################
 
-run_all_procedures = function(nsims){
+run_all_procedures <- function(nsims){
   
   ########################## FIXED SAMPLING PROCEDURE FUNCTION ########################## 
   
-  run_fixed = function(d_actual, d_forpower, target_power) {
+  run_fixed <- function(d_actual, d_forpower, target_power) {
     
     ########################## FIXED SAMPLE ##########################
     
     # Calculate fixed n per group for the one-tailed two-sample t-test
-    n_fixed = ceiling(
+    n_fixed <- ceiling(
       power.t.test(
         delta = d_forpower, 
-        #delta = d_actual, 
         sig.level = alpha_total, 
         power = target_power, 
         type = "two.sample", 
-        alternative = "one.sided")$n)
+        alternative = "one.sided"
+        )$n)
     
     # RUN FIXED SAMPLE NHST
     set.seed(1234*(d_actual*target_power))
-    fixed = generate_data(
+    fixed <- generate_data(
       nsims = nsims,
[…]

### 2021-11-13 figures/figures.R

@@ -23,10 +23,10 @@ df <- df %>%
   )
 
 facet.label <- c("Fixed Hypothesis and Equivalence Test", 
-                "Sequential Bayes Factor", 
-                "Independent Segments Procedure", 
-                "Pocock-like GS design",
-                "O'Brien-Fleming-like GS design")
+                 "Sequential Bayes Factor", 
+                 "Independent Segments Procedure", 
+                 "Pocock-like GS design",
+                 "O'Brien-Fleming-like GS design")
 names(facet.label) <- c(1, 2, 3, 4, 5)
 
 #=============================================================================#
@@ -52,7 +52,7 @@ gs <- function(proc, return = "n") {
   
   # Get parameters
   parameters <- getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
-  n_gs = ceiling(c(parameters$numberOfSubjects[1], diff(parameters$numberOfSubjects))/2) # n per look per group
+  n_gs <- ceiling(c(parameters$numberOfSubjects[1], diff(parameters$numberOfSubjects))/2) # n per look per group
   alpha <- parameters$criticalValuesPValueScale
   futility <- parameters$futilityBoundsPValueScale
   
@@ -93,6 +93,7 @@ futility_ofd <- round(
 d.plot <- function(n, alpha, t) {
   df <- 2*n-2
   critical_d <- round(qt(1-alpha, df = 2*n-2) * sqrt(2*(1/n)),2)
+  
   p <- ggplot(data.frame(x = c(-4, 4)), aes(x)) +
     stat_function(
       fun = dt, 
@@ -122,7 +123,6 @@ d.plot <- function(n, alpha, t) {
     geom_vline(xintercept = qt(1-alpha, df = df), linetype = 2, color = "grey") 
   return(p)
 }
-
 
 p1 <- d.plot(n = n_p, alpha = alpha_p[1], t = "Pocock (First Look)") +
   scale_x_continuous(
@@ -346,10 +346,18 @@ p3 <- d.plot(
     )
 
 tiff(file="figures/figure1a.tiff",width=2500,height=800, units = "px", res = 300)
-grid.arrange(p1, p2, p3, nrow = 1, 
-             left = textGrob("Density under H0: \u03b4 = 0", 
-                             rot = 90, hjust = 0.57,
-                             gp = gpar(fontsize = 10)))
+grid.arrange(
+  p1, 
[…]

### 2021-11-13 simulations/isp-helpers.R

@@ -1,23 +1,21 @@
 ####ISP Engine provided by Ulrich & Miller at https://github.com/milleratotago/Independent_Segments_R
 
-#setwd("/Users/shilaan/Library/Mobile Documents/com~apple~CloudDocs/PhD/ISP")
-
 # Class SegmentedHypTestEngine
 
-# Runs (computes the proababilistic results of) complete segmented hypothesis testing scenarios
+# Runs (computes the probabilistic results of) complete segmented hypothesis testing scenarios
 
-#########################################################################################################
+##########################################
 # Creation methods - following Wickham
-#########################################################################################################
-#========================================================================================================
+##########################################
+#=========================================
 # new_SegmentedHypTestEngine
-#========================================================================================================
+#=========================================
+
 new_SegmentedHypTestEngine <- function()
 {
   
   # No fields needed at creation
   new_SegmentedHypTestEngine <- list()
-  
   
   attr(new_SegmentedHypTestEngine, "class") <- "SegmentedHypTestEngine"
   
@@ -25,22 +23,22 @@ new_SegmentedHypTestEngine <- function()
   return(new_SegmentedHypTestEngine)
   
 } # new_SegmentedHypTestEngine
-#========================================================================================================
+#=========================================
 
-#========================================================================================================
+#=========================================
 # validate_SegmentedHypTestEngine
-#========================================================================================================
+#=========================================
 # Checks additional properties of elements, for example, that values are in allowed ranges -- business rules
 validate_SegmentedHypTestEngine <- function(segmented_hyp_test_engine)
 {
   # More elaborate checking in here if needed. Stop if things go wrong.
   
 } # end validate_SegmentedHypTestEngine
-#========================================================================================================
+#=========================================
 
-#========================================================================================================
+#=========================================
 # SegmentedHypTestEngine - Public ctor
-#========================================================================================================
+#=========================================
 #-----------------
 # Roxygen comments
 
@@ -65,16 +63,16 @@ SegmentedHypTestEngine <- function()
   return(instance)
   
 } # end SegmentedHypTestEngine ctor
-#========================================================================================================
+#=========================================
 
 
-#########################################################################################################
+##########################################
 # Public methods
-#########################################################################################################
+##########################################
 #' @export
-#========================================================================================================
+#=========================================
[…]

### 2021-11-13 simulations/simulations.R

@@ -13,16 +13,21 @@ library(NCmisc)
 library(here)
 
 ########################## GLOBAL HELPER FUNCTIONS ##########################
-generate_data = function(nsims, n, d) { 
+generate_data <- function(nsims, n, d) { 
   # Generate raw data
-  raw = lapply(1:nsims, function(i) {
-    data.table(nsim = i, n = n, d = d,
-               m2 = rnorm(n = n, mean = 0),
-               m1 = rnorm(n = n, mean = 0) + d)})
+  raw <- lapply(1:nsims, function(i) {
+    data.table(
+      nsim = i, 
+      n = n, 
+      d = d,
+      m2 = rnorm(n = n, mean = 0),
+      m1 = rnorm(n = n, mean = 0) + d
+      )}
+    )
   data.table::rbindlist(raw)
 }
 
-run_tests = function(df, n, proc, segment = 1, alpha) {
+run_tests <- function(df, n, proc, segment = 1, alpha) {
   
   # Run t-tests and create data-frame
   df %>% 
@@ -44,33 +49,35 @@ run_tests = function(df, n, proc, segment = 1, alpha) {
 } 
 
 ########################## GLOBAL PARAMETERS ##########################
-max_n_segments = 3
-alpha_total = 0.05
-target_power = 0.8 #might also add 0.9
-d_forpower = 0.5
+max_n_segments <- 3
+alpha_total <- 0.05
+target_power <- 0.8 
+d_forpower <- 0.5
 
 ########################## FULL PROCEDURE FUNCTION ##########################
 
-run_all_procedures = function(nsims){
+run_all_procedures <- function(nsims){
   
   ########################## FIXED SAMPLING PROCEDURE FUNCTION ########################## 
   
-  run_fixed = function(d_actual) {
+  run_fixed <- function(d_actual) {
     
     ########################## FIXED SAMPLE ##########################
     
     # Calculate fixed n per group for the one-tailed two-sample t-test
-    n_fixed = ceiling(
+    n_fixed <- ceiling(
       power.t.test(
         delta = d_forpower, 
         sig.level = alpha_total, 
         power = target_power, 
         type = "two.sample", 
-        alternative = "one.sided")$n)
+        alternative = "one.sided"
+        )$n
+      )
     
     # RUN FIXED SAMPLE NHST
     set.seed(1234*(d_forpower*target_power))
-    fixed = generate_data(
+    fixed <- generate_data(
       nsims = nsims,
       n = n_fixed, 
[…]
